### 2020-09-18 paper/paper.tex
Add scaffold of tex files

% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.20 of 2017/10/04
%
\documentclass[runningheads]{llncs}
%
\usepackage{graphicx}
% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following line
% to display URLs in blue roman font according to Springer's eBook style:
% \renewcommand\UrlFont{\color{blue}\rmfamily}

\begin{document}
%
\title{Detecting Fair Queuing to Enable Low-Latency and High-Throughput Congestion Control}
%
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Maximilian Bachl\inst{1}\orcidID{0000-0001-9482-7428} \and
Joachim Fabini\inst{2,3}\orcidID{0000-0002-8285-1591} \and
Tanja Zseby\inst{3}\orcidID{0000-0002-5391-467X}}
%
\authorrunning{M. Bachl et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{TU Wien, Vienna, Austria
\email{firstname.lastname@tuwien.ac.at}}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
Low delay is an explicit requirement for applications such as cloud gaming and video conferencing. Delay-based congestion control can achieve the same throughput but significantly smaller delay than loss-based one and is thus ideal for these scenarios. However, when a delay- and a loss-based flow share a bottleneck, the loss-based one can ``steal'' all the bandwidth and the delay-based one ``starves''. Fair queuing at the bottleneck link, which isolates each flow and provides the same bandwidth for everyone, solves this problem. However, so far no end host based algorithm to detect fair queuing exists. Our contribution is the development of an algorithm that detects fair queuing at flow startup and chooses delay-based congestion control in this case. Otherwise, loss-based congestion control can be used as a backup option. Results show that our algorithm reliably detects fair queuing and can decide between delay- and loss-based congestion control to enable better congestion control.

\keywords{Congestion Control  \and Active Measurements \and Fair Queuing \and Queue Management.}
\end{abstract}
%
%
%
\section{Introduction}

Bla.

\bibliography{fq_detection}

\end{document}

### 2020-09-18 src/plot_rtt_and_bandwidth.py

@@ -98,7 +98,7 @@ plt.plot(*zip(*with_correct_time))
 plt.tight_layout()
 
 plt.ylim(bottom=0)
-plt.savefig(f"plots/throughput_{interval}_{('_'.join(pcap_file.split('_')[1:]))}.pdf", bbox_inches = 'tight', pad_inches = 0)
+# plt.savefig(f"plots/throughput_{interval}_{('_'.join(pcap_file.split('_')[1:]))}.pdf", bbox_inches = 'tight', pad_inches = 0)
 plt.savefig(f"plots/throughput_{interval}_{('_'.join(pcap_file.split('_')[1:]))}.png", bbox_inches = 'tight', pad_inches = 0, dpi=200)
 
 plt.close()
@@ -116,7 +116,7 @@ plt.plot(ack_timestamps, rtts)
 plt.tight_layout()
 
 plt.ylim(bottom=0)
-plt.savefig(f"plots/rtt_{interval}_{('_'.join(pcap_file.split('_')[1:]))}.pdf", bbox_inches = 'tight', pad_inches = 0)
+# plt.savefig(f"plots/rtt_{interval}_{('_'.join(pcap_file.split('_')[1:]))}.pdf", bbox_inches = 'tight', pad_inches = 0)
 plt.savefig(f"plots/rtt_{interval}_{('_'.join(pcap_file.split('_')[1:]))}.png", bbox_inches = 'tight', pad_inches = 0, dpi=200)
 # plt.show()
 

### 2020-09-18 src/test.py

@@ -296,10 +296,10 @@ elif opt.run_scenario == "accuracy":
 elif opt.run_scenario == "competing_flow":
 	opt.competing_flow = True
 
-	opt.time = 10
+	opt.time = 20
 	opt.store_pcaps = True
-	opt.buffer_size = 50
-	opt.rate = 8
+	opt.buffer_size = 100
+	opt.rate = 50
 	opt.delay = 10
 
 	try:
@@ -319,9 +319,6 @@ elif opt.run_scenario == "competing_flow":
 	opt.two_iperfs = False
 	with virtnet.Manager() as context:
 		client_output = run(context, "competing_flow_fq")
-	quit()
-
-
 
 	try:
 		del os.environ["START_VEGAS"]
@@ -382,9 +379,9 @@ elif opt.run_scenario == "competing_flow":
 
 elif opt.run_scenario == "just_one_flow":
 
-	opt.time = 30
+	opt.time = 20
 	opt.store_pcaps = True
-	opt.buffer_size = 50
+	opt.buffer_size = 100
 	opt.rate = 10
 	opt.delay = 10
 

### 2020-09-24 paper/figures/fq_common_end.tex
Updated paper

%\draw[draw,decorate,decoration={brace,mirror}] (0.75,0) -- (1.75,0) node [inner sep=0, outer sep=0,midway,below,yshift=-1.25mm,font=\fontsize{3pt}{0}\selectfont] {interval};
%\draw[densely dotted] (1.75,0) -- (1.75,0.95);

\newcommand{\xaxislength}{1.8}

\draw[->, line width=0.05mm, -dot3-={0.5 cm}{1 rtt}, -dot3-={1.0 cm}{2 rtts}, -dot3-={1.5 cm}{3 rtts}] (0,0) -- ++(\xaxislength , 0);
\node[align=center,font=\fontsize{3pt}{0}\selectfont,inner sep=0mm] at (1.65, 0.05) (time) {time};

\draw[->, line width=0.05mm] (0,0) -- ++(0,2.35);
\draw[draw=none, line width=0.05mm] (-0.05,-0.0) -- ++(0,1);
\draw[draw=none, line width=0.05mm] (0,-0.05) -- ++(0,1);

\node (red_label) at (0.1,1.95) [inner sep=0,fill=red,rectangle,minimum width=0.1cm,minimum height=0.1cm,label={[align=left,label distance=-0.75mm]right:\fontsize{1mm}{0mm}\selectfont{}flow 2}] {};

\node (green_label) at (0.1,1.8) [inner sep=0,fill=green,rectangle,minimum width=0.1cm,minimum height=0.1cm,label={[align=left,label distance=-0.75mm]right:\fontsize{1mm}{0mm}\selectfont{}flow 1}] {};

\end{tikzpicture}
\end{document}

### 2020-09-24 paper/figures/fq_illustration_goodput_fq.tex
Updated paper

\input{fq_common_beginning.tex}

%\draw[draw,decorate,decoration={brace, mirror}] (2.5,0) -- ++(0.0,1) node [midway,right,align=left,font=\fontsize{3pt}{0}\selectfont] {buffer\\ size};
%
%\draw[dashed] (0,0) -- ++(2.5,0);
%\node[align=center,font=\fontsize{3pt}{0}\selectfont,inner sep=0mm] at (1.95,-0.25) (optimal) {optimal minimum\\ window (BDP)};
%\draw[->, line width=0.05mm] (optimal) -- ++(0,0.25);

\node at (\barwidth /2, \barheight /2) [inner sep=0,fill=green,rectangle,minimum width=\barwidth cm,minimum height=\barheight cm] {};
\node at (\barwidth /2, 2*\barheight) [inner sep=0,fill=red,rectangle,minimum width=\barwidth cm,minimum height=2*\barheight cm] {};

\node at (\barwidth /2 + \barwidth, \barheight) [inner sep=0,fill=green,rectangle,minimum width=\barwidth cm,minimum height=2*
\barheight cm] {};
\node at (\barwidth /2 + \barwidth, 4*\barheight) [inner sep=0,fill=red,rectangle,minimum width=\barwidth cm,minimum height=4*\barheight cm] {};

\node at (\barwidth /2 + 2*\barwidth, 0.3) [inner sep=0,fill=green,rectangle,minimum width=\barwidth cm,minimum height=0.6cm] {};
\node at (\barwidth /2 + 2*\barwidth, 0.9) [inner sep=0,fill=red,rectangle,minimum width=\barwidth cm,minimum height=0.6cm] {};

\draw[dashed] (0,1.2) -- ++(1.8,0);
\node[align=center,font=\fontsize{2.75pt}{0}\selectfont,inner sep=0mm] at (0.375, 1.35) (maximum) {link speed};
\draw[->, line width=0.05mm] (maximum) -- (0.375,1.2);

\node[anchor=north west, align=center,font=\fontsize{3pt}{0}\selectfont,inner sep=0mm] at (0.01, 2.3) (rate) {receiving rate};

\input{fq_common_end.tex}

### 2020-09-24 paper/figures/fq_illustration_goodput_no_fq.tex
Updated paper

--- 2020-09-24 paper/figures/fq_illustration_goodput_fq.tex
@@
 
-\node at (\barwidth /2 + 2*\barwidth, 0.3) [inner sep=0,fill=green,rectangle,minimum width=\barwidth cm,minimum height=0.6cm] {};
-\node at (\barwidth /2 + 2*\barwidth, 0.9) [inner sep=0,fill=red,rectangle,minimum width=\barwidth cm,minimum height=0.6cm] {};
+\node at (\barwidth /2 + 2*\barwidth, 0.2) [inner sep=0,fill=green,rectangle,minimum width=\barwidth cm,minimum height=0.4 cm] {};
+\node at (\barwidth /2 + 2*\barwidth, 0.8) [inner sep=0,fill=red,rectangle,minimum width=\barwidth cm,minimum height=0.8 cm] {};
 

### 2020-09-24 paper/figures/fq_illustration_throughput.tex
Updated paper

--- 2020-09-24 paper/figures/fq_illustration_goodput_fq.tex
@@
 
-\node at (\barwidth /2 + 2*\barwidth, 0.3) [inner sep=0,fill=green,rectangle,minimum width=\barwidth cm,minimum height=0.6cm] {};
-\node at (\barwidth /2 + 2*\barwidth, 0.9) [inner sep=0,fill=red,rectangle,minimum width=\barwidth cm,minimum height=0.6cm] {};
+\node at (\barwidth /2 + 2*\barwidth, 2*\barheight) [inner sep=0,fill=green,rectangle,minimum width=\barwidth cm,minimum height=4*
+\barheight cm] {};
+\node at (\barwidth /2 + 2*\barwidth, 8*\barheight) [inner sep=0,fill=red,rectangle,minimum width=\barwidth cm,minimum height=8*\barheight cm] {};
 
-\draw[dashed] (0,1.2) -- ++(1.8,0);
-\node[align=center,font=\fontsize{2.75pt}{0}\selectfont,inner sep=0mm] at (0.375, 1.35) (maximum) {link speed};
-\draw[->, line width=0.05mm] (maximum) -- (0.375,1.2);
+\node (green_label) at (0.1,1.8) [inner sep=0,fill=green,rectangle,minimum width=0.1cm,minimum height=0.1cm,label={[align=left,label distance=-0.75mm]right:\fontsize{1mm}{0mm}\selectfont{}flow 1}] {};
 
-\node[anchor=north west, align=center,font=\fontsize{3pt}{0}\selectfont,inner sep=0mm] at (0.01, 2.3) (rate) {receiving rate};
+\node[anchor=north west, align=center,font=\fontsize{3pt}{0}\selectfont,inner sep=0mm] at (0.01, 2.3) (rate) {sending rate};
 

### 2020-09-24 paper/paper.tex
Updated paper

@@
 %
+\usepackage{subfig}
 \usepackage{graphicx}
+\usepackage{cite}
+\usepackage{hyperref}
+\usepackage{booktabs}
+\usepackage[english]{babel}
+\usepackage{xcolor}
+\usepackage[nomain, toc, acronym]{glossaries}
 % Used for displaying a sample figure. If possible, figure files should
@@
 % to display URLs in blue roman font according to Springer's eBook style:
-% \renewcommand\UrlFont{\color{blue}\rmfamily}
+\renewcommand\UrlFont{\color{blue}\rmfamily}
+
+\newacronym{cc}{CC}{Congestion Control}
+\newacronym{cca}{CCA}{Congestion Control Algorithm}
+\newacronym{aqm}{AQM}{Active Queue Management}
+\newacronym{rtt}{RTT}{Round Trip Time}
+\newacronym{fq}{FQ}{Fair Queuing}
+\newacronym{bdp}{BDP}{Bandwidth Delay Product}
+\newacronym{qoe}{QoE}{Quality of Experience}
+
+\usepackage{amsmath,amssymb,amsfonts}
+\newcommand{\mynote}[3]{
+    \fbox{\bfseries\sffamily\scriptsize#1}
+    {\small$\blacktriangleright$\textsf{\emph{\color{#3}{#2}}}$\blacktriangleleft$}}
+
+\newcommand{\todo}[1]{\mynote{TODO}{#1}{red}}
+\newcommand{\noteMax}[1]{\mynote{Max}{#1}{green}}
 
@@
 %
-\title{Detecting Fair Queuing to Enable Low-Latency and High-Throughput Congestion Control}
+\title{Detecting Fair Queuing to Enable Better Congestion Control}
 %
@@
 %
-\institute{TU Wien, Vienna, Austria
+\institute{TU Wien, Vienna, Austria; 
 \email{firstname.lastname@tuwien.ac.at}}
@@
 Otherwise, loss-based congestion control can be used as a backup option.
-Results show that our algorithm reliably detects fair queuing and can decide between delay- and loss-based congestion control to enable better congestion control.
+Results show that our algorithm reliably detects fair queuing and can achieve low delay and high throughput in case fair queuing is detected.
 
@@
 \section{Introduction}
+\label{sec:introduction}
 
-Bla.
+Emerging applications such as cloud gaming \cite{jarschel_evaluation_2011} and virtual reality \cite{elbamby_toward_2018} applications require high throughput and low delay.
+Furthermore, for 5G, ultra low latency communications have been made a priority \cite{li_5g_2018}.
+We argue that for achieving high throughput as well as low delay, congestion control must be taken into account.
 
+\textit{Delay-based} \glspl{cca} were proposed to provide high throughput and low delay for network flows.
+They use the queuing delay as an indication of congestion and lower their throughput as soon as the delay increases.
+Such approaches have been proposed in the last century \cite{brakmo_tcp_1995} and recently have seen a surge in popularity \cite{arun_copa_2018,hock_tcp_2017,mittal_timely_2015,cardwell_bbr:_2016}.
+While loss-based \glspl{cca} fulfill their goal of low delay and high throughput, they cannot handle competing flows with different \glspl{cca} well.
+This is especially prevalent when they have to compete against loss-based \glspl{cca}.
+The problem is that loss-based \glspl{cca} are more ``aggressive''.
+While the delay-based ones already back away as soon as queuing delay increases, the loss-based ones continue increasing their throughput until the buffer is full and packet loss occurs.
+This results in the loss-based flows gaining all the bandwidth and the delay-based ones ``starving'' \cite{hock_toward_2016,yuan-cheng_lai_improving_2001,awdeh_comparing_2004}.
+
+This unfairness can be mitigated by using \gls{fq} at the bottleneck link, isolating each flow from all other flows and assigning each flow an equal share of bandwidth \cite{dumazet_pkt_sched:_2013}.
+Then, the loss-based flows cannot obtain the bandwidth of the loss-based flows anymore.
+In this case, delay-based \glspl{cca} fulfill their goal of high bandwidth and low delay.
+
+While \gls{fq} solves many problems regarding \gls{cc}, it is still not ubiquitously deployed in every bottleneck link.
+If, however, flows could know if there's \gls{fq} or not, they could dynamically adapt their congestion control.
+If there is \gls{fq} they would use a delay-based \gls{cca} and otherwise a loss-based one.
+Our contribution is the design and evaluation of such a mechanism.
+This mechanism determines the presence of \gls{fq} during the startup phase of a flow's \gls{cca} and, at the end of the startup, sets the \gls{cca} accordingly.
+
+We show that this solution reliably determines the presence of fair queuing and that by using this approach, queuing delay can be considerably lowered while keeping throughput high.
+While our mechanism is beneficial for network flows in general, we argue that it is most useful for long-running delay-sensitive flows, such as cloud gaming, video conferencing etc.
+
+To make our work reproducible and to encourage further research, we make the code, the results and the figures publicly available\footnote{\url{https://github.com/CN-TU/PCC-Uspace}}.
+
+\section{Related work}
+
+\section{Concept}
+\label{sec:concept}
+
+The overall concept is the following:
+\begin{enumerate}
+\item During flow startup, determine if \gls{fq} is used at the bottleneck or not.
+\item If this is determined, change to a delay-based \gls{cca} if \gls{fq} is used and to a loss-based one otherwise.
+\end{enumerate}
+
+\subsection{Algorithm to determine the presence of \gls{fq}}
+
+%\begin{figure}
+%\centering
+%\includegraphics[width=1\columnwidth]{{"figures/fq_illustration_throughput"}.pdf}
+%\caption{Sending rate}
+%\label{fig:throughput}
+%\end{figure}
+%
+%\begin{figure}
+%\includegraphics[width=1\columnwidth]{{"figures/fq_illustration_goodput_no_fq"}.pdf}
+%\caption{Receiving rate (goodput) in case the bottleneck link doesn't use \gls{fq}. Flow 2, which sends more also manages to get more data through the bottleneck.}
+%\label{fig:goodput_no_fq}
+%\end{figure}
+%
+%\begin{figure}
+%\includegraphics[width=1\columnwidth]{{"figures/fq_illustration_goodput_fq"}.pdf}
+%\caption{Receiving rate (goodput) in case the bottleneck link uses \gls{fq}. Flow 2, which sends more gets as much data through the bottleneck as flow 1.}
+%\label{fig:goodput_fq}
+%\end{figure}
+
+\begin{figure}
+\centering
+\subfloat[Sending rate]{\includegraphics[width=0.33\columnwidth]{{"figures/fq_illustration_throughput"}.pdf}
+\label{fig:throughput}}
+\subfloat[Receiving rate (no \gls{fq}).]{\includegraphics[width=0.33\columnwidth]{{"figures/fq_illustration_goodput_no_fq"}.pdf}
+\label{fig:goodput_no_fq}}
+\subfloat[Receiving rate (\gls{fq}).]{\includegraphics[width=0.33\columnwidth]{{"figures/fq_illustration_goodput_fq"}.pdf}
+\label{fig:goodput_fq}}
+\caption{An example illustrating our proposed flow startup mechanism.
+Figure \ref{fig:throughput} shows the sending rate, \ref{fig:goodput_no_fq} the receiving rate in case there's no \gls{fq} and \ref{fig:goodput_fq} the receiving rate if there is \gls{fq}.}
+\label{fig:illustration}
+\end{figure}
+
+The algorithm to determine the presence of \gls{fq} works as follows (seen from the sender's point of view):
+\begin{enumerate}
+\item Launch two concurrent flows, flow 1 and flow 2. Initialize flow 1 with some starting sending rate and flow 2 with two times that rate. 
+\item After every \gls{rtt}, if no packet loss occurred, double the sending rate of both flows. 
+\item If packet loss occurred in both flows in the previous \gls{rtt}, calculate the following metric: 
+\begin{align}
+\textit{loss ratio} &= \frac{\left(\frac{\textit{sending rate of flow 1}}{\textit{receiving rate of flow 1}}\right)}{\left(\frac{\textit{sending rate of flow 2}}{\textit{receiving rate of flow 2}}\right)} \\
+&= \frac{\textit{sending rate of flow 1}\times \textit{receiving rate of flow 2}}{\textit{receiving rate of flow 1} \times \textit{sending rate of flow 2}}
+\end{align}
+Basically, this metric indicates if flow 2, which has a higher sending rate, also achieves a higher receiving rate (goodput) than flow 1. If this is the case, there's no \gls{fq}, otherwise there is. 
+\item If \gls{fq} is used, the $\textit{loss ratio}$ is 2, otherwise it is 1. We choose 1.5 as the cutoff value. 
+\end{enumerate}
+
+Figure~\ref{fig:throughput}, \ref{fig:goodput_no_fq} and \ref{fig:goodput_fq} schematically illustrate this mechanism.
+We assume a sender being connected to a receiver, with a bottleneck link between them.
+On this bottleneck link, the queue is either managed by \gls{fq} or by a drop tail buffer (no \gls{fq}).
+\autoref{fig:throughput} shows how the sending rate before the bottleneck link is doubled after each \gls{rtt}.
+Furthermore, flow 2 has double the sending rate of flow 1 \noteMax{Figures take up too much space.
+Is it better to put them next to each other?
+The actual problem is the nonsensical Springer format.}.
+
+Figure~\ref{fig:goodput_no_fq} shows the receiving rate at the receiver after the bottleneck link, if this bottleneck link has a drop tail buffer without \gls{fq}.
+In the third \gls{rtt} the sending rate exceeds the link speed of the bottleneck and thus the receiving rate is smaller than the sending rate.
+Because no \gls{fq} is employed, flow 2 manages to have a receiving rate two times the one of flow 1.
+
+Figure~\ref{fig:goodput_fq}, on the other side, shows the receiving rate of the two flows after the bottleneck link in case \gls{fq} is deployed.
+In the third \gls{rtt}, flow 1 and flow 2 achieve the same receiving rate because of the \gls{fq}.
+
+\subsection{Simple delay-based \gls{cc}}
+
+If \gls{fq} is detected, a simple delay-based \gls{cc} is used from then on. Our algorithm is based on the two following simple rules:
+\begin{enumerate}
+\item If $\textit{current rtt} \leq \textit{smallest rtt ever measured on the connection} + 5\textit{ms}$ then increase the sending rate by 1\%.
+\item Otherwise ($\textit{current rtt} > \textit{smallest rtt ever measured on the connection} + 5\textit{ms}$) decrease the sending rate by 5\%.
+\end{enumerate}
+
+This \gls{cca} is meant as a simple proof-of-concept to demonstrate the efficacy of our \gls{fq} detection mechanism but we do not claim it to be optimal in any way.
+
+\subsection{Fallback loss-based \gls{cc}}
+
+If the absence of \gls{fq} detected, a loss based \gls{cc} is used as a fallback.
+We use the algorithm \textit{PCC} proposed by \cite{dong_pcc_2015}.
+In simple terms, this \gls{cca} aims to maximize a utility function, which strives for high throughput and low packet loss.
+However, it is reasonably aggressive and does not starve when competing with other loss-based \glspl{cca} such as Cubic \cite{ha_cubic_2008}.
+
+\section{Implementation}
+
+We base our implementation on the code of \textit{PCC Vivace} \cite{dong_pcc_2018}, which in turn is an improved version of the \cite{dong_pcc_2015}.
+The PCC code is based on the \textit{UDT} library \cite{gu_udt_2007}, which is a library that provides a reliable transport protocol similar to TCP on top of UDP.
+This approach is similar to the QUIC \cite{iyengar_quic_2018}.
+An advantage of implementing our approach on top of UDP is that it is easier to develop and modify since no kernel module needs to be created and that it is more secure since the code is isolated in a user space process.
+
+\subsection{Deployment}
+
+Our approach of detecting \gls{fq} relies on using two concurrent flows between the sender and the receiver.
+This means that during flow startup, data is transmitted in two flows.
+To be able to recombine the data from the both flows, the receiver has to be aware of the fact that the two flows belong together.
+Thus, the receiver has to be aware of the fact that the two flows belong together and also has to be able to recombine the data.
+This seems like a problem regarding deployment since solutions that require multiple parties to implement are generally hard to deploy on a wide scale in the internet.
+This can be seen for example when looking at the deployment of IPv6, which requires the participation of every router on the way and every end device.
+However, since multipath TCP \cite{handley_tcp_2012} and QUIC are becoming increasingly prevalent, we think that the actual obstacles to deployment of our solution are not too high and are constantly decreasing \noteMax{Reference needed}.
+
+\section{Results}
+
+For the evaluation we use the network emulation library \texttt{py-virtnet}\footnote{\url{https://pypi.org/project/py-virtnet/}}.
+
+\subsection{Accuracy of the \gls{fq} detection mechanism}
+
+To evaluate if our proposed mechanism correctly recognizes, we perform experiments using a wide range of network configurations: We vary bandwidth from 1 to 50\,Mbps, delay from 10 to 100\,ms and the buffer size from 1 to 50 packets, using a grid of these parameters.
+For each parameter, we take 5 values.
+Thus, we run experiments with 125 different configurations.
+For each run each configuration once with \gls{fq} and once without.
+Results show that for experiments without \gls{fq} (using drop tail), in 95\% of cases, the absence of \gls{fq} was correctly detected and in 5\% cases, our algorithm detected \gls{fq} even though there was no \gls{fq}.
+For experiments, during which \gls{fq} was actually deployed on the bottleneck link, this was correctly detected in 92\% of cases and wrongly in 8\% of the time.
+
+\subsection{Delay- vs.~loss-based \gls{cc}}
+
+\begin{figure}[h]
+\centering
+\subfloat[Throughput (receiving rate)]{%
+  \includegraphics[width=\columnwidth]{{"figures/plots/throughput_1_just_one_flow_cubic_tcp_port9010_fq_10_10_20_1600183264284.pcap"}.pdf}%
+  \label{fig:just_one_flow_tcp_throughput}
+}\\
+\subfloat[\gls{rtt}]{%
+  \includegraphics[width=\columnwidth]{{"figures/plots/rtt_1_just_one_flow_cubic_tcp_port9010_fq_10_10_20_1600183264284.pcap"}.pdf}%
+  \label{fig:just_one_flow_tcp_delay}
+}
+\caption{Throughput and delay of a flow controlled by the loss-based \gls{cca} Cubic on a link with a speed of 10\,Mbps, a delay of 10\,ms and a buffer size of 100 packets.}
+\label{fig:just_one_flow_tcp}
+\end{figure}
+
+\begin{figure}[h]
+\centering
+\subfloat[Throughput (receiving rate)]{%
+  \includegraphics[width=\columnwidth]{{"figures/plots/throughput_1_just_one_flow_vegas_udp_port9000_fq_10_10_20_1600183241479.pcap"}.pdf}%
+}\\
+\subfloat[\gls{rtt}]{%
+  \includegraphics[width=\columnwidth]{{"figures/plots/rtt_1_just_one_flow_vegas_udp_port9000_fq_10_10_20_1600183241479.pcap"}.pdf}%
+}
+\caption{Throughput and delay of a flow controlled by our delay-based \gls{cca} on a link with a speed of 10\,Mbps, a delay of 10\,ms and a buffer size of 100 packets.}
+\label{fig:just_one_flow_udp}
+\end{figure}
+
+First, we want to demonstrate that delay-based \gls{cc} actually has a benefit over loss-based one.
+\autoref{fig:just_one_flow_tcp} shows that the loss-based \gls{cc} achieves a throughput of almost 10\,Mbps, which means full utilization and an \gls{rtt} that is generally higher than 100\,ms.
+On the other side, \autoref{fig:just_one_flow_udp} shows an example flow using our \gls{cca} and a slow start algorithm similar to the one used by the classic TCP \gls{cca} \textit{New Reno} \cite{allman_tcp_1999}.
+While the delay-based flow also achieves very high utilization, delay rarely exceeds 20\,ms, meaning that it is more than 5 times lower than the one of the loss-based \gls{cca} of \autoref{fig:just_one_flow_tcp}.
+
+\subsection{Comparing flows with different Congestion Controls and AQM mechanisms}
+In this scenario, one flow with our mechanism shares a bottleneck link with a flow using a loss-based congestion control.
+
+\subsubsection{Drop tail buffer (no \gls{fq})}
+\autoref{fig:pfifo_tcp} and \autoref{fig:pfifo_udp} show the loss-based \gls{cca} Cubic compete against our delay-based one on a link with a drop tail queue and thus without \gls{fq}.
+In the scenario, the loss-based flow starts 5 seconds earlier and after 5 seconds, the delay-based flow joins.
+While the delay-based flow gains some share of the link during startup, it is pushed away by the loss-based flow later on and ``starves''.
+
+\begin{figure}[h]
+\centering
+\subfloat[Throughput (receiving rate)]{%
+  \includegraphics[width=\columnwidth]{{"figures/plots/throughput_1_competing_flow_pfifo_vegas_tcp_port9010_pfifo_10_50_20_1600182743408.pcap"}.pdf}%
+}\\
+\subfloat[\gls{rtt}]{%
+  \includegraphics[width=\columnwidth]{{"figures/plots/rtt_1_competing_flow_pfifo_vegas_tcp_port9010_pfifo_10_50_20_1600182743408.pcap"}.pdf}%
+}
+\caption{Throughput and delay of a flow controlled by the loss-based \gls{cca} Cubic on a link with a speed of 50\,Mbps, a delay of 10\,ms and a buffer size of 100 packets.
+The bottleneck is controlled by drop tail (no \gls{fq}).}
+\label{fig:pfifo_tcp}
+\end{figure}
+
+\begin{figure}[h]
+\centering
+\subfloat[Throughput (receiving rate)]{%
+  \includegraphics[width=\columnwidth]{{"figures/plots/throughput_1_competing_flow_pfifo_vegas_udp_port9000_pfifo_10_50_20_1600182743408.pcap"}.pdf}%
+}\\
+\subfloat[\gls{rtt}]{%
+  \includegraphics[width=\columnwidth]{{"figures/plots/rtt_1_competing_flow_pfifo_vegas_udp_port9000_pfifo_10_50_20_1600182743408.pcap"}.pdf}%
+}
+\caption{Throughput and delay of a flow controlled by our delay-based \gls{cca} on a link with a speed of 50\,Mbps, a delay of 10\,ms and a buffer size of 100 packets.
+The bottleneck is controlled by drop tail (no \gls{fq}).}
+\label{fig:pfifo_udp}
+\end{figure}
+
+\subsubsection{\gls{fq}}
+
+\begin{figure}[h]
+\centering
+\subfloat[Throughput (receiving rate)]{%
+  \includegraphics[width=\columnwidth]{{"figures/plots/throughput_1_competing_flow_fq_tcp_port9010_fq_10_50_20_1600182657352.pcap"}.pdf}%
+}\\
+\subfloat[\gls{rtt}]{%
+  \includegraphics[width=\columnwidth]{{"figures/plots/rtt_1_competing_flow_fq_tcp_port9010_fq_10_50_20_1600182657352.pcap"}.pdf}%
+}
+\caption{Throughput and delay of a flow controlled by the loss-based \gls{cca} Cubic on a link with a speed of 50\,Mbps, a delay of 10\,ms and a buffer size of 100 packets.
+The bottleneck is controlled by \gls{fq}.}
+\label{fig:fq_tcp}
+\end{figure}
+
+\begin{figure}[h]
+\centering
+\subfloat[Throughput (receiving rate)]{%
+  \includegraphics[width=\columnwidth]{{"figures/plots/throughput_1_competing_flow_fq_udp_port9000_fq_10_50_20_1600182657352.pcap"}.pdf}%
+}\\
+\subfloat[\gls{rtt}]{%
+  \includegraphics[width=\columnwidth]{{"figures/plots/rtt_1_competing_flow_fq_udp_port9000_fq_10_50_20_1600182657352.pcap"}.pdf}%
+}
+\caption{Throughput and delay of a flow controlled by our delay-based \gls{cca} on a link with a speed of 50\,Mbps, a delay of 10\,ms and a buffer size of 100 packets.
+The bottleneck is controlled by \gls{fq}.}
+\label{fig:fq_udp}
+\end{figure}
+
+For a link with fair queuing, the following results show that throughput is identical for our flow (\autoref{fig:fq_udp}) and the Cubic flow (\autoref{fig:fq_tcp}), just that the Cubic flow's delay is more than two times higher than the loss-based one's.
+Thus, our mechanism is effective to detect if the bottleneck uses fair queuing and use a delay-based congestion control if it is the case and a loss-based one otherwise.
+
+\section{Discussion}
+
+During the implementation of our algorithm we noticed peculiar behavior when not using \gls{fq}: During startup, we make flow 1 have half the sending rate of flow 2.
+We would expect the receiving rates to be the same: flow 1 achieving around half the rate of flow 2.
+However, we noticed that usually flow 1 would have a receiving rate of 0.
+After investigating the issue we noticed that the reason was pacing: PCC (and other \glspl{cca}) such as BBR send packets in regular intervals, for example every millisecond, because sending regularly keeps the queues shorter.
+This behavior is called pacing.
+However, when there are two flows where one flow's sending rate is a multiple of the other's, an interlocking effect occurs: flow 1 would always send each packet slightly after flow 2 sends a packet.
+Then, when the packets arrive at the bottleneck, flow 2's packet, which arrives a bit earlier, fills up the buffer and flow 1's packet is consequently dropped.
+As a solution, we do not send packets at regular intervals but always add a small random value: For example, if the sending interval were 1\,ms, we would send the packet randomly in the interval from 0.5 to 1.5\,ms.
+This solves the aforementioned problem.
+The insight we gained from this is that while pacing is generally good, it is actually harmful if several flows compete at a bottleneck without \gls{fq} and one flow's sending rate is a multiple of the other's.
+
+Finally, we think that our new flow startup method can help to increase the deployment of delay-based \gls{cc}.
+Especially for applications that are very delay-sensitive, such as video conferencing and cloud gaming, the adoption of delay-based \gls{cc} can result in a significant increase of \gls{qoe}.
+We demonstrated that our method works using a prototype implementation using PCC but we are also confident that a similar flow startup method can also be integrated with other \gls{cca} such as BBR \cite{cardwell_bbr:_2016}.
+
+\bibliographystyle{splncs04}
 \bibliography{fq_detection}

### 2020-09-24 src/plot_rtt_and_bandwidth.py

@@ -16,6 +16,8 @@ pcap_file = sys.argv[1]
 if pcap_file.startswith("pcaps/"):
 	pcap_file = pcap_file[len("pcaps/"):]
 
+is_secondary_flow = "port9010" in pcap_file
+
 receiver_pcap = 'receiver_'+('_'.join(pcap_file.split('_')[1:]))
 
 # new_rtt_command = f"tshark -r pcaps/{pcap_file} -Tfields -e frame.time_relative -e tcp.analysis.ack_rtt"
@@ -88,7 +90,7 @@ bytes_results = [(float(item[0][0]), float(item[0][1]), float(item[1])) for item
 # divided = [(first[0], first[1], first[2]/second[2]) for first, second in zip(bytes_results, packets_results)]
 # print("divided", divided)
 
-with_correct_time = [((item[0] + item[1])/2, item[2]/1000000*8) for item in bytes_results]
+with_correct_time = [((item[0] + item[1])/2 - (5 if is_secondary_flow else 0), item[2]/1000000*8) for item in bytes_results]
 os.makedirs("tex/plots", exist_ok=True)
 
 plt.figure(figsize=(5,2))
@@ -98,8 +100,9 @@ plt.plot(*zip(*with_correct_time))
 plt.tight_layout()
 
 plt.ylim(bottom=0)
-# plt.savefig(f"plots/throughput_{interval}_{('_'.join(pcap_file.split('_')[1:]))}.pdf", bbox_inches = 'tight', pad_inches = 0)
-plt.savefig(f"plots/throughput_{interval}_{('_'.join(pcap_file.split('_')[1:]))}.png", bbox_inches = 'tight', pad_inches = 0, dpi=200)
+plt.grid()
+plt.savefig(f"plots/throughput_{interval}_{('_'.join(pcap_file.split('_')[1:]))}.pdf", bbox_inches = 'tight', pad_inches = 0)
+# plt.savefig(f"plots/throughput_{interval}_{('_'.join(pcap_file.split('_')[1:]))}.png", bbox_inches = 'tight', pad_inches = 0, dpi=200)
 
 plt.close()
 
@@ -109,6 +112,8 @@ plt.figure(figsize=(5,2))
 plt.xlabel("Time [s]")
 plt.ylabel(f"RTT [ms]")
 # plt.plot(*zip(*rtt_results))
+
[…]

### 2020-09-24 src/test.py

@@ -172,8 +172,8 @@ def run(vnet, prefix=""):
 			tcpdump_receiver_popens = []
 
 			if not opt.only_iperf:
-				tcpdump_sender_popens.append(hosts[0].Popen(f"/usr/sbin/tcpdump -s {opt.bytes_to_capture} -i eth0 -w pcaps/sender_{prefix}_{protocol_for_main_flow}_port{opt.cport}_{opt.qdisc}_{opt.delay}_{opt.rate}_{opt.time}_{start_time}.pcap dst port {opt.cport} or src port {opt.cport}".split(" "), stdout=subprocess.PIPE, stderr=subprocess.PIPE))
-				tcpdump_receiver_popens.append(hosts[1].Popen(f"/usr/sbin/tcpdump -s {opt.bytes_to_capture} -i eth0 -w pcaps/receiver_{prefix}_{protocol_for_main_flow}_port{opt.cport}_{opt.qdisc}_{opt.delay}_{opt.rate}_{opt.time}_{start_time}.pcap dst port {opt.cport} or src port {opt.cport}".split(" "), stdout=subprocess.PIPE, stderr=subprocess.PIPE))
+				tcpdump_sender_popens.append(hosts[0].Popen(f"/usr/sbin/tcpdump -s {opt.bytes_to_capture} -i eth0 -w pcaps/sender_{prefix}_{protocol_for_main_flow}_port{opt.cport}_{opt.qdisc}_{opt.delay}_{opt.rate}_{opt.buffer_size}_{opt.time}_{start_time}.pcap dst port {opt.cport} or src port {opt.cport}".split(" "), stdout=subprocess.PIPE, stderr=subprocess.PIPE))
+				tcpdump_receiver_popens.append(hosts[1].Popen(f"/usr/sbin/tcpdump -s {opt.bytes_to_capture} -i eth0 -w pcaps/receiver_{prefix}_{protocol_for_main_flow}_port{opt.cport}_{opt.qdisc}_{opt.delay}_{opt.rate}_{opt.buffer_size}_{opt.time}_{start_time}.pcap dst port {opt.cport} or src port {opt.cport}".split(" "), stdout=subprocess.PIPE, stderr=subprocess.PIPE))
 
 			if opt.competing_flow:
[…]

### 2020-09-25 paper/paper.tex
Finish paper draft

@@
 This unfairness can be mitigated by using \gls{fq} at the bottleneck link, isolating each flow from all other flows and assigning each flow an equal share of bandwidth \cite{dumazet_pkt_sched:_2013}.
-Then, the loss-based flows cannot obtain the bandwidth of the loss-based flows anymore.
+Then, the loss-based flows cannot obtain the bandwidth that would actually belong to the loss-based flows anymore.
 In this case, delay-based \glspl{cca} fulfill their goal of high bandwidth and low delay.
@@
 While \gls{fq} solves many problems regarding \gls{cc}, it is still not ubiquitously deployed in every bottleneck link.
-If, however, flows could know if there's \gls{fq} or not, they could dynamically adapt their congestion control.
+If, however, flows could know if there's \gls{fq} or not, they could dynamically adapt their \gls{cc}.
 If there is \gls{fq} they would use a delay-based \gls{cca} and otherwise a loss-based one.
-Our contribution is the design and evaluation of such a mechanism.
-This mechanism determines the presence of \gls{fq} during the startup phase of a flow's \gls{cca} and, at the end of the startup, sets the \gls{cca} accordingly.
+Our contribution is the design and evaluation of such a mechanism that determines the presence of \gls{fq} during the startup phase of a flow's \gls{cca} and, at the end of the startup, sets the \gls{cca} accordingly.
 
-We show that this solution reliably determines the presence of fair queuing and that by using this approach, queuing delay can be considerably lowered while keeping throughput high.
+We show that this solution reliably determines the presence of fair queuing and that by using this approach, queuing delay can be considerably lowered while keeping throughput high if \gls{fq} is detected.
+In case no \gls{fq} is detected, our algorithm detects this as well and a loss-based \gls{cc} is used, which does not starve when facing competition from the delay-based flow.
 While our mechanism is beneficial for network flows in general, we argue that it is most useful for long-running delay-sensitive flows, such as cloud gaming, video conferencing etc.
@@
 
-\section{Related work}
+%\section{Related work}
 
@@
 \item If \gls{fq} is used, the $\textit{loss ratio}$ is 2, otherwise it is 1. We choose 1.5 as the cutoff value. 
+\item Since the presence or absence of \gls{fq} is now known the appropriate \gls{cca} can be launched. 
 \end{enumerate}
@@
 On this bottleneck link, the queue is either managed by \gls{fq} or by a drop tail buffer (no \gls{fq}).
-\autoref{fig:throughput} shows how the sending rate before the bottleneck link is doubled after each \gls{rtt}.
-Furthermore, flow 2 has double the sending rate of flow 1 \noteMax{Figures take up too much space.
+\autoref{fig:throughput} shows how the sending rate (observed before the bottleneck link) is doubled after each \gls{rtt}.
+Furthermore, it shows that flow 2 has double the sending rate of flow 1.
+%\noteMax{Figures take up too much space.
 Is it better to put them next to each other?
@@
 Figure~\ref{fig:goodput_fq}, on the other side, shows the receiving rate of the two flows after the bottleneck link in case \gls{fq} is deployed.
-In the third \gls{rtt}, flow 1 and flow 2 achieve the same receiving rate because of the \gls{fq}.
+In the third \gls{rtt}, flow 1 and flow 2 achieve the same receiving rate because of the \gls{fq}, which makes sure that each flow only gets its fair share.
 
@@
 We use the algorithm \textit{PCC} proposed by \cite{dong_pcc_2015}.
-In simple terms, this \gls{cca} aims to maximize a utility function, which strives for high throughput and low packet loss.
+In simplified terms, this \gls{cca} aims to maximize a utility function, which strives for high throughput and low packet loss.
 However, it is reasonably aggressive and does not starve when competing with other loss-based \glspl{cca} such as Cubic \cite{ha_cubic_2008}.
@@
 
-We base our implementation on the code of \textit{PCC Vivace} \cite{dong_pcc_2018}, which in turn is an improved version of the \cite{dong_pcc_2015}.
+We base our implementation on the code of \textit{PCC Vivace} \cite{dong_pcc_2018}, which in turn is an improved version of PCC.
 The PCC code is based on the \textit{UDT} library \cite{gu_udt_2007}, which is a library that provides a reliable transport protocol similar to TCP on top of UDP.
 This approach is similar to the QUIC \cite{iyengar_quic_2018}.
-An advantage of implementing our approach on top of UDP is that it is easier to develop and modify since no kernel module needs to be created and that it is more secure since the code is isolated in a user space process.
+An advantage of implementing our approach on top of UDP as opposed to TCP is that it is easier to develop and modify since no kernel module needs to be created and that it is more secure since the code is isolated in a user space process.
 
@@
 This means that during flow startup, data is transmitted in two flows.
-To be able to recombine the data from the both flows, the receiver has to be aware of the fact that the two flows belong together.
-Thus, the receiver has to be aware of the fact that the two flows belong together and also has to be able to recombine the data.
+To be able to recombine the data from the both flows, the receiver has to be aware of the fact that the two flows belong together and also has to be able to recombine the data.
 This seems like a problem regarding deployment since solutions that require multiple parties to implement are generally hard to deploy on a wide scale in the internet.
 This can be seen for example when looking at the deployment of IPv6, which requires the participation of every router on the way and every end device.
-However, since multipath TCP \cite{handley_tcp_2012} and QUIC are becoming increasingly prevalent, we think that the actual obstacles to deployment of our solution are not too high and are constantly decreasing \noteMax{Reference needed}.
+However, multipath TCP \cite{handley_tcp_2012} and QUIC are becoming increasingly prevalent and Google's QUIC implementation has multipath support \cite{google_quiche_2020}.
+We thus think that the actual obstacles to deployment of our solution are constantly decreasing since QUIC (mostly Google traffic) already accounts for around $\frac{1}{10}$ of all Internet traffic \cite{ruth_first_2018} and is growing and, for example, Apple's macOS and iOS support multipath TCP \cite{apple_about_2018}.
 
@@
 
-To evaluate if our proposed mechanism correctly recognizes, we perform experiments using a wide range of network configurations: We vary bandwidth from 1 to 50\,Mbps, delay from 10 to 100\,ms and the buffer size from 1 to 50 packets, using a grid of these parameters.
+To evaluate if our proposed mechanism correctly recognizes, we perform experiments using a wide range of network configurations: We vary bandwidth from 1 to 50\,Mbps, delay from 10 to 100\,ms and the buffer size from 1 to 100 packets, using a grid of these parameters.
 For each parameter, we take 5 values.
 Thus, we run experiments with 125 different configurations.
-For each run each configuration once with \gls{fq} and once without.
-Results show that for experiments without \gls{fq} (using drop tail), in 95\% of cases, the absence of \gls{fq} was correctly detected and in 5\% cases, our algorithm detected \gls{fq} even though there was no \gls{fq}.
-For experiments, during which \gls{fq} was actually deployed on the bottleneck link, this was correctly detected in 92\% of cases and wrongly in 8\% of the time.
+We run each configuration once with \gls{fq} and once without.
+Results show that for experiments without \gls{fq} (using drop tail), in over 99\% of cases, the absence of \gls{fq} was correctly detected and in less than 1\% of cases, our algorithm detected \gls{fq} even though there was no \gls{fq}.
+For experiments, during which \gls{fq} was actually deployed on the bottleneck link, this was correctly detected in 97\% of cases and wrongly in 3\% of the time.
+The overall accuracy is thus 98\%, which we consider sufficient.
 
@@
 First, we want to demonstrate that delay-based \gls{cc} actually has a benefit over loss-based one.
-\autoref{fig:just_one_flow_tcp} shows that the loss-based \gls{cc} achieves a throughput of almost 10\,Mbps, which means full utilization and an \gls{rtt} that is generally higher than 100\,ms.
-On the other side, \autoref{fig:just_one_flow_udp} shows an example flow using our \gls{cca} and a slow start algorithm similar to the one used by the classic TCP \gls{cca} \textit{New Reno} \cite{allman_tcp_1999}.
+\autoref{fig:just_one_flow_tcp} shows that the loss-based \gls{cc} achieves a throughput of almost 10\,Mbps on a 10\,Mbps link, which means full utilization and an \gls{rtt} that is generally higher than 100\,ms.
+On the other side, \autoref{fig:just_one_flow_udp} shows an example flow using our simple delay-based \gls{cca} and a slow start algorithm similar to the one used by the classic TCP \gls{cca} \textit{New Reno} \cite{allman_tcp_1999}.
 While the delay-based flow also achieves very high utilization, delay rarely exceeds 20\,ms, meaning that it is more than 5 times lower than the one of the loss-based \gls{cca} of \autoref{fig:just_one_flow_tcp}.
@@
 Then, when the packets arrive at the bottleneck, flow 2's packet, which arrives a bit earlier, fills up the buffer and flow 1's packet is consequently dropped.
-As a solution, we do not send packets at regular intervals but always add a small random value: For example, if the sending interval were 1\,ms, we would send the packet randomly in the interval from 0.5 to 1.5\,ms.
+As a solution, we do not send packets at regular intervals during startup but always add a small random value: For example, if the sending interval were 1\,ms, we would send the packet randomly in the interval from 0.5 to 1.5\,ms.
 This solves the aforementioned problem.
@@
 Especially for applications that are very delay-sensitive, such as video conferencing and cloud gaming, the adoption of delay-based \gls{cc} can result in a significant increase of \gls{qoe}.
-We demonstrated that our method works using a prototype implementation using PCC but we are also confident that a similar flow startup method can also be integrated with other \gls{cca} such as BBR \cite{cardwell_bbr:_2016}.
+We demonstrated that our method works using a prototype implementation using PCC but we are also confident that a similar flow startup method can also be integrated with other \glspl{cca} such as BBR \cite{cardwell_bbr:_2016}.
 

### 2020-09-25 src/plot_rtt_and_bandwidth.py

@@ -17,6 +17,7 @@ if pcap_file.startswith("pcaps/"):
 	pcap_file = pcap_file[len("pcaps/"):]
 
 is_secondary_flow = "port9010" in pcap_file
+just_one_flow = "just_one_flow" in pcap_file
 
 receiver_pcap = 'receiver_'+('_'.join(pcap_file.split('_')[1:]))
 
@@ -90,13 +91,13 @@ bytes_results = [(float(item[0][0]), float(item[0][1]), float(item[1])) for item
 # divided = [(first[0], first[1], first[2]/second[2]) for first, second in zip(bytes_results, packets_results)]
 # print("divided", divided)
 
-with_correct_time = [((item[0] + item[1])/2 - (5 if is_secondary_flow else 0), item[2]/1000000*8) for item in bytes_results]
+with_correct_time = [((item[0] + item[1])/2 - (5 if is_secondary_flow and not just_one_flow else 0), item[2]/1000000*8) for item in bytes_results]
 os.makedirs("tex/plots", exist_ok=True)
 
 plt.figure(figsize=(5,2))
 plt.xlabel("Time [s]")
 plt.ylabel(f"Throughput [Mbit/s]")
-plt.plot(*zip(*with_correct_time))
+plt.plot(*zip(*[(a, b) for a, b in with_correct_time if a <= 20]))
 plt.tight_layout()
 
 plt.ylim(bottom=0)
@@ -113,8 +114,8 @@ plt.xlabel("Time [s]")
 plt.ylabel(f"RTT [ms]")
 # plt.plot(*zip(*rtt_results))
 
-ack_timestamps = [ts - (5 if is_secondary_flow else 0) for ts in ack_timestamps]
-plt.plot(ack_timestamps, rtts)
+ack_timestamps = [ts - (5 if is_secondary_flow and not just_one_flow else 0) for ts in ack_timestamps]
+plt.plot(*zip(*[(a, b) for a, b in zip(ack_timestamps, rtts) if a <= 20]))
 # print("rtt_results", *zip(*rtt_results))
 # plt.scatter(retransmissions_results, [0]*len(retransmissions_results), marker=".", linestyle="None", color="r", s=2, edgecolors="none")
 

### 2020-09-28 paper/paper.tex
Fix bug and modify paper a little

@@
 We assume a sender being connected to a receiver, with a bottleneck link between them.
-On this bottleneck link, the queue is either managed by \gls{fq} or by a drop tail buffer (no \gls{fq}).
+On this bottleneck link, the queue is either managed by \gls{fq} or by a shared queue (no \gls{fq}).
 \autoref{fig:throughput} shows how the sending rate (observed before the bottleneck link) is doubled after each \gls{rtt}.
@@
 
-Figure~\ref{fig:goodput_no_fq} shows the receiving rate at the receiver after the bottleneck link, if this bottleneck link has a drop tail buffer without \gls{fq}.
+Figure~\ref{fig:goodput_no_fq} shows the receiving rate at the receiver after the bottleneck link, if this bottleneck link has a a shared queue without \gls{fq}.
 In the third \gls{rtt} the sending rate exceeds the link speed of the bottleneck and thus the receiving rate is smaller than the sending rate.
@@
 For the evaluation we use the network emulation library \texttt{py-virtnet}\footnote{\url{https://pypi.org/project/py-virtnet/}}.
+For the experiments with \gls{fq} we use the \texttt{fq} module for Linux, for the experiments with a shared queue we use \texttt{pfifo}.
 
@@
 
-To evaluate if our proposed mechanism correctly recognizes, we perform experiments using a wide range of network configurations: We vary bandwidth from 1 to 50\,Mbps, delay from 10 to 100\,ms and the buffer size from 1 to 100 packets, using a grid of these parameters.
+To evaluate if our proposed mechanism correctly recognizes, we perform experiments using a wide range of network configurations: We vary bandwidth from 5 to 50\,Mbps, delay from 10 to 100\,ms and the buffer size from 1 to 100 packets, using a grid of these parameters.
 For each parameter, we take 5 values.
@@
 We run each configuration once with \gls{fq} and once without.
-Results show that for experiments without \gls{fq} (using drop tail), in over 99\% of cases, the absence of \gls{fq} was correctly detected and in less than 1\% of cases, our algorithm detected \gls{fq} even though there was no \gls{fq}.
+Results show that for experiments without \gls{fq} (using a shared queue), in over 99\% of cases, the absence of \gls{fq} was correctly detected and in less than 1\% of cases, our algorithm detected \gls{fq} even though there was no \gls{fq}.
 For experiments, during which \gls{fq} was actually deployed on the bottleneck link, this was correctly detected in 97\% of cases and wrongly in 3\% of the time.
 The overall accuracy is thus 98\%, which we consider sufficient.
+
+To verify the behavior of our algorithm under the presence of cross traffic, we repeated the above experiments but with one bulk traffic flow using Cubic running at the same time, which we start 5 seconds before so that it can gain sufficient bandwidth.
+Also under these circumstances we achieve 98\% detection accuracy using our algorithm.
 
@@
 
-\subsubsection{Drop tail buffer (no \gls{fq})}
-\autoref{fig:pfifo_tcp} and \autoref{fig:pfifo_udp} show the loss-based \gls{cca} Cubic compete against our delay-based one on a link with a drop tail queue and thus without \gls{fq}.
+\subsubsection{No \gls{fq}}
+\autoref{fig:pfifo_tcp} and \autoref{fig:pfifo_udp} show the loss-based \gls{cca} Cubic compete against our delay-based one on a link without \gls{fq}.
 In the scenario, the loss-based flow starts 5 seconds earlier and after 5 seconds, the delay-based flow joins.
@@
 \caption{Throughput and delay of a flow controlled by the loss-based \gls{cca} Cubic on a link with a speed of 50\,Mbps, a delay of 10\,ms and a buffer size of 100 packets.
-The bottleneck is controlled by drop tail (no \gls{fq}).}
+The bottleneck is controlled by a shared queue (no \gls{fq}).}
 \label{fig:pfifo_tcp}
@@
 \caption{Throughput and delay of a flow controlled by our delay-based \gls{cca} on a link with a speed of 50\,Mbps, a delay of 10\,ms and a buffer size of 100 packets.
-The bottleneck is controlled by drop tail (no \gls{fq}).}
+The bottleneck is controlled by a shared queue (no \gls{fq}).}
 \label{fig:pfifo_udp}

### 2020-09-28 src/test.py

@@ -30,22 +30,11 @@ except KeyError:
 	pass
 
 import subprocess
-# import matplotlib
-# matplotlib.use("Agg")
-# import matplotlib.pyplot as plt
 import virtnet
 import statistics
 import argparse
 
-# import numpy as np
-# from scipy.stats import norm
 
-
-# DELAY = 10000
-# SIGMA = 2000
-# SAMPLES = 1000
-# X = 10
-# NUMPING = 1000
 
 parser = argparse.ArgumentParser()
 parser.add_argument('--bytes_to_capture', type=int, default=100)
@@ -205,6 +194,7 @@ def run(vnet, prefix=""):
 				print("client out", out.decode("utf-8"))
 			if err:
 				print("client err", err.decode("utf-8"))
+			client_out = out
 			if opt.competing_flow:
 				client_popen_iperf.terminate()
 				out, err = client_popen_iperf.stdout.read(), client_popen_iperf.stderr.read()
@@ -212,8 +202,6 @@ def run(vnet, prefix=""):
 					print("client iperf out", out.decode("utf-8"))
 				if err:
 					print("client iperf err", err.decode("utf-8"))
-
-			client_out = out
 		else:
 			client_out = b""
 
@@ -260,7 +248,7 @@ if opt.run_scenario == "":
 elif opt.run_scenario == "accuracy":
 	import sklearn.metrics
 	results_dict = {}
-	for bw_index, bw in enumerate(np.linspace(1,50,opt.how_many_values_per_parameter)):
+	for bw_index, bw in enumerate(np.linspace(5,50,opt.how_many_values_per_parameter)):
 		for delay_index, delay in enumerate(np.linspace(10,100,opt.how_many_values_per_parameter)):
 			for buffer_index, buffer in enumerate(np.linspace(1,100,opt.how_many_values_per_parameter)):
 				for fq_index, fq in enumerate([False, True]):
@@ -272,8 +260,10 @@ elif opt.run_scenario == "accuracy":
 
 					with virtnet.Manager() as context:
 						client_output = run(context)
+					assert client_output != ""
 					contained_vegas = "Starting Vegas" in client_output
 					contained_pcc = "Starting PCC Classic" in client_output
+					# print("client_output", client_output)
 
 					results_dict[(bw, delay, buffer, fq)] = (contained_vegas, contained_pcc)
 

### 2020-09-29 paper/paper.tex
Implement capability of getting overall throughput and delay

@@
 
-Emerging applications such as cloud gaming \cite{jarschel_evaluation_2011} and virtual reality \cite{elbamby_toward_2018} applications require high throughput and low delay.
+Emerging applications such as cloud gaming \cite{jarschel_evaluation_2011} and remote virtual reality \cite{elbamby_toward_2018} applications require high throughput and low delay.
 Furthermore, for 5G, ultra low latency communications have been made a priority \cite{li_5g_2018}.
@@
 Such approaches have been proposed in the last century \cite{brakmo_tcp_1995} and recently have seen a surge in popularity \cite{arun_copa_2018,hock_tcp_2017,mittal_timely_2015,cardwell_bbr:_2016}.
-While loss-based \glspl{cca} fulfill their goal of low delay and high throughput, they cannot handle competing flows with different \glspl{cca} well.
+While delay-based \glspl{cca} fulfill their goal of low delay and high throughput, they cannot handle competing flows with different \glspl{cca} well \cite{turkovic_fifty_2019, turkovic_interactions_2019}.
 This is especially prevalent when they have to compete against loss-based \glspl{cca}.
@@
 \begin{align}
-\textit{loss ratio} &= \frac{\left(\frac{\textit{sending rate of flow 1}}{\textit{receiving rate of flow 1}}\right)}{\left(\frac{\textit{sending rate of flow 2}}{\textit{receiving rate of flow 2}}\right)} \\
-&= \frac{\textit{sending rate of flow 1}\times \textit{receiving rate of flow 2}}{\textit{receiving rate of flow 1} \times \textit{sending rate of flow 2}}
+\textit{loss ratio} &= \frac{\left(\frac{\textit{receiving rate of flow 1}}{\textit{sending rate of flow 1}}\right)}{\left(\frac{\textit{receiving rate of flow 2}}{\textit{sending rate of flow 2}}\right)} \\
+&= \frac{\textit{receiving rate of flow 1}\times \textit{sending rate of flow 2}}{\textit{sending rate of flow 1} \times \textit{receiving rate of flow 2}}
 \end{align}
 Basically, this metric indicates if flow 2, which has a higher sending rate, also achieves a higher receiving rate (goodput) than flow 1. If this is the case, there's no \gls{fq}, otherwise there is.
+
+It is necessary to wait for packet loss in both connections because only then it is certain that the link is saturated and our algorithm assumes that the link is saturated. 
 \item If \gls{fq} is used, the $\textit{loss ratio}$ is 2, otherwise it is 1. We choose 1.5 as the cutoff value. 
@@
 To be able to recombine the data from the both flows, the receiver has to be aware of the fact that the two flows belong together and also has to be able to recombine the data.
-This seems like a problem regarding deployment since solutions that require multiple parties to implement are generally hard to deploy on a wide scale in the internet.
+This seems like a problem regarding deployment since solutions that require multiple parties to implement are generally hard to deploy on a wide scale in the Internet.
 This can be seen for example when looking at the deployment of IPv6, which requires the participation of every router on the way and every end device.
@@
 
-To verify the behavior of our algorithm under the presence of cross traffic, we repeated the above experiments but with one bulk traffic flow using Cubic running at the same time, which we start 5 seconds before so that it can gain sufficient bandwidth.
+To verify the behavior of our algorithm under the presence of cross traffic, we repeated the above experiments but with one bulk traffic flow using Cubic running at the same time, which we start 5 seconds before so that it has sufficient time to gain full bandwidth.
 Also under these circumstances we achieve 98\% detection accuracy using our algorithm.
@@
 \subfloat[Throughput (receiving rate)]{%
-  \includegraphics[width=\columnwidth]{{"figures/plots/throughput_1_just_one_flow_cubic_tcp_port9010_fq_10_10_20_1600183264284.pcap"}.pdf}%
+  \includegraphics[width=0.48\columnwidth]{{"figures/plots/throughput_1_just_one_flow_cubic_tcp_port9010_fq_10_10_20_1600183264284.pcap"}.pdf}%
   \label{fig:just_one_flow_tcp_throughput}
-}\\
+}\hfill
 \subfloat[\gls{rtt}]{%
-  \includegraphics[width=\columnwidth]{{"figures/plots/rtt_1_just_one_flow_cubic_tcp_port9010_fq_10_10_20_1600183264284.pcap"}.pdf}%
+  \includegraphics[width=0.48\columnwidth]{{"figures/plots/rtt_1_just_one_flow_cubic_tcp_port9010_fq_10_10_20_1600183264284.pcap"}.pdf}%
   \label{fig:just_one_flow_tcp_delay}
@@
 \subfloat[Throughput (receiving rate)]{%
-  \includegraphics[width=\columnwidth]{{"figures/plots/throughput_1_just_one_flow_vegas_udp_port9000_fq_10_10_20_1600183241479.pcap"}.pdf}%
-}\\
+  \includegraphics[width=0.48\columnwidth]{{"figures/plots/throughput_1_just_one_flow_vegas_udp_port9000_fq_10_10_20_1600183241479.pcap"}.pdf}%
+}\hfill
 \subfloat[\gls{rtt}]{%
-  \includegraphics[width=\columnwidth]{{"figures/plots/rtt_1_just_one_flow_vegas_udp_port9000_fq_10_10_20_1600183241479.pcap"}.pdf}%
+  \includegraphics[width=0.48\columnwidth]{{"figures/plots/rtt_1_just_one_flow_vegas_udp_port9000_fq_10_10_20_1600183241479.pcap"}.pdf}%
 }
@@
 
-First, we want to demonstrate that delay-based \gls{cc} actually has a benefit over loss-based one.
+First, we want to empirically demonstrate that delay-based \gls{cc} actually has a benefit over loss-based one.
 \autoref{fig:just_one_flow_tcp} shows that the loss-based \gls{cc} achieves a throughput of almost 10\,Mbps on a 10\,Mbps link, which means full utilization and an \gls{rtt} that is generally higher than 100\,ms.
@@
 While the delay-based flow gains some share of the link during startup, it is pushed away by the loss-based flow later on and ``starves''.
+We start the loss-based flow before to make sure that it has sufficient time to fill up the link.
 
@@
 \subfloat[Throughput (receiving rate)]{%
-  \includegraphics[width=\columnwidth]{{"figures/plots/throughput_1_competing_flow_pfifo_vegas_tcp_port9010_pfifo_10_50_20_1600182743408.pcap"}.pdf}%
-}\\
+  \includegraphics[width=0.48\columnwidth]{{"figures/plots/throughput_1_competing_flow_pfifo_vegas_tcp_port9010_pfifo_10_50_20_1600182743408.pcap"}.pdf}%
+}\hfill
 \subfloat[\gls{rtt}]{%
-  \includegraphics[width=\columnwidth]{{"figures/plots/rtt_1_competing_flow_pfifo_vegas_tcp_port9010_pfifo_10_50_20_1600182743408.pcap"}.pdf}%
+  \includegraphics[width=0.48\columnwidth]{{"figures/plots/rtt_1_competing_flow_pfifo_vegas_tcp_port9010_pfifo_10_50_20_1600182743408.pcap"}.pdf}%
 }
 \caption{Throughput and delay of a flow controlled by the loss-based \gls{cca} Cubic on a link with a speed of 50\,Mbps, a delay of 10\,ms and a buffer size of 100 packets.
-The bottleneck is controlled by a shared queue (no \gls{fq}).}
+The bottleneck is controlled by a shared queue (no \gls{fq}) and is shared with the flow of \autoref{fig:pfifo_udp}.}
 \label{fig:pfifo_tcp}
@@
 \subfloat[Throughput (receiving rate)]{%
-  \includegraphics[width=\columnwidth]{{"figures/plots/throughput_1_competing_flow_pfifo_vegas_udp_port9000_pfifo_10_50_20_1600182743408.pcap"}.pdf}%
-}\\
+  \includegraphics[width=0.48\columnwidth]{{"figures/plots/throughput_1_competing_flow_pfifo_vegas_udp_port9000_pfifo_10_50_20_1600182743408.pcap"}.pdf}%
+}\hfill
 \subfloat[\gls{rtt}]{%
-  \includegraphics[width=\columnwidth]{{"figures/plots/rtt_1_competing_flow_pfifo_vegas_udp_port9000_pfifo_10_50_20_1600182743408.pcap"}.pdf}%
+  \includegraphics[width=0.48\columnwidth]{{"figures/plots/rtt_1_competing_flow_pfifo_vegas_udp_port9000_pfifo_10_50_20_1600182743408.pcap"}.pdf}%
 }
 \caption{Throughput and delay of a flow controlled by our delay-based \gls{cca} on a link with a speed of 50\,Mbps, a delay of 10\,ms and a buffer size of 100 packets.
-The bottleneck is controlled by a shared queue (no \gls{fq}).}
+The bottleneck is controlled by a shared queue (no \gls{fq}) and is shared with the flow of \autoref{fig:fq_tcp} and is shared with the flow of \autoref{fig:pfifo_tcp}.}
 \label{fig:pfifo_udp}
@@
 \subfloat[Throughput (receiving rate)]{%
-  \includegraphics[width=\columnwidth]{{"figures/plots/throughput_1_competing_flow_fq_tcp_port9010_fq_10_50_20_1600182657352.pcap"}.pdf}%
-}\\
+  \includegraphics[width=0.48\columnwidth]{{"figures/plots/throughput_1_competing_flow_fq_tcp_port9010_fq_10_50_20_1600182657352.pcap"}.pdf}%
+}\hfill
 \subfloat[\gls{rtt}]{%
-  \includegraphics[width=\columnwidth]{{"figures/plots/rtt_1_competing_flow_fq_tcp_port9010_fq_10_50_20_1600182657352.pcap"}.pdf}%
+  \includegraphics[width=0.48\columnwidth]{{"figures/plots/rtt_1_competing_flow_fq_tcp_port9010_fq_10_50_20_1600182657352.pcap"}.pdf}%
 }
 \caption{Throughput and delay of a flow controlled by the loss-based \gls{cca} Cubic on a link with a speed of 50\,Mbps, a delay of 10\,ms and a buffer size of 100 packets.
-The bottleneck is controlled by \gls{fq}.}
+The bottleneck is controlled by \gls{fq} and is shared with the flow of \autoref{fig:fq_udp}.}
 \label{fig:fq_tcp}
@@
 \subfloat[Throughput (receiving rate)]{%
-  \includegraphics[width=\columnwidth]{{"figures/plots/throughput_1_competing_flow_fq_udp_port9000_fq_10_50_20_1600182657352.pcap"}.pdf}%
-}\\
+  \includegraphics[width=0.48\columnwidth]{{"figures/plots/throughput_1_competing_flow_fq_udp_port9000_fq_10_50_20_1600182657352.pcap"}.pdf}%
+}\hfill
 \subfloat[\gls{rtt}]{%
-  \includegraphics[width=\columnwidth]{{"figures/plots/rtt_1_competing_flow_fq_udp_port9000_fq_10_50_20_1600182657352.pcap"}.pdf}%
+  \includegraphics[width=0.48\columnwidth]{{"figures/plots/rtt_1_competing_flow_fq_udp_port9000_fq_10_50_20_1600182657352.pcap"}.pdf}%
 }
 \caption{Throughput and delay of a flow controlled by our delay-based \gls{cca} on a link with a speed of 50\,Mbps, a delay of 10\,ms and a buffer size of 100 packets.
-The bottleneck is controlled by \gls{fq}.}
+The bottleneck is controlled by \gls{fq} and is shared with the flow of \autoref{fig:fq_tcp}.}
 \label{fig:fq_udp}
@@
 
-For a link with fair queuing, the following results show that throughput is identical for our flow (\autoref{fig:fq_udp}) and the Cubic flow (\autoref{fig:fq_tcp}), just that the Cubic flow's delay is more than two times higher than the loss-based one's.
+For a link with fair queuing, the following results show that throughput is identical for our flow (\autoref{fig:fq_udp}) and the Cubic flow (\autoref{fig:fq_tcp}), just that the Cubic flow's delay is more than two times higher than the delay-based one's.
 Thus, our mechanism is effective to detect if the bottleneck uses fair queuing and use a delay-based congestion control if it is the case and a loss-based one otherwise.
@@
 This behavior is called pacing.
-However, when there are two flows where one flow's sending rate is a multiple of the other's, an interlocking effect occurs: flow 1 would always send each packet slightly after flow 2 sends a packet.
+However, when there are two flows where one flow's sending rate is a multiple of the other's, an interleaving effect occurs: flow 1 would always send each packet slightly after flow 2 sends a packet.
 Then, when the packets arrive at the bottleneck, flow 2's packet, which arrives a bit earlier, fills up the buffer and flow 1's packet is consequently dropped.
@@
 The insight we gained from this is that while pacing is generally good, it is actually harmful if several flows compete at a bottleneck without \gls{fq} and one flow's sending rate is a multiple of the other's.
+
+One problem of our proposed solution is when the bottleneck link changes while a flow is running.
+For example, the bottleneck link in the beginning could be the home router, while later is it becoming the another router deeper in the core network.
+The problem occurs if the home router supports \gls{fq} and our algorithm detects this while the router in the core network, which become the bottleneck later on, doesn't support \gls{fq}.
+Then, our algorithm is going to use a delay-based \gls{cc} because that's the case in the beginning of the flow.
+However, later when the bottleneck changes, our algorithm is still using delay-based \gls{cc} even though the current bottleneck doesn't support \gls{fq} anymore.
+As a result, it might be possible that our algorithm uses the wrong \gls{cc}.
+One way to prevent this problem from occurring is to use periodic measurements.
+For example, our algorithm to determine \gls{fq} could be used every 5 or 10 seconds to measure again if the current bottleneck supports \gls{fq}.
+
+An alternative to our algorithm could be that the routers tell the endpoints if they support \gls{fq} or not.
+Such a solution could, for example, use an IP or a TCP extension, which every router uses to indicate if they're capability of \gls{fq}.
+Then, if every device on a path supports \gls{fq}, the client would use a delay-based \gls{cca}.
+However, while this solutions would achieve good results, the problem is that solutions, which require participation of every device in the Internet generally suffer from insufficient deployment because not every hardware vendor and network operator can be forced to implement the proposed solution for informing about \gls{fq} capability.
 

### 2020-09-29 src/plot_rtt_and_bandwidth.py

@@ -9,6 +9,7 @@ import pandas as pd
 import fnmatch
 plt.rcParams["font.family"] = "serif"
 plt.rcParams['pdf.fonttype'] = 42
+plt.rcParams["font.size"] = 8
 
 interval = 1
 pcap_file = sys.argv[1]
@@ -25,12 +26,10 @@ receiver_pcap = 'receiver_'+('_'.join(pcap_file.split('_')[1:]))
 new_rtt_command = f"./wintracker pcaps/{pcap_file}"
 
 # retransmissions_command = f"tshark -Y ip.src==192.168.0.1&&tcp.analysis.retransmission -r pcaps/{pcap_file} -Tfields -e frame.time_relative"
-
 # packets_command = f"tshark -Y ip.src==192.168.0.1 -r pcaps/{receiver_pcap} -q -z io,stat,{interval},ip.src==192.168.0.1"
 
 bytes_command = f"tshark -r pcaps/{receiver_pcap} -q -z io,stat,{interval},SUM(ip.len)ip.len&&ip.src==192.168.0.1"
-
-# bytes_command_total = f"tshark -r pcaps/{receiver_pcap} -q -z io,stat,0,SUM(ip.len)ip.len&&ip.src==192.168.0.1"
+bytes_command_total = f"tshark -r pcaps/{receiver_pcap} -q -z io,stat,0,SUM(ip.len)ip.len&&ip.src==192.168.0.1"
 
 # rtt_out = subprocess.run(rtt_command.split(), check=True, stdout=subprocess.PIPE, stderr=subprocess.STDOUT)
 if "tcp" in pcap_file:
@@ -46,7 +45,7 @@ else:
 		if fnmatch.fnmatch(f, f'*{timestamp}.txt'):
 			rtt_text_file = f
 			break
-	print("rtt_text_file", rtt_text_file)
+	# print("rtt_text_file", rtt_text_file)
 	df = pd.read_csv(f"pcaps/{rtt_text_file}")
 
 	ack_timestamps = df['timestamp'].tolist()
@@ -60,7 +59,6 @@ else:
 # # print("rtt_out", rtt_out)
 # rtt_results = [(float(item[0]), 1000*float(item[1])) for item in rtt_out if item[1] != ""]
 # print("average rtt", statistics.mean(list(zip(*rtt_results))[1]))
-print("average rtt", statistics.mean(rtts))
 # print("rtt_results", rtt_results[:100])
 
[…]

### 2020-09-29 src/test.py

@@ -10,6 +10,7 @@ import time
 import math
 import signal
 import numpy as np
+import fnmatch
 sys.path.insert(0, os.path.dirname(os.path.abspath(os.path.dirname(__file__))))
 
 try:
@@ -240,7 +241,7 @@ def run(vnet, prefix=""):
 
 			subprocess.check_output("chmod -R o+rw pcaps".split())
 
-		return client_out.decode("utf-8")
+		return client_out.decode("utf-8"), start_time
 
 if opt.run_scenario == "":
 	with virtnet.Manager() as context:
@@ -259,11 +260,25 @@ elif opt.run_scenario == "accuracy":
 					opt.time = 10
 
 					with virtnet.Manager() as context:
-						client_output = run(context)
+						client_output, timestamp = run(context, "accuracy")
 					assert client_output != ""
 					contained_vegas = "Starting Vegas" in client_output
 					contained_pcc = "Starting PCC Classic" in client_output
-					# print("client_output", client_output)
+
+					if opt.store_pcaps:
+						files = []
+						for file in os.listdir('pcaps'):
+							if fnmatch.fnmatch(file, f'sender_*{timestamp}.pcap'):
+								files.append(file)
+						assert len(files) == 1, len(files)
+					command = f"python3 ./plot_rtt_and_bandwidth.py {files[0]} no_plotting"
+					print("command", command)
+					output = subprocess.check_output(command.split())
+					output_lines = output.decode("utf-8").split("\n")[:2]
+					throughput = float(output_lines[0].split(" ")[-1])
+					rtt = float(output_lines[1].split(" ")[-1])
+
+					print("throughput", throughput, "rtt", rtt)
 
 					results_dict[(bw, delay, buffer, fq)] = (contained_vegas, contained_pcc)
 
@@ -308,7 +323,7 @@ elif opt.run_scenario == "competing_flow":
 	opt.qdisc = "fq"
 	opt.two_iperfs = False
 	with virtnet.Manager() as context:
-		client_output = run(context, "competing_flow_fq")
+		client_output, timestamp = run(context, "competing_flow_fq")
 
 	try:
 		del os.environ["START_VEGAS"]
@@ -326,7 +341,7 @@ elif opt.run_scenario == "competing_flow":
 	print("Starting pfifo experiment")
 	opt.qdisc = "pfifo"
 	with virtnet.Manager() as context:
[…]

### 2020-09-30 src/plot_rtt_and_bandwidth.py
Add evaluation function

@@ -98,8 +98,8 @@ bytes_results = [(float(item[0][0]), float(item[0][1]), float(item[1])) for item
 # divided = [(first[0], first[1], first[2]/second[2]) for first, second in zip(bytes_results, packets_results)]
 # print("divided", divided)
 
-print("avg_rtt", statistics.mean(rtts))
 print("throughput", sum_part/(end_time-start_time))
+print("avg_rtt", statistics.mean(rtts))
 
 if len(sys.argv) > 2 and sys.argv[2] == "no_plotting":
 	quit()

### 2020-09-30 src/test.py

@@ -265,38 +265,70 @@ elif opt.run_scenario == "accuracy":
 					contained_vegas = "Starting Vegas" in client_output
 					contained_pcc = "Starting PCC Classic" in client_output
 
-					if opt.store_pcaps:
-						files = []
-						for file in os.listdir('pcaps'):
-							if fnmatch.fnmatch(file, f'sender_*{timestamp}.pcap'):
-								files.append(file)
-						assert len(files) == 1, len(files)
-					command = f"python3 ./plot_rtt_and_bandwidth.py {files[0]} no_plotting"
-					print("command", command)
-					output = subprocess.check_output(command.split())
-					output_lines = output.decode("utf-8").split("\n")[:2]
-					throughput = float(output_lines[0].split(" ")[-1])
-					rtt = float(output_lines[1].split(" ")[-1])
-
-					print("throughput", throughput, "rtt", rtt)
-
 					results_dict[(bw, delay, buffer, fq)] = (contained_vegas, contained_pcc)
 
-	invalid = 0
+	invalids = []
+	false_predictions = []
 	predictions = []
 	for (bw, delay, buffer, fq), (is_vegas, is_pcc) in results_dict.items():
 		is_invalid = (not is_vegas and not is_pcc)
-		invalid += is_invalid
+		if is_invalid:
+			invalids.append(((bw, delay, buffer, fq), (is_vegas, is_pcc)))
 		if not is_invalid:
 			predictions.append((fq, is_vegas))
+			if fq != is_vegas:
+				false_predictions.append(((bw, delay, buffer, fq), is_vegas))
 
-	print("invalid", invalid, "total", len(results_dict))#, "results", results_dict)
+	print("invalids", len(invalids), "total", len(results_dict))
+	print("invalids", invalids)
 	confusion_matrix_input = list(zip(*predictions))
-	# print("confusion_matrix_input", confusion_matrix_input)
 	accuracy_score = sklearn.metrics.accuracy_score(*confusion_matrix_input)
 	print("accuracy_score", accuracy_score)
 	confusion_matrix = sklearn.metrics.confusion_matrix(*confusion_matrix_input)
 	print("confusion_matrix", confusion_matrix)
+	print("false_predictions", false_predictions)
+
+elif opt.run_scenario == "evaluation":
+	results_dict = {}
+	opt.store_pcaps = True
+	for bw_index, bw in enumerate(np.linspace(5,50,opt.how_many_values_per_parameter)):
[…]

### 2020-09-30 paper/paper.tex
Add related work

@@
 
-%\section{Related work}
+\section{Related work}
+
+Since our work involves using active measurements for the startup phase of \gls{cc} as well as proposing a new measurement technique that detects \gls{aqm}, we will present related work related to these two fields.
+
+\subsubsection{\gls{cc} using Active Measurements}
+Several approaches have been recently proposed which aim to incorporate active measurements into \glspl{cca}.
+\cite{dong_pcc_2015,dong_pcc_2018} introduced the \gls{cca} \textit{PCC}, which uses active measurements to find the sending rate which optimizes a given reward function.
+\cite{cardwell_bbr:_2016} introduce the \textit{BBR} \gls{cca}, which uses active measurement to create a model of the network connection.
+This model is then used to determine the optimal sending rate.
+\cite{bachl_rax_2019} use Reinforcement Learning to learn policies \gls{cc}.
+This is done by performing active measurements by varying the congestion window to evaluate which value gives the highest throughput and lowest packet loss.
+
+\subsubsection{Flow startup techniques} Besides the classical TCP slow start algorithm \cite{stevens_tcp_1997} also some other proposals were made such as Quick Start \cite{jain_quick-start_2019}, which aims to speed up flow start by routers telling the end point, which rate they deem appropriate.
+\cite{jarvinen_congestion_2019} investigate the impact of \gls{aqm} and \gls{cc} during flow startup.
+They inspect a variety of different \gls{aqm} algorithms and also investigate \gls{fq}.
+
+\subsubsection{Detecting \gls{aqm} mechanisms} We know of only one work \cite{kargar_bideh_tada_2016} which proposes a method to detect and distinguish certain popular \gls{aqm} mechanisms.
+However, unlike our work, it doesn't consider \gls{fq}.
 

### 2020-09-30 src/test.py

@@ -299,7 +299,7 @@ elif opt.run_scenario == "evaluation":
 				opt.delay = int(round(delay))
 				opt.buffer_size = int(round(buffer))
 				opt.qdisc = "fq" if fq else "pfifo"
-				opt.time = 10
+				opt.time = 20
 
 				with virtnet.Manager() as context:
 					client_output, timestamp = run(context, "accuracy")

### 2020-09-30 paper/paper.tex
Append related work

@@
 %
-\author{Maximilian Bachl\inst{1}\orcidID{0000-0001-9482-7428} \and
-Joachim Fabini\inst{2,3}\orcidID{0000-0002-8285-1591} \and
-Tanja Zseby\inst{3}\orcidID{0000-0002-5391-467X}}
+\author{Anonymous authors}
 %
-\authorrunning{M. Bachl et al.}
+\authorrunning{Anonymous authors}
 % First names are abbreviated in the running head.
@@
 %
-\institute{TU Wien, Vienna, Austria; 
-\email{firstname.lastname@tuwien.ac.at}}
+%\author{Maximilian Bachl\inst{1}\orcidID{0000-0001-9482-7428} \and
+%Joachim Fabini\inst{2,3}\orcidID{0000-0002-8285-1591} \and
+%Tanja Zseby\inst{3}\orcidID{0000-0002-5391-467X}}
+%%
+%\authorrunning{M. Bachl et al.}
+%% First names are abbreviated in the running head.
+%% If there are more than two authors, 'et al.' is used.
+%%
+%\institute{TU Wien, Vienna, Austria; 
+%\email{firstname.lastname@tuwien.ac.at}}
 %
@@
 They inspect a variety of different \gls{aqm} algorithms and also investigate \gls{fq}.
+\cite{kuhlewind_chirping_2010} propose ``chirping'' for flow startup to estimate the available bandwidth.
+This works by sending packets with different gaps between them to verify what the actual link speed is.
 
-\subsubsection{Detecting \gls{aqm} mechanisms} We know of only one work \cite{kargar_bideh_tada_2016} which proposes a method to detect and distinguish certain popular \gls{aqm} mechanisms.
+%\subsubsection{Detecting \gls{aqm} mechanisms} We know of only one work \cite{kargar_bideh_tada_2016} which proposes a method to detect and distinguish certain popular \gls{aqm} mechanisms.
 However, unlike our work, it doesn't consider \gls{fq}.

### 2020-10-01 paper/paper.tex
Amend paper

@@
 However, so far no end host based algorithm to detect fair queuing exists.
-Our contribution is the development of an algorithm that detects fair queuing at flow startup and chooses delay-based congestion control in this case.
+Our contribution is the development of an algorithm that detects fair queuing at flow startup and chooses delay-based congestion control if there is fair queuing.
 Otherwise, loss-based congestion control can be used as a backup option.
@@
 This unfairness can be mitigated by using \gls{fq} at the bottleneck link, isolating each flow from all other flows and assigning each flow an equal share of bandwidth \cite{dumazet_pkt_sched:_2013}.
-Then, the loss-based flows cannot obtain the bandwidth that would actually belong to the loss-based flows anymore.
+Then, the loss-based flows cannot obtain the bandwidth that would actually belong to the delay-based flows anymore.
 In this case, delay-based \glspl{cca} fulfill their goal of high bandwidth and low delay.
 
-While \gls{fq} solves many problems regarding \gls{cc}, it is still not ubiquitously deployed in every bottleneck link.
+While \gls{fq} solves many problems regarding \gls{cc}, it is still not ubiquitously deployed at every bottleneck link.
 If, however, flows could know if there's \gls{fq} or not, they could dynamically adapt their \gls{cc}.
@@
 \section{Related work}
-
-Since our work involves using active measurements for the startup phase of \gls{cc} as well as proposing a new measurement technique that detects \gls{aqm}, we will present related work related to these two fields.
+Since our work involves using active measurements for the startup phase of \gls{cc} as well as proposing a new measurement technique that detects \gls{aqm}, we will present related work related to these fields.
 
@@
 This model is then used to determine the optimal sending rate.
-\cite{bachl_rax_2019} use Reinforcement Learning to learn policies \gls{cc}.
+\cite{bachl_rax_2019} use Reinforcement Learning to learn \gls{cc} policies.
 This is done by performing active measurements by varying the congestion window to evaluate which value gives the highest throughput and lowest packet loss.
@@
 This works by sending packets with different gaps between them to verify what the actual link speed is.
+\cite{mittal_recursively_2014} propose speeding up flow startup by using packets of different priority.
+The link is flooded with low-priority packets and in case the link is not fully utilized during startup, these low-priority packets are going to be transmitted, allowing the starting flow to achieve high throughput.
+If the link is already saturated, the low-priority packets are going to be discarded and no other flow suffers.
 
-%\subsubsection{Detecting \gls{aqm} mechanisms} We know of only one work \cite{kargar_bideh_tada_2016} which proposes a method to detect and distinguish certain popular \gls{aqm} mechanisms.
+\subsubsection{Detecting \gls{aqm} mechanisms} There are very few works which investigate methods to detect the presence of an \gls{aqm}.
+\cite{kargar_bideh_tada_2016} propose a method to detect and distinguish certain popular \gls{aqm} mechanisms.
 However, unlike our work, it doesn't consider \gls{fq}.
+\cite{baykal_detection_2017} propose a machine learning based approach to fingerprint \gls{aqm} algorithm.
+Also, \gls{fq} is not considered.
 
@@
 
-It is necessary to wait for packet loss in both connections because only then it is certain that the link is saturated and our algorithm assumes that the link is saturated. 
+It is necessary to wait for packet loss in both connections because only then it is certain that the link is saturated. Our algorithm assumes that the link is saturated. 
 \item If \gls{fq} is used, the $\textit{loss ratio}$ is 2, otherwise it is 1. We choose 1.5 as the cutoff value. 
@@
 On this bottleneck link, the queue is either managed by \gls{fq} or by a shared queue (no \gls{fq}).
-\autoref{fig:throughput} shows how the sending rate (observed before the bottleneck link) is doubled after each \gls{rtt}.
+Figure~\ref{fig:throughput} shows how the sending rate (observed before the bottleneck link) is doubled after each \gls{rtt}.
 Furthermore, it shows that flow 2 has double the sending rate of flow 1.
@@
 This can be seen for example when looking at the deployment of IPv6, which requires the participation of every router on the way and every end device.
-However, multipath TCP \cite{handley_tcp_2012} and QUIC are becoming increasingly prevalent and Google's QUIC implementation has multipath support \cite{google_quiche_2020}.
+However, multipath TCP \cite{handley_tcp_2012} and QUIC, which enable recombining data of several flows, are becoming increasingly prevalent and Google's QUIC implementation has multipath support \cite{google_quiche_2020}.
 We thus think that the actual obstacles to deployment of our solution are constantly decreasing since QUIC (mostly Google traffic) already accounts for around $\frac{1}{10}$ of all Internet traffic \cite{ruth_first_2018} and is growing and, for example, Apple's macOS and iOS support multipath TCP \cite{apple_about_2018}.
@@
 For the experiments with \gls{fq} we use the \texttt{fq} module for Linux, for the experiments with a shared queue we use \texttt{pfifo}.
-
-\subsection{Accuracy of the \gls{fq} detection mechanism}
-
-To evaluate if our proposed mechanism correctly recognizes, we perform experiments using a wide range of network configurations: We vary bandwidth from 5 to 50\,Mbps, delay from 10 to 100\,ms and the buffer size from 1 to 100 packets, using a grid of these parameters.
-For each parameter, we take 5 values.
-Thus, we run experiments with 125 different configurations.
-We run each configuration once with \gls{fq} and once without.
-Results show that for experiments without \gls{fq} (using a shared queue), in over 99\% of cases, the absence of \gls{fq} was correctly detected and in less than 1\% of cases, our algorithm detected \gls{fq} even though there was no \gls{fq}.
-For experiments, during which \gls{fq} was actually deployed on the bottleneck link, this was correctly detected in 97\% of cases and wrongly in 3\% of the time.
-The overall accuracy is thus 98\%, which we consider sufficient.
-
-To verify the behavior of our algorithm under the presence of cross traffic, we repeated the above experiments but with one bulk traffic flow using Cubic running at the same time, which we start 5 seconds before so that it has sufficient time to gain full bandwidth.
-Also under these circumstances we achieve 98\% detection accuracy using our algorithm.
 
@@
 }
-\caption{Throughput and delay of a flow controlled by the loss-based \gls{cca} Cubic on a link with a speed of 10\,Mbps, a delay of 10\,ms and a buffer size of 100 packets.}
+\caption{Throughput and delay of a flow controlled by the loss-based \gls{cca} Cubic on a link with a speed of 10\,Mbps, a delay of 10\,ms and a buffer size of 100 packets (\texttt{pfifo}'s default value).}
 \label{fig:just_one_flow_tcp}
@@
 \subsection{Comparing flows with different Congestion Controls and AQM mechanisms}
-In this scenario, one flow with our mechanism shares a bottleneck link with a flow using a loss-based congestion control.
+In this scenario, we show an example of a flow with our mechanism that shares a bottleneck link with a flow using a loss-based congestion control.
 
 \subsubsection{No \gls{fq}}
-\autoref{fig:pfifo_tcp} and \autoref{fig:pfifo_udp} show the loss-based \gls{cca} Cubic compete against our delay-based one on a link without \gls{fq}.
+The examples in \autoref{fig:pfifo_tcp} and in \autoref{fig:pfifo_udp} show the loss-based \gls{cca} Cubic compete against our delay-based \gls{cca} on a link without \gls{fq}.
+It is important to note that we do not use our \gls{fq} detection mechanism here but instead choose delay-based \gls{cc} from the beginning to demonstrate the starvation of delay-based flows that happens when they compete against loss-based ones without \gls{fq}.
 In the scenario, the loss-based flow starts 5 seconds earlier and after 5 seconds, the delay-based flow joins.
@@
 
-For a link with fair queuing, the following results show that throughput is identical for our flow (\autoref{fig:fq_udp}) and the Cubic flow (\autoref{fig:fq_tcp}), just that the Cubic flow's delay is more than two times higher than the delay-based one's.
+For a link with fair queuing, the following example results show that throughput is identical for our flow (\autoref{fig:fq_udp}) and the Cubic flow (\autoref{fig:fq_tcp}), just that the Cubic flow's delay is more than two times higher than the delay-based one's.
 Thus, our mechanism is effective to detect if the bottleneck uses fair queuing and use a delay-based congestion control if it is the case and a loss-based one otherwise.
+
+\subsection{Accuracy of the \gls{fq} detection mechanism}
+
+To evaluate if our proposed mechanism correctly recognizes \gls{fq} in a systematic manner, we perform experiments using a wide range of network configurations: We vary bandwidth from 5 to 50\,Mbps, delay from 10 to 100\,ms and the buffer size from 1 to 100 packets, using a grid of these parameters.
+For each parameter, we take 5 values, evenly spaced.
+Thus, we run experiments with 125 different configurations.
+We run each configuration once with \gls{fq} and once without.
+Results show that for experiments without \gls{fq} (using a shared queue), in over 99\% of cases, the absence of \gls{fq} was correctly detected and in less than 1\% of cases, our algorithm detected \gls{fq} even though there was no \gls{fq}.
+For experiments, during which \gls{fq} was actually deployed on the bottleneck link, this was correctly detected in 97\% of cases and wrongly in 3\% of the time.
+The overall accuracy is thus 98\%, which we consider sufficient.
+
+To verify the behavior of our algorithm under the presence of cross traffic, we repeated the above experiments but with one bulk traffic flow using Cubic running at the same time, which we start 5 seconds before so that it has sufficient time to gain full bandwidth.
+Also under these circumstances we also achieve 98\% detection accuracy using our algorithm.
+
+\subsection{Systematic evaluation}
+
+Besides the examples we provided, we also conducted a more systematic comparison of the throughput and delay that our algorithm achieves: We use the same scenario as above and again vary bandwidth from 5 to 50\,Mbps, delay from 10 to 100\,ms and the buffer size from 1 to 100 packets.
+For each parameter, we take 3 values, evenly spaced.
+This means that there are 27 distinct scenarios.
+We use \gls{fq} at the bottleneck link.
+We run each scenario both using Cubic as well as our algorithm.
+We average the achieved throughputs and average delays from all experiments for both Cubic as well as our algorithm.
+For Cubic, the mean throughput is 18\,Mbit/s and the mean \gls{rtt} is 109\;ms while for our algorithm it is 25\,Mbit/s and 74\,ms.
+This shows that our algorithm can correctly determine that there is \gls{fq} and switch to the delay-based \gls{cc} and that our algorithm achieves significantly lower delay.
+Surprisingly, our algorithm also achieves higher throughput, which is astonishing considering that loss-based \glspl{cca} such as Cubic should be more aggressive in filling the link.
+When looking at the results, we found out that Cubic performs very poorly in scenarios with a buffer of only 1 packet since it doesn't pace its packets and thus each time when two packets are sent right after another, packet loss occurs and Cubic lowers its rate.
+
+%\begin{table}[t]
+%\caption{Comparison of for loss-based \glspl{cca} on a link of 100\;Mbit/s with an \gls{rtt} of 50\;ms.
+Utilization in \% of maximum capacity; \gls{rtt} in milliseconds.
+For Cubic fq achieves good performance.
+This is because its default buffer size of 100 packets coincidentially works well for this specific link speed and \gls{rtt}.
+We take the mean of 10 runs for each scenario.} \label{tab:performance_results}
+%\centering
+%\begin{tabular}{l r r r r r r} \toprule
+%& \multicolumn{2}{l}{fq} & \multicolumn{2}{l}{fq\_codel} & \multicolumn{2}{l}{eggs} \\
+%\cmidrule(r){2-3} \cmidrule(lr){4-5} \cmidrule(l){6-7}
+%& Util. & \gls{rtt} & Util. & \gls{rtt} & Util. & \gls{rtt} \\ \midrule
+%Cubic	& 96.2 & 54 & 92.6 & 52 & 98.5 & 66.5 \\
+%Reno	& 84.1 & 52.7 & 81.1 & 51.5 & 97.9 & 98.2 \\
+%\bottomrule
+%\end{tabular}
+%\end{table}
 
@@
 However, when there are two flows where one flow's sending rate is a multiple of the other's, an interleaving effect occurs: flow 1 would always send each packet slightly after flow 2 sends a packet.
-Then, when the packets arrive at the bottleneck, flow 2's packet, which arrives a bit earlier, fills up the buffer and flow 1's packet is consequently dropped.
+Then, when the packets arrive at the bottleneck, flow 2's packet, which arrives a bit earlier, fills up the buffer and flow 1's packet is consequently always dropped.
 As a solution, we do not send packets at regular intervals during startup but always add a small random value: For example, if the sending interval were 1\,ms, we would send the packet randomly in the interval from 0.5 to 1.5\,ms.
@@
 
-One problem of our proposed solution is when the bottleneck link changes while a flow is running.
-For example, the bottleneck link in the beginning could be the home router, while later is it becoming the another router deeper in the core network.
-The problem occurs if the home router supports \gls{fq} and our algorithm detects this while the router in the core network, which become the bottleneck later on, doesn't support \gls{fq}.
-Then, our algorithm is going to use a delay-based \gls{cc} because that's the case in the beginning of the flow.
+One problem of our proposed solution occurs when the bottleneck link changes while a flow is running.
+For example, the bottleneck link in the beginning could be the home router, while later is it becoming another router deeper in the core network.
+The problem occurs if the home router supports \gls{fq} and our algorithm detects this while the router in the core network, which becomes the bottleneck later on, doesn't support \gls{fq}.
+Then, our algorithm is going to use a delay-based \gls{cc} because that's what it determined to be correct in the beginning of the flow.
 However, later when the bottleneck changes, our algorithm is still using delay-based \gls{cc} even though the current bottleneck doesn't support \gls{fq} anymore.
@@
 An alternative to our algorithm could be that the routers tell the endpoints if they support \gls{fq} or not.
-Such a solution could, for example, use an IP or a TCP extension, which every router uses to indicate if they're capability of \gls{fq}.
+Such a solution could, for example, use an IP or a TCP extension, which every router uses to indicate if they're capable of \gls{fq}.
 Then, if every device on a path supports \gls{fq}, the client would use a delay-based \gls{cca}.
@@
 Especially for applications that are very delay-sensitive, such as video conferencing and cloud gaming, the adoption of delay-based \gls{cc} can result in a significant increase of \gls{qoe}.
-We demonstrated that our method works using a prototype implementation using PCC but we are also confident that a similar flow startup method can also be integrated with other \glspl{cca} such as BBR \cite{cardwell_bbr:_2016}.
+We demonstrated that our method works using a prototype implementation using PCC but we are also confident that a similar flow startup method can also be integrated with other \glspl{cca} such as BBR.
 

### 2020-10-02 paper/paper.tex
Review JF

@@
 %
-\title{Detecting Fair Queuing to Enable Better Congestion Control}
-%
+%JF: \title{Detecting Fair Queuing to Enable Better Congestion Control}
+%JF: Alternative Vorschläge für Titel, etwas flüssiger:
+\title{Optimizing Congestion Control Through Fair Queuing Detection}
+%JF: \title{Optimum Congestion Control Selection Through Fair Queuing Detection}
+
 %\titlerunning{Abbreviated paper title}
@@
 %
+
+%JF: FIXME: A fundamental problem that we should perhaps address in a brief statement: The proposed method does not work for cases in which load balancers are on the path.
+Distinct flows are likely to result in distinct hashes/flowIDs, meaning that the balancer will potentially assign them to distinct paths.
+So we may add a restrictive statement to the discussion that we assume that all flows share the same bottleneck?
+In addition we may state that no service/port priorization is in place.
+%JF Any other idea?
+
 \begin{abstract}
 Low delay is an explicit requirement for applications such as cloud gaming and video conferencing.
-Delay-based congestion control can achieve the same throughput but significantly smaller delay than loss-based one and is thus ideal for these scenarios.
-However, when a delay- and a loss-based flow share a bottleneck, the loss-based one can ``steal'' all the bandwidth and the delay-based one ``starves''.
-Fair queuing at the bottleneck link, which isolates each flow and provides the same bandwidth for everyone, solves this problem.
+Delay-based congestion control can achieve the same throughput but significantly smaller delay than loss-based one and is thus ideal for these applications.
+However, when a delay- and a loss-based flow compete for a bottleneck, the loss-based one can monopolize all the bandwidth and starve the delay-based one.
+Fair queuing at the bottleneck link solves this problem by assigning an equal share of the available bandwidth to each flow.
 However, so far no end host based algorithm to detect fair queuing exists.
@@
 Emerging applications such as cloud gaming \cite{jarschel_evaluation_2011} and remote virtual reality \cite{elbamby_toward_2018} applications require high throughput and low delay.
-Furthermore, for 5G, ultra low latency communications have been made a priority \cite{li_5g_2018}.
+Furthermore, ultra low latency communications have been made a priority for 5G \cite{li_5g_2018}.
 We argue that for achieving high throughput as well as low delay, congestion control must be taken into account.
@@
 While delay-based \glspl{cca} fulfill their goal of low delay and high throughput, they cannot handle competing flows with different \glspl{cca} well \cite{turkovic_fifty_2019, turkovic_interactions_2019}.
-This is especially prevalent when they have to compete against loss-based \glspl{cca}.
-The problem is that loss-based \glspl{cca} are more ``aggressive''.
-While the delay-based ones already back away as soon as queuing delay increases, the loss-based ones continue increasing their throughput until the buffer is full and packet loss occurs.
-This results in the loss-based flows gaining all the bandwidth and the delay-based ones ``starving'' \cite{hock_toward_2016,yuan-cheng_lai_improving_2001,awdeh_comparing_2004}.
+This is especially prevalent when they compete against loss-based \glspl{cca}, the latter being more ``aggressive''.
+While the delay-based \glspl{cca} back off as soon as queuing delay increases, the loss-based ones continue increasing their throughput until the buffer is full and packet loss occurs.
+This results in the loss-based flows monopolize the available bandwidth and starve the delay-based ones \cite{hock_toward_2016,yuan-cheng_lai_improving_2001,awdeh_comparing_2004}.
 
 This unfairness can be mitigated by using \gls{fq} at the bottleneck link, isolating each flow from all other flows and assigning each flow an equal share of bandwidth \cite{dumazet_pkt_sched:_2013}.
-Then, the loss-based flows cannot obtain the bandwidth that would actually belong to the delay-based flows anymore.
-In this case, delay-based \glspl{cca} fulfill their goal of high bandwidth and low delay.
+Consequently, loss-based flows can no longer acquire bandwidth that delay-based flows have released due to their network-friendly behavior.
+In this case, delay-based \glspl{cca} achieve their goal of high bandwidth and low delay.
 
-While \gls{fq} solves many problems regarding \gls{cc}, it is still not ubiquitously deployed at every bottleneck link.
-If, however, flows could know if there's \gls{fq} or not, they could dynamically adapt their \gls{cc}.
-If there is \gls{fq} they would use a delay-based \gls{cca} and otherwise a loss-based one.
-Our contribution is the design and evaluation of such a mechanism that determines the presence of \gls{fq} during the startup phase of a flow's \gls{cca} and, at the end of the startup, sets the \gls{cca} accordingly.
+While \gls{fq} solves many problems regarding \gls{cc}, it is still not ubiquitously deployed.
+Flows can benefit from knowledge of \gls{fq} enabled bottleneck links on the path to dynamically adapt their \gls{cc}.
+In the case of \gls{fq} they can use a delay-based \gls{cca} and otherwise revert to a loss-based one.
+Our contribution is the design and evaluation of such a mechanism that determines the presence of \gls{fq} during the startup phase of a flow's \gls{cca} and sets the \gls{cca} accordingly at the end of the startup.
 
-We show that this solution reliably determines the presence of fair queuing and that by using this approach, queuing delay can be considerably lowered while keeping throughput high if \gls{fq} is detected.
-In case no \gls{fq} is detected, our algorithm detects this as well and a loss-based \gls{cc} is used, which does not starve when facing competition from the delay-based flow.
+We show that this solution reliably determines the presence of fair queuing and that by using this approach, queuing delay can be considerably lowered while maintaining throughput high if \gls{fq} is detected.
+In case no \gls{fq} is enabled, our algorithm detects this as well and a loss-based \gls{cc} is used, which does not starve when facing competition from other loss-based flows.
+%JF: delay-based was imo mistaken here!
 While our mechanism is beneficial for network flows in general, we argue that it is most useful for long-running delay-sensitive flows, such as cloud gaming, video conferencing etc.
@@
 \section{Related work}
-Since our work involves using active measurements for the startup phase of \gls{cc} as well as proposing a new measurement technique that detects \gls{aqm}, we will present related work related to these fields.
+Our work depends on active measurements for the startup phase of \gls{cc} and proposes a new measurement technique to detect \gls{aqm}. This section discusses preliminary work related to these fields.
 
+%JF: Evtl.statt subsubsection Absätze mit Titel fett verwenden um doch einigen Platz zu sparen?
 \subsubsection{\gls{cc} using Active Measurements}
@@
 
-\subsubsection{Flow startup techniques} Besides the classical TCP slow start algorithm \cite{stevens_tcp_1997} also some other proposals were made such as Quick Start \cite{jain_quick-start_2019}, which aims to speed up flow start by routers telling the end point, which rate they deem appropriate.
+%JF: Evtl.statt subsubsection Absätze mit Titel fett verwenden um doch einigen Platz zu sparen?
+\subsubsection{Flow startup techniques} Besides the classical TCP slow start algorithm \cite{stevens_tcp_1997} some other proposals were made such as Quick Start \cite{jain_quick-start_2019}, which aims to speed up flow start by routers informing the end point, which rate they deem appropriate.
 \cite{jarvinen_congestion_2019} investigate the impact of \gls{aqm} and \gls{cc} during flow startup.
@@
 
-\subsubsection{Detecting \gls{aqm} mechanisms} There are very few works which investigate methods to detect the presence of an \gls{aqm}.
+%JF: Evtl.statt subsubsection Absätze mit Titel fett verwenden um doch einigen Platz zu sparen?
+\subsubsection{Detecting \gls{aqm} mechanisms} There are few publications that investigate methods to detect the presence of an \gls{aqm}.
 \cite{kargar_bideh_tada_2016} propose a method to detect and distinguish certain popular \gls{aqm} mechanisms.
@@
 \cite{baykal_detection_2017} propose a machine learning based approach to fingerprint \gls{aqm} algorithm.
-Also, \gls{fq} is not considered.
+\gls{fq} is not considered, too.
 
@@
 \begin{enumerate}
-\item During flow startup, determine if \gls{fq} is used at the bottleneck or not.
-\item If this is determined, change to a delay-based \gls{cca} if \gls{fq} is used and to a loss-based one otherwise.
+\item During flow startup, determine whether the bottleneck link deploys \gls{fq} or not.
+\item If \gls{fq} is used change to a delay-based \gls{cca} or revert to a loss-based \gls{cca} otherwise.
 \end{enumerate}
@@
 \end{align}
-Basically, this metric indicates if flow 2, which has a higher sending rate, also achieves a higher receiving rate (goodput) than flow 1. If this is the case, there's no \gls{fq}, otherwise there is.
+
+%JF: FIXME: I'd delete eq (2) above or fit it into one row - no substantial information.
+%JF: Perhaps abbreviate: sendRate flow1, recvRate flow1...
+This loss ratio metric indicates if flow 2, which has a higher sending rate, achieves a higher receiving rate (goodput) than flow 1. If this is the case, there's no \gls{fq}, otherwise there is.
 
@@
 
-Figure~\ref{fig:goodput_fq}, on the other side, shows the receiving rate of the two flows after the bottleneck link in case \gls{fq} is deployed.
-In the third \gls{rtt}, flow 1 and flow 2 achieve the same receiving rate because of the \gls{fq}, which makes sure that each flow only gets its fair share.
+Figure~\ref{fig:goodput_fq} shows the receiving rate of the two flows measured after the bottleneck link in case \gls{fq} is deployed.
+In the third \gls{rtt}, flow 1 and flow 2 achieve the same receiving rate because of the \gls{fq}, which makes sure that each flow gets an equal share of the total capacity.
 
@@
 
-If the absence of \gls{fq} detected, a loss based \gls{cc} is used as a fallback.
+If the absence of \gls{fq} is detected, a loss based \gls{cc} is used as a fallback.
 We use the algorithm \textit{PCC} proposed by \cite{dong_pcc_2015}.
@@
 The PCC code is based on the \textit{UDT} library \cite{gu_udt_2007}, which is a library that provides a reliable transport protocol similar to TCP on top of UDP.
-This approach is similar to the QUIC \cite{iyengar_quic_2018}.
+This approach is similar to QUIC \cite{iyengar_quic_2018}.
 An advantage of implementing our approach on top of UDP as opposed to TCP is that it is easier to develop and modify since no kernel module needs to be created and that it is more secure since the code is isolated in a user space process.
@@
 Our approach of detecting \gls{fq} relies on using two concurrent flows between the sender and the receiver.
-This means that during flow startup, data is transmitted in two flows.
-To be able to recombine the data from the both flows, the receiver has to be aware of the fact that the two flows belong together and also has to be able to recombine the data.
-This seems like a problem regarding deployment since solutions that require multiple parties to implement are generally hard to deploy on a wide scale in the Internet.
-This can be seen for example when looking at the deployment of IPv6, which requires the participation of every router on the way and every end device.
+During flow startup, user data is transmitted in two flows that the receiver must detect and reassemble.
+This may be a challenge for deployments since solutions that require cooperation of several instances are more difficult to deploy at a wide scale in the Internet.
+%This can be seen for example when looking at the deployment of IPv6, which requires the participation of every router on the way and every end device.
+%JF: FIXME: I'm sceptical that the examples below work well in all scenarios. Distinct flows end up on distinct paths, so they do not necessarily use the same bottleneck.
 However, multipath TCP \cite{handley_tcp_2012} and QUIC, which enable recombining data of several flows, are becoming increasingly prevalent and Google's QUIC implementation has multipath support \cite{google_quiche_2020}.
-We thus think that the actual obstacles to deployment of our solution are constantly decreasing since QUIC (mostly Google traffic) already accounts for around $\frac{1}{10}$ of all Internet traffic \cite{ruth_first_2018} and is growing and, for example, Apple's macOS and iOS support multipath TCP \cite{apple_about_2018}.
+We thus argue that the actual obstacles to deployment of our solution are constantly decreasing since QUIC (mostly Google traffic) already accounts for around %$\frac{1}{10}$
+$10\%$ of all Internet traffic \cite{ruth_first_2018} and is growing. Moreover, Apple's macOS and iOS support multipath TCP \cite{apple_about_2018} support.
 
@@
 
-First, we want to empirically demonstrate that delay-based \gls{cc} actually has a benefit over loss-based one.
-\autoref{fig:just_one_flow_tcp} shows that the loss-based \gls{cc} achieves a throughput of almost 10\,Mbps on a 10\,Mbps link, which means full utilization and an \gls{rtt} that is generally higher than 100\,ms.
-On the other side, \autoref{fig:just_one_flow_udp} shows an example flow using our simple delay-based \gls{cca} and a slow start algorithm similar to the one used by the classic TCP \gls{cca} \textit{New Reno} \cite{allman_tcp_1999}.
+First, we demonstrate empirically that delay-based \gls{cc} actually has a benefit over loss-based one.
+On one hand, \autoref{fig:just_one_flow_tcp} shows that the loss-based \gls{cc} achieves a throughput of almost 10\,Mbps on a 10\,Mbps link, which means full utilization and an \gls{rtt} that is  higher than 100\,ms.
+On the other hand, \autoref{fig:just_one_flow_udp} shows an example flow using our simple delay-based \gls{cca} and a slow start algorithm similar to the one used by the classic TCP \gls{cca} \textit{New Reno} \cite{allman_tcp_1999}.
 While the delay-based flow also achieves very high utilization, delay rarely exceeds 20\,ms, meaning that it is more than 5 times lower than the one of the loss-based \gls{cca} of \autoref{fig:just_one_flow_tcp}.
@@
 
+%JF: More compelling would be the use case when the delay-based flow is started first and the late-comer loss-based preempts it.
 \subsubsection{No \gls{fq}}
@@
 \caption{Throughput and delay of a flow controlled by our delay-based \gls{cca} on a link with a speed of 50\,Mbps, a delay of 10\,ms and a buffer size of 100 packets.
-The bottleneck is controlled by a shared queue (no \gls{fq}) and is shared with the flow of \autoref{fig:fq_tcp} and is shared with the flow of \autoref{fig:pfifo_tcp}.}
+The bottleneck is controlled by a shared queue (no \gls{fq}) and is shared with the flow of \autoref{fig:pfifo_tcp}.}
 \label{fig:pfifo_udp}
@@
 Besides the examples we provided, we also conducted a more systematic comparison of the throughput and delay that our algorithm achieves: We use the same scenario as above and again vary bandwidth from 5 to 50\,Mbps, delay from 10 to 100\,ms and the buffer size from 1 to 100 packets.
-For each parameter, we take 3 values, evenly spaced.
-This means that there are 27 distinct scenarios.
+For each parameter, we take 3 values, evenly spaced, totaling 27 distinct scenarios.
 We use \gls{fq} at the bottleneck link.
@@
 Surprisingly, our algorithm also achieves higher throughput, which is astonishing considering that loss-based \glspl{cca} such as Cubic should be more aggressive in filling the link.
-When looking at the results, we found out that Cubic performs very poorly in scenarios with a buffer of only 1 packet since it doesn't pace its packets and thus each time when two packets are sent right after another, packet loss occurs and Cubic lowers its rate.
+When analyzing the results we identified that Cubic performs very poorly in scenarios with a buffer of only 1 packet since it doesn't pace its packets and thus each time when two packets are sent right after another, packet loss occurs and Cubic lowers its rate.
 
@@
 This solves the aforementioned problem.
-The insight we gained from this is that while pacing is generally good, it is actually harmful if several flows compete at a bottleneck without \gls{fq} and one flow's sending rate is a multiple of the other's.
+The insight we gained from this is that while pacing is generally supportive, it is actually harmful if several flows compete at a bottleneck without \gls{fq} and one flow's sending rate is a multiple of the other's.
 
@@
 
-An alternative to our algorithm could be that the routers tell the endpoints if they support \gls{fq} or not.
+%JF: FIXME: Downside (if room left): it's not clear which element is the bottleneck link.
+If all of them are FQ or non-FW, then it's fine.
+Routing may result in other elements to be used that have not been probed.
+All routers along the path may need to provide periodic reports on queueing policy *and* on queue fill status, so the source can guess the bottleneck link.
+%JF: And one interesting observation (perhaps follow-up work, but likely discussed already): I'm curious how a series of FQ and non-FQ routers impacts on competing flows. 
+An alternative to our algorithm could be that all routers inform the endpoints if they support \gls{fq} or not.
 Such a solution could, for example, use an IP or a TCP extension, which every router uses to indicate if they're capable of \gls{fq}.
@@
 
-Finally, we think that our new flow startup method can help to increase the deployment of delay-based \gls{cc}.
+Concluding, we argue that our new flow startup method can help to increase the deployment of delay-based \gls{cc}.
 Especially for applications that are very delay-sensitive, such as video conferencing and cloud gaming, the adoption of delay-based \gls{cc} can result in a significant increase of \gls{qoe}.

### 2020-10-02 paper/paper.tex
Add Anomaly Detection meeting presentation

--- 2020-10-01 paper/paper.tex
@@
 We run each scenario both using Cubic as well as our algorithm.
-We average the achieved throughputs and average delays from all experiments for both Cubic as well as our algorithm.
-For Cubic, the mean throughput is 18\,Mbit/s and the mean \gls{rtt} is 109\;ms while for our algorithm it is 25\,Mbit/s and 74\,ms.
+We average the achieved throughputs and average delays from all experiments for both Cubic as well as our algorithm for 20 seconds each.
+For Cubic, the mean throughput is 18\,Mbit/s and the mean \gls{rtt} is 109\,ms while for our algorithm it is 25\,Mbit/s and 74\,ms.
 This shows that our algorithm can correctly determine that there is \gls{fq} and switch to the delay-based \gls{cc} and that our algorithm achieves significantly lower delay.

### 2020-10-02 paper/presentation.tex
Add Anomaly Detection meeting presentation

%
% Beispieldokument für TU beamer theme
%
% v1.0: 12.10.2014
% 
\documentclass[xcolor={dvipsnames}]{beamer}
\usetheme{TU}

\setbeamertemplate{caption}[numbered]
\usepackage{hyperref}
\usepackage{booktabs}
\usepackage{tabularx}
\usepackage{subfig}
% Macro to make entire row bold from https://tex.stackexchange.com/questions/309833/format-whole-row-of-table-as-bold
\newcommand\setrow[1]{\gdef\rowmac{#1}#1\ignorespaces}
\newcommand\clearrow{\global\let\rowmac\relax}
\clearrow

\title{Detecting Fair Queuing to Enable Better Congestion Control}

\author[M. Bachl et al.]{%
	\underline{Maximilian Bachl}\email{maximilian.bachl@tuwien.ac.at} \and Tanja Zseby \and Joachim Fabini
}

\institute{%
	Technische Universität Wien, Vienna, Austria
}

%Kann angepasst werden, wie es beliebt. Entweder das Datum der Präsentation oder das Datum der aktuellen Präsentations-Version.
%\date[\the\day.\the\month.\the\year]{\today}
\date[October 5, 2020]{October 5, 2020}

\begin{document}

\maketitle

\section{Introduction}

\begin{frame}{Loss-based Congestion Control}
\begin{itemize}
\item Currently loss-based congestion control (Cubic) is used in the Internet.
\item Constantly increases sending rate until buffer full and packet loss occurs. 
\item If packet loss occurs, lower sending rate. 
\end{itemize}
\pause
\begin{block}{Insight:}
Buffer always gets filled to the maximum $\rightarrow$ high delay.
\end{block}
\end{frame}

\begin{frame}{Delay-based Congestion Control}
\begin{itemize}
\item Increase sending rate until buffer fills and delay increases. 
\item Decrease sending rate if delay is too high.
\end{itemize}
\pause
\begin{block}{Insight:}
Same throughput but lower delay than loss-based congestion control. 
\end{block}
\end{frame}

\begin{frame}{Loss-based Congestion Control -- Example}
\begin{figure}[h]
\centering
\subfloat[Throughput (receiving rate)]{%
  \includegraphics[width=0.48\columnwidth]{{"figures/plots/throughput_1_just_one_flow_cubic_tcp_port9010_fq_10_10_20_1600183264284.pcap"}.pdf}%
  \label{fig:just_one_flow_tcp_throughput}
}\hfill
\subfloat[delay]{%
  \includegraphics[width=0.48\columnwidth]{{"figures/plots/rtt_1_just_one_flow_cubic_tcp_port9010_fq_10_10_20_1600183264284.pcap"}.pdf}%
  \label{fig:just_one_flow_tcp_delay}
}
\caption{Flow controlled by the loss-based Cubic on a link with a speed of 10\,Mbps, a delay of 10\,ms and a buffer size of 100 packets (\texttt{pfifo}'s default value).}
\label{fig:just_one_flow_tcp}
\end{figure}
\end{frame}

\begin{frame}{Delay-based Congestion Control -- Example}
\begin{figure}[h]
\centering
\subfloat[Throughput (receiving rate)]{%
  \includegraphics[width=0.48\columnwidth]{{"figures/plots/throughput_1_just_one_flow_vegas_udp_port9000_fq_10_10_20_1600183241479.pcap"}.pdf}%
}\hfill
\subfloat[delay]{%
  \includegraphics[width=0.48\columnwidth]{{"figures/plots/rtt_1_just_one_flow_vegas_udp_port9000_fq_10_10_20_1600183241479.pcap"}.pdf}%
}
\caption{Flow controlled by our delay-based algorithm on a link with a speed of 10\,Mbps, a delay of 10\,ms and a buffer size of 100 packets.}
\label{fig:just_one_flow_udp}
\end{figure}
\end{frame}

\begin{frame}{Comparing delay- and loss-based}
\begin{block}{Insight:}
Delay-based Congestion Control achieves the same throughput but much lower delay.\end{block}
\end{frame}

\begin{frame}{Delay- vs.~Loss-based -- Loss-based}
\begin{figure}[h]
\centering
\subfloat[Throughput (receiving rate)]{%
  \includegraphics[width=0.48\columnwidth]{{"figures/plots/throughput_1_competing_flow_pfifo_vegas_tcp_port9010_pfifo_10_50_20_1600182743408.pcap"}.pdf}%
}\hfill
\subfloat[delay]{%
  \includegraphics[width=0.48\columnwidth]{{"figures/plots/rtt_1_competing_flow_pfifo_vegas_tcp_port9010_pfifo_10_50_20_1600182743408.pcap"}.pdf}%
}
\caption{Flow controlled by the loss-based Cubic on a link with a speed of 50\,Mbps, a delay of 10\,ms and a buffer size of 100 packets. The bottleneck is controlled by a shared queue (no FQ) and is shared with the flow of Figure~\ref{fig:pfifo_udp}.}
\label{fig:pfifo_tcp}
\end{figure}
\end{frame}

\begin{frame}{Delay- vs.~Loss-based together -- Delay-based}
\begin{figure}[h]
\centering
\subfloat[Throughput (receiving rate)]{%
  \includegraphics[width=0.48\columnwidth]{{"figures/plots/throughput_1_competing_flow_pfifo_vegas_udp_port9000_pfifo_10_50_20_1600182743408.pcap"}.pdf}%
}\hfill
\subfloat[delay]{%
  \includegraphics[width=0.48\columnwidth]{{"figures/plots/rtt_1_competing_flow_pfifo_vegas_udp_port9000_pfifo_10_50_20_1600182743408.pcap"}.pdf}%
}
\caption{Flow (started 5\,s later) controlled by our delay-based algorithm on a link with a speed of 50\,Mbps, a delay of 10\,ms and a buffer size of 100 packets. The bottleneck is controlled by a shared queue (no FQ) and is shared with the flow of Figure~\ref{fig:pfifo_tcp}.}
\label{fig:pfifo_udp}
\end{figure}
\end{frame}

\begin{frame}{Delay- vs.~Loss-based}
\begin{block}{Insight:}
When loss-based and delay-based congestion control compete, loss-based one eats delay-based one's lunch!\end{block}
\end{frame}

\begin{frame}{Fair Queuing}
\begin{itemize}
\item Deployed in switches and routers
\item Each flow gets separate queue
\item All flows peacefully coexist, loss-based and delay-based ones
\end{itemize}
\begin{block}{Insight:}
Fair queuing enables the use of delay-based congestion control!
\end{block}
\end{frame}

\begin{frame}{Summary}
\begin{itemize}
\item Delay-based congestion control offers same throughput but lower delay than loss-based ones
\item When competing, loss-based flows will gain all bandwidth. Delay-based flows ``starve''. 
\item Fair queuing solves this
\end{itemize}
\begin{block}{Insight:}
If flows could detect the presence of fair queuing, they could switch to delay-based congestion control!
\end{block}
\end{frame}

\section{Concept}
\begin{frame}{Overall Algorithm}
\begin{enumerate}
\item During flow startup, determine if fair queuing is used at the bottleneck or not.
\item If this is determined, change to a delay-based congestion control algorithm if fair queuing is used and to a loss-based one otherwise.
\end{enumerate}
\end{frame}

\begin{frame}{Fair queuing detection algorithm}
\begin{enumerate}
\item Launch two concurrent flows, flow 1 and flow 2. Initialize flow 1 with some starting sending rate and flow 2 with two times that rate. 
\item After every delay, if no packet loss occurred, double the sending rate of both flows. 
\item If packet loss occurred in both flows in the previous delay, calculate the following metric if each flow got an equal share of bandwidth (fair queuing) or not (no fair queuing).
\end{enumerate}
\end{frame}

\begin{frame}{Fair queuing detection algorithm -- Example}
\begin{figure}
\centering
\subfloat[Sending rate]{\includegraphics[width=0.33\columnwidth]{{"figures/fq_illustration_throughput"}.pdf}
\label{fig:throughput}}
\subfloat[Receiving rate\\ (no fair queuing).]{\includegraphics[width=0.33\columnwidth]{{"figures/fq_illustration_goodput_no_fq"}.pdf}
\label{fig:goodput_no_fq}}
\subfloat[Receiving rate\\ (fair queuing).]{\includegraphics[width=0.33\columnwidth]{{"figures/fq_illustration_goodput_fq"}.pdf}
\label{fig:goodput_fq}}
\caption{Example of our proposed flow startup mechanism.}
\label{fig:illustration}
\end{figure}
\end{frame}

\begin{frame}{Simple delay-based congestion control}
\begin{enumerate}
\item If $\textit{current rtt}$ $\leq$ $\textit{smallest rtt ever measured on the connection}$ $+$ $5\textit{ms}$ then increase the sending rate by 1\%.
\item Otherwise ($\textit{current rtt}$ $>$ $\textit{smallest rtt ever measured on the connection}$ $+$ $5\textit{ms}$) decrease the sending rate by 5\%.
\end{enumerate}
\end{frame}

\section{Results}
\begin{frame}{Cubic -- Example}
\begin{figure}[h]
\centering
\subfloat[Throughput (receiving rate)]{%
  \includegraphics[width=0.48\columnwidth]{{"figures/plots/throughput_1_competing_flow_fq_tcp_port9010_fq_10_50_20_1600182657352.pcap"}.pdf}%
}\hfill
\subfloat[delay]{%
  \includegraphics[width=0.48\columnwidth]{{"figures/plots/rtt_1_competing_flow_fq_tcp_port9010_fq_10_50_20_1600182657352.pcap"}.pdf}%
}
\caption{Flow controlled by the loss-based Cubic on a link with a speed of 50\,Mbps, a delay of 10\,ms and a buffer size of 100 packets. The bottleneck is controlled by fair queuing and is shared with the flow of Figure~\ref{fig:fq_udp}.}
\label{fig:fq_tcp}
\end{figure}
\end{frame}

\begin{frame}{Our Algorithm -- Example}
\begin{figure}[h]
\centering
\subfloat[Throughput (receiving rate)]{%
  \includegraphics[width=0.48\columnwidth]{{"figures/plots/throughput_1_competing_flow_fq_udp_port9000_fq_10_50_20_1600182657352.pcap"}.pdf}%
}\hfill
\subfloat[delay]{%
  \includegraphics[width=0.48\columnwidth]{{"figures/plots/rtt_1_competing_flow_fq_udp_port9000_fq_10_50_20_1600182657352.pcap"}.pdf}%
}
\caption{Flow (started 5\;s later) controlled by our delay-based congestion control on a link with a speed of 50\,Mbps, a delay of 10\,ms and a buffer size of 100 packets. The bottleneck is controlled by fair queuing and is shared with the flow of Figure~\ref{fig:fq_tcp}.}
\label{fig:fq_udp}
\end{figure}
\end{frame}

\begin{frame}{Accuracy of our fair queuing detection}
\begin{itemize}
\item Vary bandwidth from 5 to 50\,Mbps, delay from 10 to 100\,ms and the buffer size from 1 to 100 packets
\item 125 different configurations
\item Each configuration once with fair queuing and once without
\item Accuracy 98\%
\item Same as above but with cross traffic: also 98\% accuracy
\end{itemize}
\end{frame}

\begin{frame}{Systematic evaluation}
\begin{itemize}
\item Vary bandwidth from 5 to 50\,Mbps, delay from 10 to 100\,ms and the buffer size from 1 to 100 packets
\item 27 different configurations
\item Always fair queuing
\item Compare our algorithm with Cubic
\item Cubic: Throughput 18\,Mbit/s; mean delay 109\,ms
\item Our algorithm: Throughput 25\,Mbit/s; mean delay 74\,ms
\end{itemize}
\end{frame}

\section{Discussion}
\begin{frame}{Pacing}
\begin{itemize}
\item During flow startup two concurrent flows (flow 2 having double the rate of flow 1)
\item Our observation: flow 1 has receiving rate of 0!
\item Cause: Pacing: Packets are sent in regular intervals. 
\begin{enumerate}
\item Flow 2 sends a packet that fills the last place in the buffer
\item Flow 1's packet arrives a bit later: Buffer full, packet lost.
\item This happens always since they send at regular intervals with flow 1's packet coincidentally always coming a bit after flow 2's. 
\end{enumerate}
\item Solution: Randomize packet sending times a little bit. 
\end{itemize}
\end{frame}

\begin{frame}{Conclusion}
\begin{itemize}
\item If fair queuing is deployed, we can have significantly smaller delay. 
\item Our proposal seems to work well.
\item Especially useful for long delay-sensitive flows like cloud gaming, video conferencing. 
\end{itemize}
\end{frame}

\makelastslide

\end{document}

### 2020-10-02 paper/paper.tex

--- 2020-10-02 paper/paper.tex
@@
 We use \gls{fq} at the bottleneck link.
-We run each scenario both using Cubic as well as our algorithm.
+We run each scenario both using Cubic as well as our algorithm for 20 seconds each.
 We average the achieved throughputs and average delays from all experiments for both Cubic as well as our algorithm.
-For Cubic, the mean throughput is 18\,Mbit/s and the mean \gls{rtt} is 109\;ms while for our algorithm it is 25\,Mbit/s and 74\,ms.
+For Cubic, the mean throughput is 18\,Mbit/s and the mean \gls{rtt} is 109\,ms while for our algorithm it is 25\,Mbit/s and 74\,ms.
 This shows that our algorithm can correctly determine that there is \gls{fq} and switch to the delay-based \gls{cc} and that our algorithm achieves significantly lower delay.

### 2020-10-05 paper/paper.tex
Change paper and presentation

@@
 
+\def\anonymize{yes}
+\def\tempYes{yes}
+\def\tempNo{no}
+
 \begin{document}
@@
 %
+\ifx\anonymize\tempYes
 \author{Anonymous authors}
@@
 %
-%\author{Maximilian Bachl\inst{1}\orcidID{0000-0001-9482-7428} \and
-%Joachim Fabini\inst{2,3}\orcidID{0000-0002-8285-1591} \and
-%Tanja Zseby\inst{3}\orcidID{0000-0002-5391-467X}}
-%%
-%\authorrunning{M. Bachl et al.}
-%% First names are abbreviated in the running head.
-%% If there are more than two authors, 'et al.' is used.
-%%
-%\institute{TU Wien, Vienna, Austria; 
-%\email{firstname.lastname@tuwien.ac.at}}
+\else
+\author{Maximilian Bachl\inst{1}\orcidID{0000-0001-9482-7428} \and
+Joachim Fabini\inst{2,3}\orcidID{0000-0002-8285-1591} \and
+Tanja Zseby\inst{3}\orcidID{0000-0002-5391-467X}}
 %
+\authorrunning{M. Bachl et al.}
+% First names are abbreviated in the running head.
+% If there are more than two authors, 'et al.' is used.
+%
+\institute{TU Wien, Vienna, Austria; 
+\email{firstname.lastname@tuwien.ac.at}}
+\fi
+
 \maketitle              % typeset the header of the contribution
-%
 
@@
 
-To make our work reproducible and to encourage further research, we make the code, the results and the figures publicly available\footnote{\url{https://github.com/CN-TU/PCC-Uspace}}.
+To make our work reproducible and to encourage further research, we make the code, the results and the figures publicly available\footnote{\ifx\anonymize\tempYes $\langle$anonymized$\rangle$\else\url{https://github.com/CN-TU/PCC-Uspace}\fi}.
 

### 2020-10-05 paper/presentation.tex
Change paper and presentation

@@
 \begin{block}{Insight:}
-When loss-based and delay-based congestion control compete, loss-based one eats delay-based one's lunch!\end{block}
+When loss-based and delay-based congestion control compete, loss-based gets no bandwidth!\end{block}
 \end{frame}
@@
 \item Delay-based congestion control offers same throughput but lower delay than loss-based ones
-\item When competing, loss-based flows will gain all bandwidth. Delay-based flows ``starve''. 
+\item When competing, loss-based flows will gain all bandwidth. Delay-based flows ``starve''
 \item Fair queuing solves this
@@
 \begin{enumerate}
-\item During flow startup, determine if fair queuing is used at the bottleneck or not.
-\item If this is determined, change to a delay-based congestion control algorithm if fair queuing is used and to a loss-based one otherwise.
+\item During flow startup, determine if fair queuing is used at the bottleneck or not
+\item If this is determined, change to a delay-based congestion control algorithm if fair queuing is used and to a loss-based one otherwise
 \end{enumerate}
@@
 \begin{enumerate}
-\item Launch two concurrent flows, flow 1 and flow 2. Initialize flow 1 with some starting sending rate and flow 2 with two times that rate. 
-\item After every delay, if no packet loss occurred, double the sending rate of both flows. 
-\item If packet loss occurred in both flows in the previous delay, calculate the following metric if each flow got an equal share of bandwidth (fair queuing) or not (no fair queuing).
+\item Launch two concurrent flows, flow 1 and flow 2. Initialize flow 1 with some starting sending rate and flow 2 with two times that rate
+\item After every delay, if no packet loss occurred, double the sending rate of both flows
+\item If packet loss occurred in both flows in the previous delay, calculate the following metric if each flow got an equal share of bandwidth (fair queuing) or not (no fair queuing)
 \end{enumerate}

### 2020-10-06 paper/paper.tex
Add load balancing section

@@
 \begin{align}
-\textit{loss ratio} &= \frac{\left(\frac{\textit{receiving rate of flow 1}}{\textit{sending rate of flow 1}}\right)}{\left(\frac{\textit{receiving rate of flow 2}}{\textit{sending rate of flow 2}}\right)} \\
-&= \frac{\textit{receiving rate of flow 1}\times \textit{sending rate of flow 2}}{\textit{sending rate of flow 1} \times \textit{receiving rate of flow 2}}
-\end{align}
+\textit{loss ratio} &= \frac{\left(\frac{\textit{receiving rate of flow 1}}{\textit{sending rate of flow 1}}\right)}{\left(\frac{\textit{receiving rate of flow 2}}{\textit{sending rate of flow 2}}\right)}\end{align}
 
-%JF: FIXME: I'd delete eq (2) above or fit it into one row - no substantial information.
 %JF: Perhaps abbreviate: sendRate flow1, recvRate flow1...
@@
 In this scenario, we show an example of a flow with our mechanism that shares a bottleneck link with a flow using a loss-based congestion control. 
+%JF: More compelling would be the use case when the delay-based flow is started first and the late-comer loss-based preempts it.
 
-%JF: More compelling would be the use case when the delay-based flow is started first and the late-comer loss-based preempts it.
 \subsubsection{No \gls{fq}}
@@
 
+Another scenario that is worth noting is how our proposed solution deals with load balancing: In this paper we assume that both of our flows, which we use during startup to determine the presence of \gls{fq} take the same path to the host on the other side.
+However, with load balancing it could happen that flow 1 takes a different path than flow 2.
+Fortunately, several load balancing algorithms have been proposed whose aim is to make sure that connections, which use multiple paths using Multipath TCP or QUIC, stay on the same path \cite{olivier_bonaventure_multipath_2018,olteanu_datacenter_2016,lienardy_towards_2016}.
+Since we envision our solution to be based on Multipath TCP or QUIC load balancing should not be a problem.
+
 %JF: FIXME: Downside (if room left): it's not clear which element is the bottleneck link.

### 2020-10-07 src/plot_rtt_and_bandwidth.py
create 'plots' directory automatically

@@ -11,6 +11,8 @@ plt.rcParams["font.family"] = "serif"
 plt.rcParams['pdf.fonttype'] = 42
 plt.rcParams["font.size"] = 8
 
+os.makedirs("plots", exist_ok=True)
+
 interval = 1
 pcap_file = sys.argv[1]
 # assert pcap_file.startswith("sender_"), pcap_file

### 2020-10-10 paper/paper.tex
Add reference
Update evaluation section

@@
 Besides the examples we provided, we also conducted a more systematic comparison of the throughput and delay that our algorithm achieves: We use the same scenario as above and again vary bandwidth from 5 to 50\,Mbps, delay from 10 to 100\,ms and the buffer size from 1 to 100 packets.
-For each parameter, we take 3 values, evenly spaced, totaling 27 distinct scenarios.
+For each parameter, we take 5 values, evenly spaced, totaling 125 distinct scenarios.
 We use \gls{fq} at the bottleneck link.
-We run each scenario both using Cubic as well as our algorithm for 20 seconds each.
+We run each scenario both using Cubic as well as our algorithm for 30 seconds each.
 We average the achieved throughputs and average delays from all experiments for both Cubic as well as our algorithm.
-For Cubic, the mean throughput is 18\,Mbit/s and the mean \gls{rtt} is 109\,ms while for our algorithm it is 25\,Mbit/s and 74\,ms.
+For Cubic, the mean throughput is 21\,Mbit/s and the mean \gls{rtt} is 96\,ms while for our algorithm it is 26\,Mbit/s and 64\,ms.
 This shows that our algorithm can correctly determine that there is \gls{fq} and switch to the delay-based \gls{cc} and that our algorithm achieves significantly lower delay.
@@
 Since we envision our solution to be based on Multipath TCP or QUIC load balancing should not be a problem.
+\textit{Apple Music}, for example, successfully load balances multipath connections \cite{olivier_bonaventure_apple_2019}.
 

### 2020-10-10 src/test.py

@@ -299,7 +299,7 @@ elif opt.run_scenario == "evaluation":
 				opt.delay = int(round(delay))
 				opt.buffer_size = int(round(buffer))
 				opt.qdisc = "fq" if fq else "pfifo"
-				opt.time = 20
+				opt.time = 30
 
 				with virtnet.Manager() as context:
 					client_output, timestamp = run(context, "accuracy")

### 2020-10-10 paper/paper.tex
Improve evaluation section

@@
 For Cubic, the mean throughput is 21\,Mbit/s and the mean \gls{rtt} is 96\,ms while for our algorithm it is 26\,Mbit/s and 64\,ms.
-This shows that our algorithm can correctly determine that there is \gls{fq} and switch to the delay-based \gls{cc} and that our algorithm achieves significantly lower delay.
-Surprisingly, our algorithm also achieves higher throughput, which is astonishing considering that loss-based \glspl{cca} such as Cubic should be more aggressive in filling the link.
-When analyzing the results we identified that Cubic performs very poorly in scenarios with a buffer of only 1 packet since it doesn't pace its packets and thus each time when two packets are sent right after another, packet loss occurs and Cubic lowers its rate.
+This shows that our algorithm can correctly determine that there is \gls{fq} and switch to the delay-based \gls{cc} and that our algorithm achieves significantly lower delay (Cubic's delay is 50\% higher) and higher throughput.
+This is a bit surprising, our algorithm also achieves higher throughput, which is astonishing considering that loss-based \glspl{cca} such as Cubic should be more aggressive in filling the link.
+When analyzing the results we identified that Cubic performs very poorly in scenarios with small buffers because the multiplicative decrease of Cubic is 0.7, meaning that the bandwidth is reduced by 30\% after every packet loss while our algorithm only reduces it by only 5\%.
+Thus Cubic reduces bandwidth too much after packet loss which results in underutilization of the link.
 

### 2020-10-12 paper/paper.tex
Add references

@@
 For the evaluation we use the network emulation library \texttt{py-virtnet}\footnote{\url{https://pypi.org/project/py-virtnet/}}.
-For the experiments with \gls{fq} we use the \texttt{fq} module for Linux, for the experiments with a shared queue we use \texttt{pfifo}.
+For the experiments with \gls{fq} we use the \texttt{fq} module for Linux.
+This module keeps one drop tail queue for each flow with a configurable maximum size of packets.
+However, the design of our \gls{fq} detection algorithm does not assume any particular queue management algorithm being used for the queue of each flow and thus is also compatible with more sophisticated \gls{aqm} mechanisms which were explicitly developed to be used in conjunction with \gls{fq} \cite{taht_flow_2018,hoiland-jorgensen_piece_2018,bachl_cocoa_2019,bachl_lfq_2020}.
+For the experiments with a shared queue we use \texttt{pfifo}, which is a simple drop tail queue which can contain up to a fixed number of packets.
 

### 2020-10-20 paper/paper.tex
Add reference

@@
 
-\def\anonymize{yes}
+\def\anonymize{no}
 \def\tempYes{yes}
@@
 
-%JF: Evtl.statt subsubsection Absätze mit Titel fett verwenden um doch einigen Platz zu sparen?
 \subsubsection{\gls{cc} using Active Measurements}
@@
 This is done by performing active measurements by varying the congestion window to evaluate which value gives the highest throughput and lowest packet loss.
+Another approach proposed by \cite{goyal_elasticity_2018} aims to differentiate bulk transfer flows, which they deem ``elastic'' because they try to grab as much bandwidth as available, from other flows that send at a constant rate, which they call ``unelastic''.
+They use this elasticity detection to determine if delay-based \gls{cc} can be used or not.
+If cross traffic is deemed elastic then they use classic loss-based \gls{cc} and otherwise delay-based \gls{cc}.
 
-%JF: Evtl.statt subsubsection Absätze mit Titel fett verwenden um doch einigen Platz zu sparen?
 \subsubsection{Flow startup techniques} Besides the classical TCP slow start algorithm \cite{stevens_tcp_1997} some other proposals were made such as Quick Start \cite{jain_quick-start_2019}, which aims to speed up flow start by routers informing the end point, which rate they deem appropriate.
@@
 For Cubic, the mean throughput is 21\,Mbit/s and the mean \gls{rtt} is 96\,ms while for our algorithm it is 26\,Mbit/s and 64\,ms.
-This shows that our algorithm can correctly determine that there is \gls{fq} and switch to the delay-based \gls{cc} and that our algorithm achieves significantly lower delay (Cubic's delay is 50\% higher) and higher throughput.
-This is a bit surprising, our algorithm also achieves higher throughput, which is astonishing considering that loss-based \glspl{cca} such as Cubic should be more aggressive in filling the link.
+This shows that our algorithm can correctly determine that there is \gls{fq} and switch to the delay-based \gls{cc} and that our algorithm achieves significantly lower delay (loss-based Cubic's delay is 50\% higher) and -- surprisingly -- higher throughput.
+This is astonishing considering that loss-based \glspl{cca} such as Cubic should be more aggressive in filling the link.
 When analyzing the results we identified that Cubic performs very poorly in scenarios with small buffers because the multiplicative decrease of Cubic is 0.7, meaning that the bandwidth is reduced by 30\% after every packet loss while our algorithm only reduces it by only 5\%.

### 2021-02-11 paper/paper_ieee.tex
Add paper in IEEE format

--- 2020-10-10 paper/paper.tex
@@
-% This is samplepaper.tex, a sample chapter demonstrating the
-% LLNCS macro package for Springer Computer Science proceedings;
-% Version 2.20 of 2017/10/04
-%
-\documentclass[runningheads]{llncs}
-%
+%%
+%% This is file `sample-sigconf.tex',
+%% generated with the docstrip utility.
+%%
+%% The original source files were:
+%%
+%% samples.dtx  (with options: `sigconf')
+%%
+%% IMPORTANT NOTICE:
+%%
+%% For the copyright see the source file.
+%%
+%% Any modified versions of this file must be renamed
+%% with new filenames distinct from sample-sigconf.tex.
+%%
+%% For distribution of the original source see the terms
+%% for copying and modification in the file samples.dtx.
+%%
+%% This generated file may be distributed as long as the
+%% original source files, as listed above, are part of the
+%% same distribution. (The sources need not necessarily be
+%% in the same archive or directory.)
+%%
+%% The first command in your LaTeX source must be the \documentclass command.
+\documentclass[conference]{IEEEtran}
+
+\usepackage[utf8]{inputenc}
+\usepackage{multirow}
+\usepackage[inline]{enumitem}
+\usepackage{xcolor}
+\usepackage{booktabs}
+\usepackage[pdfauthor={Maximilian Bachl, Joachim Fabini, Tanja Zseby}]{hyperref}
+\usepackage{amsmath}
+\usepackage{amssymb}
+\usepackage{graphicx}
+\usepackage[numbers]{natbib}
 \usepackage{subfig}
-\usepackage{graphicx}
-\usepackage{cite}
-\usepackage{hyperref}
-\usepackage{booktabs}
-\usepackage[english]{babel}
-\usepackage{xcolor}
+\usepackage{tikz}
+\usepackage{float}
+
 \usepackage[nomain, toc, acronym]{glossaries}
-% Used for displaying a sample figure. If possible, figure files should
-% be included in EPS format.
-%
-% If you use the hyperref package, please uncomment the following line
-% to display URLs in blue roman font according to Springer's eBook style:
-\renewcommand\UrlFont{\color{blue}\rmfamily}
+\glsdisablehyper
+
+\newcommand\note[2]{{\color{#1}#2}}
+\newcommand\todo[1]{{\note{red}{TODO: #1}}}
+
+\def\anonymize{no}
+\def\tempYes{yes}
+\def\tempNo{no}
+
+%%
+%% end of the preamble, start of the body of the document source.
+\begin{document}
+
+%%
+%% The "title" command has an optional parameter,
+%% allowing the author to define a "short title" to be used in page headers.
+\title{Detecting Fair Queuing for Better Congestion Control}
+
+\author{\IEEEauthorblockN{Maximilian Bachl, Joachim Fabini, Tanja Zseby}
+\IEEEauthorblockA{Technische Universität Wien\\
+firstname.lastname@tuwien.ac.at}}
+
+%\date{\today}
+
+% \IEEEoverridecommandlockouts
+% \IEEEpubid{\begin{minipage}[t]{\textwidth}\ \\[10pt]
+%         \centering\normalsize{xxx-x-xxxx-xxxx-x/xx/\$31.00 \copyright 2018 IEEE}
+% \end{minipage}}
+
+% \renewcommand*{\bibfont}{\footnotesize}
+
+%\newcommand\copyrighttext{%
+%  \footnotesize \textcopyright 2020 IEEE. Personal use of this material is permitted.
+%  Permission from IEEE must be obtained for all other uses, in any current or future
+%  media, including reprinting/republishing this material for advertising or promotional
+%  purposes, creating new collective works, for resale or redistribution to servers or
+%  lists, or reuse of any copyrighted component of this work in other works.}
+%\newcommand\copyrightnotice{%
+%\begin{tikzpicture}[remember picture,overlay]
+%\node[anchor=south,yshift=10pt] at (current page.south) {\fbox{\parbox{\dimexpr\textwidth-\fboxsep-\fboxrule\relax}{\copyrighttext}}};
+%\end{tikzpicture}%
+%}
+
+\maketitle%
+%\copyrightnotice
+
+ \thispagestyle{plain}
+ \pagestyle{plain}
 
@@
 \newacronym{qoe}{QoE}{Quality of Experience}
-
-\usepackage{amsmath,amssymb,amsfonts}
-\newcommand{\mynote}[3]{
-    \fbox{\bfseries\sffamily\scriptsize#1}
-    {\small$\blacktriangleright$\textsf{\emph{\color{#3}{#2}}}$\blacktriangleleft$}}
-
-\newcommand{\todo}[1]{\mynote{TODO}{#1}{red}}
-\newcommand{\noteMax}[1]{\mynote{Max}{#1}{green}}
-
-\def\anonymize{yes}
-\def\tempYes{yes}
-\def\tempNo{no}
-
-\begin{document}
-%
-%JF: \title{Detecting Fair Queuing to Enable Better Congestion Control}
-%JF: Alternative Vorschläge für Titel, etwas flüssiger:
-\title{Optimizing Congestion Control Through Fair Queuing Detection}
-%JF: \title{Optimum Congestion Control Selection Through Fair Queuing Detection}
-
-%\titlerunning{Abbreviated paper title}
-% If the paper title is too long for the running head, you can set
-% an abbreviated paper title here
-%
-\ifx\anonymize\tempYes
-\author{Anonymous authors}
-%
-\authorrunning{Anonymous authors}
-% First names are abbreviated in the running head.
-% If there are more than two authors, 'et al.' is used.
-%
-\else
-\author{Maximilian Bachl\inst{1}\orcidID{0000-0001-9482-7428} \and
-Joachim Fabini\inst{2,3}\orcidID{0000-0002-8285-1591} \and
-Tanja Zseby\inst{3}\orcidID{0000-0002-5391-467X}}
-%
-\authorrunning{M. Bachl et al.}
-% First names are abbreviated in the running head.
-% If there are more than two authors, 'et al.' is used.
-%
-\institute{TU Wien, Vienna, Austria; 
-\email{firstname.lastname@tuwien.ac.at}}
-\fi
-
-\maketitle              % typeset the header of the contribution
-
-%JF: FIXME: A fundamental problem that we should perhaps address in a brief statement: The proposed method does not work for cases in which load balancers are on the path.
-Distinct flows are likely to result in distinct hashes/flowIDs, meaning that the balancer will potentially assign them to distinct paths.
-So we may add a restrictive statement to the discussion that we assume that all flows share the same bottleneck?
-In addition we may state that no service/port priorization is in place.
-%JF Any other idea?
 
@@
 Results show that our algorithm reliably detects fair queuing and can achieve low delay and high throughput in case fair queuing is detected.
+\end{abstract}
 
-\keywords{Congestion Control  \and Active Measurements \and Fair Queuing \and Queue Management.}
-\end{abstract}
-%
-%
-%
+%%
+%% The abstract is a short summary of the work to be presented in the
+%% article.
+%\begin{abstract}
+%  A clear and well-documented \LaTeX\ document is presented as an
+%  article formatted for publication by ACM in a conference proceedings
+%  or journal publication. Based on the ``acmart'' document class, this
+%  article presents and explains many of the common variations, as well
+%  as many of the formatting elements an author may use in the
+%  preparation of the documentation of their work.
+%\end{abstract}
+
+\maketitle
+
+\glsresetall
+\newacronym{cc}{CC}{Congestion Control}
+\newacronym{cca}{CCA}{Congestion Control Algorithm}
+\newacronym{aqm}{AQM}{Active Queue Management}
+\newacronym{rtt}{RTT}{Round Trip Time}
+\newacronym{fq}{FQ}{Fair Queuing}
+\newacronym{bdp}{BDP}{Bandwidth Delay Product}
+\newacronym{qoe}{QoE}{Quality of Experience}
+
+\def\anonymize{no}
+\def\tempYes{yes}
+\def\tempNo{no}
+
 \section{Introduction}
-\label{sec:introduction}
+\label{sec:introduction_fq}
 
@@
 
-%JF: Evtl.statt subsubsection Absätze mit Titel fett verwenden um doch einigen Platz zu sparen?
 \subsubsection{\gls{cc} using Active Measurements}
 Several approaches have been recently proposed which aim to incorporate active measurements into \glspl{cca}.
-\cite{dong_pcc_2015,dong_pcc_2018} introduced the \gls{cca} \textit{PCC}, which uses active measurements to find the sending rate which optimizes a given reward function.
-\cite{cardwell_bbr:_2016} introduce the \textit{BBR} \gls{cca}, which uses active measurement to create a model of the network connection.
+\citep{dong_pcc_2015,dong_pcc_2018} introduced the \gls{cca} \textit{PCC}, which uses active measurements to find the sending rate which optimizes a given reward function.
+\citep{cardwell_bbr:_2016} introduce the \textit{BBR} \gls{cca}, which uses active measurement to create a model of the network connection.
 This model is then used to determine the optimal sending rate.
-\cite{bachl_rax_2019} use Reinforcement Learning to learn \gls{cc} policies.
-This is done by performing active measurements by varying the congestion window to evaluate which value gives the highest throughput and lowest packet loss.
+An approach proposed by \citep{goyal_elasticity_2018} aims to differentiate bulk transfer flows, which they deem ``elastic'' because they try to grab as much bandwidth as available, from other flows that send at a constant rate, which they call ``unelastic''.
+They use this elasticity detection to determine if delay-based \gls{cc} can be used or not.
+If cross traffic is deemed elastic then they use classic loss-based \gls{cc} and otherwise delay-based \gls{cc}.
 
-%JF: Evtl.statt subsubsection Absätze mit Titel fett verwenden um doch einigen Platz zu sparen?
-\subsubsection{Flow startup techniques} Besides the classical TCP slow start algorithm \cite{stevens_tcp_1997} some other proposals were made such as Quick Start \cite{jain_quick-start_2019}, which aims to speed up flow start by routers informing the end point, which rate they deem appropriate.
-\cite{jarvinen_congestion_2019} investigate the impact of \gls{aqm} and \gls{cc} during flow startup.
+\subsubsection{Flow startup techniques} Besides the classical TCP slow start algorithm \citep{stevens_tcp_1997} some other proposals were made such as Quick Start \citep{jain_quick-start_2019}, which aims to speed up flow start by routers informing the end point, which rate they deem appropriate.
+\citep{jarvinen_congestion_2019} investigate the impact of \gls{aqm} and \gls{cc} during flow startup.
 They inspect a variety of different \gls{aqm} algorithms and also investigate \gls{fq}.
-\cite{kuhlewind_chirping_2010} propose ``chirping'' for flow startup to estimate the available bandwidth.
+\citep{kuhlewind_chirping_2010} propose ``chirping'' for flow startup to estimate the available bandwidth.
 This works by sending packets with different gaps between them to verify what the actual link speed is.
-\cite{mittal_recursively_2014} propose speeding up flow startup by using packets of different priority.
+\citep{mittal_recursively_2014} propose speeding up flow startup by using packets of different priority.
 The link is flooded with low-priority packets and in case the link is not fully utilized during startup, these low-priority packets are going to be transmitted, allowing the starting flow to achieve high throughput.
@@
 \subsubsection{Detecting \gls{aqm} mechanisms} There are few publications that investigate methods to detect the presence of an \gls{aqm}.
-\cite{kargar_bideh_tada_2016} propose a method to detect and distinguish certain popular \gls{aqm} mechanisms.
-However, unlike our work, it doesn't consider \gls{fq}.
-\cite{baykal_detection_2017} propose a machine learning based approach to fingerprint \gls{aqm} algorithm.
+\citep{kargar_bideh_tada_2016} propose a method to detect and distinguish certain popular \gls{aqm} mechanisms.
+However, unlike my work, it doesn't consider \gls{fq}.
+\citep{baykal_detection_2017} propose a machine learning based approach to fingerprint \gls{aqm} algorithm.
 \gls{fq} is not considered, too.
@@
 \section{Concept}
-\label{sec:concept}
+\label{sec:concept_fq}
 
@@
 
-%\begin{figure}
-%\centering
-%\includegraphics[width=1\columnwidth]{{"figures/fq_illustration_throughput"}.pdf}
-%\caption{Sending rate}
-%\label{fig:throughput}
-%\end{figure}
-%
-%\begin{figure}
-%\includegraphics[width=1\columnwidth]{{"figures/fq_illustration_goodput_no_fq"}.pdf}
-%\caption{Receiving rate (goodput) in case the bottleneck link doesn't use \gls{fq}. Flow 2, which sends more also manages to get more data through the bottleneck.}
-%\label{fig:goodput_no_fq}
-%\end{figure}
-%
-%\begin{figure}
-%\includegraphics[width=1\columnwidth]{{"figures/fq_illustration_goodput_fq"}.pdf}
-%\caption{Receiving rate (goodput) in case the bottleneck link uses \gls{fq}. Flow 2, which sends more gets as much data through the bottleneck as flow 1.}
-%\label{fig:goodput_fq}
-%\end{figure}
-
 \begin{figure}
@@
 \label{fig:goodput_fq}}
-\caption{An example illustrating our proposed flow startup mechanism.
+\caption{An example illustrating my proposed flow startup mechanism.
 Figure \ref{fig:throughput} shows the sending rate, \ref{fig:goodput_no_fq} the receiving rate in case there's no \gls{fq} and \ref{fig:goodput_fq} the receiving rate if there is \gls{fq}.}
@@
 \begin{enumerate}
-\item Launch two concurrent flows, flow 1 and flow 2. Initialize flow 1 with some starting sending rate and flow 2 with two times that rate. 
+\item Launch two concurrent flows, flow 1 and flow 2 from the same source host to the same destination host. Initialize flow 1 with some starting sending rate and flow 2 with two times that rate. 
 \item After every \gls{rtt}, if no packet loss occurred, double the sending rate of both flows. 
@@
 
-%JF: Perhaps abbreviate: sendRate flow1, recvRate flow1...
 This loss ratio metric indicates if flow 2, which has a higher sending rate, achieves a higher receiving rate (goodput) than flow 1. If this is the case, there's no \gls{fq}, otherwise there is.
@@
 It is necessary to wait for packet loss in both connections because only then it is certain that the link is saturated. Our algorithm assumes that the link is saturated. 
-\item If \gls{fq} is used, the $\textit{loss ratio}$ is 2, otherwise it is 1. We choose 1.5 as the cutoff value. 
+\item If \gls{fq} is used, the $\textit{loss ratio}$ is 2, otherwise it is 1. I choose 1.5 as the cutoff value. 
 \item Since the presence or absence of \gls{fq} is now known the appropriate \gls{cca} can be launched. 
@@
 Figure~\ref{fig:throughput}, \ref{fig:goodput_no_fq} and \ref{fig:goodput_fq} schematically illustrate this mechanism.
-We assume a sender being connected to a receiver, with a bottleneck link between them.
+I assume a sender being connected to a receiver, with a bottleneck link between them.
+I also assume that all flows between the source and the destination follow the same route.
 On this bottleneck link, the queue is either managed by \gls{fq} or by a shared queue (no \gls{fq}).
-Figure~\ref{fig:throughput} shows how the sending rate (observed before the bottleneck link) is doubled after each \gls{rtt}.
-Furthermore, it shows that flow 2 has double the sending rate of flow 1.
-%\noteMax{Figures take up too much space.
-Is it better to put them next to each other?
-The actual problem is the nonsensical Springer format.}.
+
+Figure~\ref{fig:throughput} shows how the sending rate (observed before the bottleneck link) is doubled after each \gls{rtt}. Furthermore, it shows that flow 2 has double the sending rate of flow 1.
 
@@
 \begin{enumerate}
-\item If $\textit{current rtt} \leq \textit{smallest rtt ever measured on the connection} + 5\textit{ms}$ then increase the sending rate by 1\%.
-\item Otherwise ($\textit{current rtt} > \textit{smallest rtt ever measured on the connection} + 5\textit{ms}$) decrease the sending rate by 5\%.
+\item If \textit{current rtt} $\leq$ \textit{smallest rtt ever measured on the connection} $+$ 5\textit{ms} then increase the sending rate by 1\%.
+I choose 5\,ms since a delay of 5\,ms is barely noticeable by humans.
+This design choice was first established in \citep{nichols_controlling_2012}.
+\item Otherwise (\textit{current rtt} > \textit{smallest rtt ever measured on the connection} + 5\textit{ms}) decrease the sending rate by 5\%.
 \end{enumerate}
 
-This \gls{cca} is meant as a simple proof-of-concept to demonstrate the efficacy of our \gls{fq} detection mechanism but we do not claim it to be optimal in any way.
+%This \gls{cca} is meant as a simple proof-of-concept to demonstrate the efficacy of my \gls{fq} detection mechanism but I do not claim it to be optimal in any way.
 
@@
 If the absence of \gls{fq} is detected, a loss based \gls{cc} is used as a fallback.
-We use the algorithm \textit{PCC} proposed by \cite{dong_pcc_2015}.
+I use the algorithm \textit{PCC} proposed by \citep{dong_pcc_2015}.
 In simplified terms, this \gls{cca} aims to maximize a utility function, which strives for high throughput and low packet loss.
-However, it is reasonably aggressive and does not starve when competing with other loss-based \glspl{cca} such as Cubic \cite{ha_cubic_2008}.
+However, it is reasonably aggressive and does not starve when competing with other loss-based \glspl{cca} such as Cubic \citep{ha_cubic_2008}.
 
@@
 
-We base our implementation on the code of \textit{PCC Vivace} \cite{dong_pcc_2018}, which in turn is an improved version of PCC.
-The PCC code is based on the \textit{UDT} library \cite{gu_udt_2007}, which is a library that provides a reliable transport protocol similar to TCP on top of UDP.
-This approach is similar to QUIC \cite{iyengar_quic_2018}.
-An advantage of implementing our approach on top of UDP as opposed to TCP is that it is easier to develop and modify since no kernel module needs to be created and that it is more secure since the code is isolated in a user space process.
+I base my implementation on the code of \textit{PCC Vivace} \citep{dong_pcc_2018}, which in turn is an improved version of PCC.
+The PCC code is based on the \textit{UDT} library \citep{gu_udt_2007}, which is a library that provides a reliable transport protocol similar to TCP on top of UDP.
+This approach is similar to QUIC \citep{iyengar_quic_2018}.
+An advantage of implementing my approach on top of UDP as opposed to TCP is that it is easier to develop and modify since no kernel module needs to be created and that it is more secure since the code is isolated in a user space process.
 
@@
 This may be a challenge for deployments since solutions that require cooperation of several instances are more difficult to deploy at a wide scale in the Internet.
-%This can be seen for example when looking at the deployment of IPv6, which requires the participation of every router on the way and every end device.
-%JF: FIXME: I'm sceptical that the examples below work well in all scenarios. Distinct flows end up on distinct paths, so they do not necessarily use the same bottleneck.
-However, multipath TCP \cite{handley_tcp_2012} and QUIC, which enable recombining data of several flows, are becoming increasingly prevalent and Google's QUIC implementation has multipath support \cite{google_quiche_2020}.
-We thus argue that the actual obstacles to deployment of our solution are constantly decreasing since QUIC (mostly Google traffic) already accounts for around %$\frac{1}{10}$
-$10\%$ of all Internet traffic \cite{ruth_first_2018} and is growing. Moreover, Apple's macOS and iOS support multipath TCP \cite{apple_about_2018} support.
+However, multipath TCP \citep{handley_tcp_2012} and QUIC, for which there are proposals to enable recombining data of several flows \citep{de_coninck_multipath_2017}, are becoming increasingly prevalent.
+I thus argue that the actual obstacles to deployment of my solution are constantly decreasing since QUIC (mostly Google traffic) already accounts for around
+$10\%$ of all Internet traffic \citep{ruth_first_2018} and is growing. Moreover, Apple's macOS and iOS support multipath TCP \citep{apple_about_2018} support.
+
+\section{Experiment setup}
+
+For the evaluation I use the network emulation library \texttt{py-virtnet}\footnote{\url{https://pypi.org/project/py-virtnet/}}.
+All experiments run on virtual hosts on one physical host, on which the necessary network components and virtual hosts are emulated using Linux network namespaces.
+The setup consists of two virtual hosts connected through a virtual switch.
+One virtual host acts as a server and one as a client and a bulk transfer is initiated from the client to the server.
+
+For the experiments with \gls{fq} I use the \texttt{fq} module for Linux.
+This module keeps one drop tail queue for each flow with a configurable maximum size of packets.
+However, the design of my \gls{fq} detection algorithm does not assume any particular queue management algorithm being used for the queue of each flow and thus is also compatible with more sophisticated \gls{aqm} mechanisms which were explicitly developed to be used in conjunction with \gls{fq} \citep{taht_flow_2018,hoiland-jorgensen_piece_2018,bachl_cocoa_2019,bachl_lfq_2020}.
+It only assumes that each flow gets the same share of bandwidth and that for a flow, whose queue is growing faster because more packets of this flow arrive at the bottleneck, packets are dropped faster than for a flow of which fewer packets appear at the bottleneck link.
+For the experiments with a shared queue I use \texttt{pfifo}, which is a simple drop tail queue which can contain up to a fixed number of packets.
 
 \section{Results}
-
-For the evaluation we use the network emulation library \texttt{py-virtnet}\footnote{\url{https://pypi.org/project/py-virtnet/}}.
-For the experiments with \gls{fq} we use the \texttt{fq} module for Linux, for the experiments with a shared queue we use \texttt{pfifo}.
 
@@
 }
-\caption{Throughput and delay of a flow controlled by our delay-based \gls{cca} on a link with a speed of 10\,Mbps, a delay of 10\,ms and a buffer size of 100 packets.}
+\caption{Throughput and delay of a flow controlled by my delay-based \gls{cca} on a link with a speed of 10\,Mbps, a delay of 10\,ms and a buffer size of 100 packets.}
 \label{fig:just_one_flow_udp}
@@
 
-First, we demonstrate empirically that delay-based \gls{cc} actually has a benefit over loss-based one.
-On one hand, \autoref{fig:just_one_flow_tcp} shows that the loss-based \gls{cc} achieves a throughput of almost 10\,Mbps on a 10\,Mbps link, which means full utilization and an \gls{rtt} that is  higher than 100\,ms.
-On the other hand, \autoref{fig:just_one_flow_udp} shows an example flow using our simple delay-based \gls{cca} and a slow start algorithm similar to the one used by the classic TCP \gls{cca} \textit{New Reno} \cite{allman_tcp_1999}.
+First, I demonstrate empirically that delay-based \gls{cc} actually has a benefit over loss-based one.
+On one side, \autoref{fig:just_one_flow_tcp} shows that the loss-based \gls{cc} achieves a throughput of almost 10\,Mbps on a 10\,Mbps link, which means full utilization and an \gls{rtt} that is  higher than 100\,ms, which is not desirable.
+On the other side, \autoref{fig:just_one_flow_udp} shows an example flow using my simple delay-based \gls{cca} and a slow start algorithm similar to the one used by the classic TCP \gls{cca} \textit{New Reno} \citep{allman_tcp_1999}.
 While the delay-based flow also achieves very high utilization, delay rarely exceeds 20\,ms, meaning that it is more than 5 times lower than the one of the loss-based \gls{cca} of \autoref{fig:just_one_flow_tcp}.
+The initial peak in \autoref{fig:just_one_flow_udp} is caused by the exponential increase of the sending rate in the beginning, which is commonly used by transport protocols and called ``slow start'' \citep{stevens_tcp_1997}.
+Our delay-based \gls{cc} also uses a mechanism similar to slow start in the beginning.
 
 \subsection{Comparing flows with different Congestion Controls and AQM mechanisms}
-In this scenario, we show an example of a flow with our mechanism that shares a bottleneck link with a flow using a loss-based congestion control. 
-%JF: More compelling would be the use case when the delay-based flow is started first and the late-comer loss-based preempts it.
+In this scenario, I show an example of a flow with my mechanism that shares a bottleneck link with a flow using a loss-based congestion control.
 
 \subsubsection{No \gls{fq}}
-The examples in \autoref{fig:pfifo_tcp} and in \autoref{fig:pfifo_udp} show the loss-based \gls{cca} Cubic compete against our delay-based \gls{cca} on a link without \gls{fq}.
-It is important to note that we do not use our \gls{fq} detection mechanism here but instead choose delay-based \gls{cc} from the beginning to demonstrate the starvation of delay-based flows that happens when they compete against loss-based ones without \gls{fq}.
-In the scenario, the loss-based flow starts 5 seconds earlier and after 5 seconds, the delay-based flow joins.
-While the delay-based flow gains some share of the link during startup, it is pushed away by the loss-based flow later on and ``starves''.
-We start the loss-based flow before to make sure that it has sufficient time to fill up the link.
+The examples in \autoref{fig:pfifo_tcp} and in \autoref{fig:pfifo_udp} show the loss-based \gls{cca} Cubic compete against my delay-based \gls{cca} on a link without \gls{fq}.
+It is important to note that I do not use my \gls{fq} detection mechanism here but instead choose delay-based \gls{cc} from the beginning to demonstrate the starvation of delay-based flows that happens when they compete against loss-based ones without \gls{fq}.
+I start the loss-based flow before to make sure that it has sufficient time to fill up the link: The loss-based flow starts 5 seconds earlier and after 5 seconds, the delay-based flow joins.
+I choose these 5 seconds so that the first flow can enter its stable state because I want to show how how the flow, even though it had the advantage of the 5 seconds in which it could use the link all for itself, loses to the loss-based \gls{cc}: While the delay-based flow gains some share of the link during startup, it is pushed away by the loss-based flow later on and ``starves''.
 
@@
 }
-\caption{Throughput and delay of a flow controlled by our delay-based \gls{cca} on a link with a speed of 50\,Mbps, a delay of 10\,ms and a buffer size of 100 packets.
+\caption{Throughput and delay of a flow controlled by my delay-based \gls{cca} on a link with a speed of 50\,Mbps, a delay of 10\,ms and a buffer size of 100 packets.
 The bottleneck is controlled by a shared queue (no \gls{fq}) and is shared with the flow of \autoref{fig:pfifo_tcp}.}
@@
 }
-\caption{Throughput and delay of a flow controlled by our delay-based \gls{cca} on a link with a speed of 50\,Mbps, a delay of 10\,ms and a buffer size of 100 packets.
+\caption{Throughput and delay of a flow controlled by my delay-based \gls{cca} on a link with a speed of 50\,Mbps, a delay of 10\,ms and a buffer size of 100 packets.
 The bottleneck is controlled by \gls{fq} and is shared with the flow of \autoref{fig:fq_tcp}.}
@@
 
-For a link with fair queuing, the following example results show that throughput is identical for our flow (\autoref{fig:fq_udp}) and the Cubic flow (\autoref{fig:fq_tcp}), just that the Cubic flow's delay is more than two times higher than the delay-based one's.
-Thus, our mechanism is effective to detect if the bottleneck uses fair queuing and use a delay-based congestion control if it is the case and a loss-based one otherwise.
+For a link with fair queuing, the following example results show that throughput is identical for my flow (\autoref{fig:fq_udp}) and the Cubic flow (\autoref{fig:fq_tcp}), just that the Cubic flow's delay is more than two times higher than the delay-based one's.
+Thus, my mechanism detects that the bottleneck uses \gls{fq} and uses a delay-based congestion control because it is better suited in a scenario with \gls{fq}.
 
@@
 
-To evaluate if our proposed mechanism correctly recognizes \gls{fq} in a systematic manner, we perform experiments using a wide range of network configurations: We vary bandwidth from 5 to 50\,Mbps, delay from 10 to 100\,ms and the buffer size from 1 to 100 packets, using a grid of these parameters.
-For each parameter, we take 5 values, evenly spaced.
-Thus, we run experiments with 125 different configurations.
-We run each configuration once with \gls{fq} and once without.
-Results show that for experiments without \gls{fq} (using a shared queue), in over 99\% of cases, the absence of \gls{fq} was correctly detected and in less than 1\% of cases, our algorithm detected \gls{fq} even though there was no \gls{fq}.
+To evaluate if my proposed mechanism correctly recognizes \gls{fq} in a systematic manner, I perform experiments using a wide range of network configurations: I vary bandwidth from 5 to 50\,Mbps, delay from 10 to 100\,ms and the buffer size from 1 to 100 packets, using a grid of these parameters.
+For each parameter, I take 5 values, evenly spaced.
+Thus, I run experiments with 125 different configurations.
+I run each configuration once with \gls{fq} and once without.
+Results show that for experiments without \gls{fq} (using a shared queue), in over 99\% of cases, the absence of \gls{fq} was correctly detected and in less than 1\% of cases, my algorithm detected \gls{fq} even though there was no \gls{fq}.
 For experiments, during which \gls{fq} was actually deployed on the bottleneck link, this was correctly detected in 97\% of cases and wrongly in 3\% of the time.
-The overall accuracy is thus 98\%, which we consider sufficient.
+The overall accuracy is thus 98\%, which I consider sufficient.
 
-To verify the behavior of our algorithm under the presence of cross traffic, we repeated the above experiments but with one bulk traffic flow using Cubic running at the same time, which we start 5 seconds before so that it has sufficient time to gain full bandwidth.
-Also under these circumstances we also achieve 98\% detection accuracy using our algorithm.
+To verify the behavior of my algorithm under the presence of cross traffic, I repeated the above experiments but with one bulk traffic flow using Cubic running at the same time, which I start 5 seconds before so that it has sufficient time to gain full bandwidth.
+Also under these circumstances I also achieve 98\% detection accuracy using my algorithm.
+I investigated the reason for the detection accuracy not being 100\% and found the following: The PCC code relies on packet scheduling using timers, which is very computationally expensive because it needs a timer for every packet and thus the program has to be put to sleep/woken up repeatedly by the operating system.
+Thus, if processor load is high, the timers fail and PCC can't send as many packets as it would want to.
+My algorithm relies on flow 1 sending half the rate of flow 2 but this cannot always be guaranteed if CPU load is high.
+I tried to minimize all tasks which run in the background and achieved the accuracy I got.
+When I ran the experiments while I was browsing the web, I got lower accuracy.
+Solutions to the problem are to not schedule packets using timers or to implement the algorithm not in user space but in kernel space.
+Then the computational cost of could decrease because fewer context switches between the PCC user space application and the kernel are necessary.
 
@@
 
-Besides the examples we provided, we also conducted a more systematic comparison of the throughput and delay that our algorithm achieves: We use the same scenario as above and again vary bandwidth from 5 to 50\,Mbps, delay from 10 to 100\,ms and the buffer size from 1 to 100 packets.
-For each parameter, we take 5 values, evenly spaced, totaling 125 distinct scenarios.
-We use \gls{fq} at the bottleneck link.
-We run each scenario both using Cubic as well as our algorithm for 30 seconds each.
-We average the achieved throughputs and average delays from all experiments for both Cubic as well as our algorithm.
-For Cubic, the mean throughput is 21\,Mbit/s and the mean \gls{rtt} is 96\,ms while for our algorithm it is 26\,Mbit/s and 64\,ms.
-This shows that our algorithm can correctly determine that there is \gls{fq} and switch to the delay-based \gls{cc} and that our algorithm achieves significantly lower delay (Cubic's delay is 50\% higher) and higher throughput.
-This is a bit surprising, our algorithm also achieves higher throughput, which is astonishing considering that loss-based \glspl{cca} such as Cubic should be more aggressive in filling the link.
-When analyzing the results we identified that Cubic performs very poorly in scenarios with small buffers because the multiplicative decrease of Cubic is 0.7, meaning that the bandwidth is reduced by 30\% after every packet loss while our algorithm only reduces it by only 5\%.
+Besides the examples I provided, I also conducted a more systematic comparison of the throughput and delay that my algorithm achieves: I use the same scenario as above and again vary bandwidth from 5 to 50\,Mbps, delay from 10 to 100\,ms and the buffer size from 1 to 100 packets.
+For each parameter, I take 5 values, evenly spaced, totaling 125 distinct scenarios.
+I use \gls{fq} at the bottleneck link.
+I run each scenario both using Cubic as well as my algorithm for 30 seconds each.
+I average the achieved throughputs and average delays from all experiments for both Cubic as well as my algorithm.
+For Cubic, the mean throughput is 21\,Mbit/s and the mean \gls{rtt} is 96\,ms while for my algorithm it is 26\,Mbit/s and 64\,ms.
+This shows that my algorithm can correctly determine that there is \gls{fq} and switch to the delay-based \gls{cc} and that my algorithm achieves significantly lower delay (loss-based Cubic's delay is 50\% higher) and -- surprisingly -- higher throughput.
+This is astonishing considering that loss-based \glspl{cca} such as Cubic should be more aggressive in filling the link.
+When analyzing the results I identified that Cubic performs very poorly in scenarios with small buffers because the multiplicative decrease of Cubic is 0.7, meaning that the bandwidth is reduced by 30\% after every packet loss while my algorithm only reduces it by only 5\%.
 Thus Cubic reduces bandwidth too much after packet loss which results in underutilization of the link.
-
-%\begin{table}[t]
-%\caption{Comparison of for loss-based \glspl{cca} on a link of 100\;Mbit/s with an \gls{rtt} of 50\;ms.
-Utilization in \% of maximum capacity; \gls{rtt} in milliseconds.
-For Cubic fq achieves good performance.
-This is because its default buffer size of 100 packets coincidentially works well for this specific link speed and \gls{rtt}.
-We take the mean of 10 runs for each scenario.} \label{tab:performance_results}
-%\centering
-%\begin{tabular}{l r r r r r r} \toprule
-%& \multicolumn{2}{l}{fq} & \multicolumn{2}{l}{fq\_codel} & \multicolumn{2}{l}{eggs} \\
-%\cmidrule(r){2-3} \cmidrule(lr){4-5} \cmidrule(l){6-7}
-%& Util. & \gls{rtt} & Util. & \gls{rtt} & Util. & \gls{rtt} \\ \midrule
-%Cubic	& 96.2 & 54 & 92.6 & 52 & 98.5 & 66.5 \\
-%Reno	& 84.1 & 52.7 & 81.1 & 51.5 & 97.9 & 98.2 \\
-%\bottomrule
-%\end{tabular}
-%\end{table}
 
@@
 
-During the implementation of our algorithm we noticed peculiar behavior when not using \gls{fq}: During startup, we make flow 1 have half the sending rate of flow 2.
-We would expect the receiving rates to be the same: flow 1 achieving around half the rate of flow 2.
-However, we noticed that usually flow 1 would have a receiving rate of 0.
-After investigating the issue we noticed that the reason was pacing: PCC (and other \glspl{cca}) such as BBR send packets in regular intervals, for example every millisecond, because sending regularly keeps the queues shorter.
+During the implementation of my algorithm I noticed peculiar behavior when not using \gls{fq}: During startup, I make flow 1 have half the sending rate of flow 2.
+I would expect the receiving rates to be the same: flow 1 achieving around half the rate of flow 2.
+However, I noticed that usually flow 1 would have a receiving rate of 0.
+After investigating the issue I noticed that the reason was pacing: PCC (and other \glspl{cca}) such as BBR send packets in regular intervals, for example every millisecond, because sending regularly keeps the queues shorter.
 This behavior is called pacing.
@@
 Then, when the packets arrive at the bottleneck, flow 2's packet, which arrives a bit earlier, fills up the buffer and flow 1's packet is consequently always dropped.
-As a solution, we do not send packets at regular intervals during startup but always add a small random value: For example, if the sending interval were 1\,ms, we would send the packet randomly in the interval from 0.5 to 1.5\,ms.
+As a solution, I do not send packets at regular intervals during startup but always add a small random value: For example, if the sending interval were 1\,ms, I would send the packet randomly in the interval from 0.5 to 1.5\,ms.
 This solves the aforementioned problem.
-The insight we gained from this is that while pacing is generally supportive, it is actually harmful if several flows compete at a bottleneck without \gls{fq} and one flow's sending rate is a multiple of the other's.
+The insight I gained from this is that while pacing is generally supportive, it is actually harmful if several flows compete at a bottleneck without \gls{fq} and one flow's sending rate is a multiple of the other's.
 
-One problem of our proposed solution occurs when the bottleneck link changes while a flow is running.
-For example, the bottleneck link in the beginning could be the home router, while later is it becoming another router deeper in the core network.
-The problem occurs if the home router supports \gls{fq} and our algorithm detects this while the router in the core network, which becomes the bottleneck later on, doesn't support \gls{fq}.
-Then, our algorithm is going to use a delay-based \gls{cc} because that's what it determined to be correct in the beginning of the flow.
-However, later when the bottleneck changes, our algorithm is still using delay-based \gls{cc} even though the current bottleneck doesn't support \gls{fq} anymore.
-As a result, it might be possible that our algorithm uses the wrong \gls{cc}.
+One problem of my proposed solution occurs when the bottleneck link changes while a flow is running.
+For example, the bottleneck link could be the home router in the beginning, while later the bottleneck link could be another router deeper in the core network.
+The problem occurs if the home router supports \gls{fq} and my algorithm detects this while the router in the core network, which becomes the bottleneck later on, doesn't support \gls{fq}.
+Then, my algorithm is going to use a delay-based \gls{cc} because that's what it determined to be correct in the beginning of the flow.
+However, later when the bottleneck changes, my algorithm is still using delay-based \gls{cc} even though the current bottleneck doesn't support \gls{fq} anymore.
+As a result, it might be possible that my algorithm uses the wrong \gls{cc}.
 One way to prevent this problem from occurring is to use periodic measurements.
-For example, our algorithm to determine \gls{fq} could be used every 5 or 10 seconds to measure again if the current bottleneck supports \gls{fq}.
+For example, my algorithm to determine \gls{fq} could be used every 5 or 10 seconds to measure again if the current bottleneck supports \gls{fq}.
 
-Another scenario that is worth noting is how our proposed solution deals with load balancing: In this paper we assume that both of our flows, which we use during startup to determine the presence of \gls{fq} take the same path to the host on the other side.
+Another scenario that is worth noting is how my proposed solution deals with load balancing: In this research I assume that both of my flows, which I use during startup to determine the presence of \gls{fq} take the same path to the host on the other side.
 However, with load balancing it could happen that flow 1 takes a different path than flow 2.
-Fortunately, several load balancing algorithms have been proposed whose aim is to make sure that connections, which use multiple paths using Multipath TCP or QUIC, stay on the same path \cite{olivier_bonaventure_multipath_2018,olteanu_datacenter_2016,lienardy_towards_2016}.
-Since we envision our solution to be based on Multipath TCP or QUIC load balancing should not be a problem.
-\textit{Apple Music}, for example, successfully load balances multipath connections \cite{olivier_bonaventure_apple_2019}.
+Fortunately, several load balancing algorithms have been proposed whose aim is to make sure that connections, which use multiple paths using Multipath TCP or QUIC, stay on the same path \citep{olivier_bonaventure_multipath_2018,olteanu_datacenter_2016,lienardy_towards_2016}.
+Since I envision my solution to be based on Multipath TCP or QUIC load balancing should not be a problem.
+\textit{Apple Music}, for example, successfully load balances multipath connections \citep{olivier_bonaventure_apple_2019}.
 
-%JF: FIXME: Downside (if room left): it's not clear which element is the bottleneck link.
-If all of them are FQ or non-FW, then it's fine.
-Routing may result in other elements to be used that have not been probed.
-All routers along the path may need to provide periodic reports on queueing policy *and* on queue fill status, so the source can guess the bottleneck link.
-%JF: And one interesting observation (perhaps follow-up work, but likely discussed already): I'm curious how a series of FQ and non-FQ routers impacts on competing flows. 
-An alternative to our algorithm could be that all routers inform the endpoints if they support \gls{fq} or not.
+An alternative to my algorithm could be that all routers inform the endpoints if they support \gls{fq} or not.
 Such a solution could, for example, use an IP or a TCP extension, which every router uses to indicate if they're capable of \gls{fq}.
@@
 
-Concluding, we argue that our new flow startup method can help to increase the deployment of delay-based \gls{cc}.
+Concluding, I argue that my new flow startup method can help to increase the deployment of delay-based \gls{cc}.
 Especially for applications that are very delay-sensitive, such as video conferencing and cloud gaming, the adoption of delay-based \gls{cc} can result in a significant increase of \gls{qoe}.
-We demonstrated that our method works using a prototype implementation using PCC but we are also confident that a similar flow startup method can also be integrated with other \glspl{cca} such as BBR.
+I demonstrated that my method works using a prototype implementation using PCC but I are also confident that a similar flow startup method can also be integrated with other \glspl{cca} such as BBR.
 
-\bibliographystyle{splncs04}
+\renewcommand*{\bibfont}{\small}
+\bibliographystyle{ieeetr}
 \bibliography{fq_detection}
@@
 \end{document}
+\endinput

### 2021-02-15 paper/paper_ieee.tex
Tiny paper changes

@@
 \citep{kargar_bideh_tada_2016} propose a method to detect and distinguish certain popular \gls{aqm} mechanisms.
-However, unlike my work, it doesn't consider \gls{fq}.
+However, unlike our work, it doesn't consider \gls{fq}.
 \citep{baykal_detection_2017} propose a machine learning based approach to fingerprint \gls{aqm} algorithm.
@@
 \label{fig:goodput_fq}}
-\caption{An example illustrating my proposed flow startup mechanism.
+\caption{An example illustrating our proposed flow startup mechanism.
 Figure \ref{fig:throughput} shows the sending rate, \ref{fig:goodput_no_fq} the receiving rate in case there's no \gls{fq} and \ref{fig:goodput_fq} the receiving rate if there is \gls{fq}.}
@@
 It is necessary to wait for packet loss in both connections because only then it is certain that the link is saturated. Our algorithm assumes that the link is saturated. 
-\item If \gls{fq} is used, the $\textit{loss ratio}$ is 2, otherwise it is 1. I choose 1.5 as the cutoff value. 
+\item If \gls{fq} is used, the $\textit{loss ratio}$ is 2, otherwise it is 1. We choose 1.5 as the cutoff value. 
 \item Since the presence or absence of \gls{fq} is now known the appropriate \gls{cca} can be launched. 
@@
 Figure~\ref{fig:throughput}, \ref{fig:goodput_no_fq} and \ref{fig:goodput_fq} schematically illustrate this mechanism.
-I assume a sender being connected to a receiver, with a bottleneck link between them.
-I also assume that all flows between the source and the destination follow the same route.
+We assume a sender being connected to a receiver, with a bottleneck link between them.
+We also assume that all flows between the source and the destination follow the same route.
 On this bottleneck link, the queue is either managed by \gls{fq} or by a shared queue (no \gls{fq}).
@@
 \item If \textit{current rtt} $\leq$ \textit{smallest rtt ever measured on the connection} $+$ 5\textit{ms} then increase the sending rate by 1\%.
-I choose 5\,ms since a delay of 5\,ms is barely noticeable by humans.
+We choose 5\,ms since a delay of 5\,ms is barely noticeable by humans.
 This design choice was first established in \citep{nichols_controlling_2012}.
@@
 
-%This \gls{cca} is meant as a simple proof-of-concept to demonstrate the efficacy of my \gls{fq} detection mechanism but I do not claim it to be optimal in any way.
+%This \gls{cca} is meant as a simple proof-of-concept to demonstrate the efficacy of our \gls{fq} detection mechanism but we do not claim it to be optimal in any way.
 
@@
 If the absence of \gls{fq} is detected, a loss based \gls{cc} is used as a fallback.
-I use the algorithm \textit{PCC} proposed by \citep{dong_pcc_2015}.
+We use the algorithm \textit{PCC} proposed by \citep{dong_pcc_2015}.
 In simplified terms, this \gls{cca} aims to maximize a utility function, which strives for high throughput and low packet loss.
@@
 
-I base my implementation on the code of \textit{PCC Vivace} \citep{dong_pcc_2018}, which in turn is an improved version of PCC.
+We base our implementation on the code of \textit{PCC Vivace} \citep{dong_pcc_2018}, which in turn is an improved version of PCC.
 The PCC code is based on the \textit{UDT} library \citep{gu_udt_2007}, which is a library that provides a reliable transport protocol similar to TCP on top of UDP.
 This approach is similar to QUIC \citep{iyengar_quic_2018}.
-An advantage of implementing my approach on top of UDP as opposed to TCP is that it is easier to develop and modify since no kernel module needs to be created and that it is more secure since the code is isolated in a user space process.
+An advantage of implementing our approach on top of UDP as opposed to TCP is that it is easier to develop and modify since no kernel module needs to be created and that it is more secure since the code is isolated in a user space process.
 
@@
 However, multipath TCP \citep{handley_tcp_2012} and QUIC, for which there are proposals to enable recombining data of several flows \citep{de_coninck_multipath_2017}, are becoming increasingly prevalent.
-I thus argue that the actual obstacles to deployment of my solution are constantly decreasing since QUIC (mostly Google traffic) already accounts for around
+We thus argue that the actual obstacles to deployment of our solution are constantly decreasing since QUIC (mostly Google traffic) already accounts for around
 $10\%$ of all Internet traffic \citep{ruth_first_2018} and is growing. Moreover, Apple's macOS and iOS support multipath TCP \citep{apple_about_2018} support.
@@
 
-For the evaluation I use the network emulation library \texttt{py-virtnet}\footnote{\url{https://pypi.org/project/py-virtnet/}}.
+For the evaluation we use the network emulation library \texttt{py-virtnet}\footnote{\url{https://pypi.org/project/py-virtnet/}}.
 All experiments run on virtual hosts on one physical host, on which the necessary network components and virtual hosts are emulated using Linux network namespaces.
@@
 
-For the experiments with \gls{fq} I use the \texttt{fq} module for Linux.
+For the experiments with \gls{fq} we use the \texttt{fq} module for Linux.
 This module keeps one drop tail queue for each flow with a configurable maximum size of packets.
-However, the design of my \gls{fq} detection algorithm does not assume any particular queue management algorithm being used for the queue of each flow and thus is also compatible with more sophisticated \gls{aqm} mechanisms which were explicitly developed to be used in conjunction with \gls{fq} \citep{taht_flow_2018,hoiland-jorgensen_piece_2018,bachl_cocoa_2019,bachl_lfq_2020}.
+However, the design of our \gls{fq} detection algorithm does not assume any particular queue management algorithm being used for the queue of each flow and thus is also compatible with more sophisticated \gls{aqm} mechanisms which were explicitly developed to be used in conjunction with \gls{fq} \citep{taht_flow_2018,hoiland-jorgensen_piece_2018,bachl_cocoa_2019,bachl_lfq_2020}.
 It only assumes that each flow gets the same share of bandwidth and that for a flow, whose queue is growing faster because more packets of this flow arrive at the bottleneck, packets are dropped faster than for a flow of which fewer packets appear at the bottleneck link.
-For the experiments with a shared queue I use \texttt{pfifo}, which is a simple drop tail queue which can contain up to a fixed number of packets.
+For the experiments with a shared queue we use \texttt{pfifo}, which is a simple drop tail queue which can contain up to a fixed number of packets.
 
@@
 }
-\caption{Throughput and delay of a flow controlled by my delay-based \gls{cca} on a link with a speed of 10\,Mbps, a delay of 10\,ms and a buffer size of 100 packets.}
+\caption{Throughput and delay of a flow controlled by our delay-based \gls{cca} on a link with a speed of 10\,Mbps, a delay of 10\,ms and a buffer size of 100 packets.}
 \label{fig:just_one_flow_udp}
@@
 
-First, I demonstrate empirically that delay-based \gls{cc} actually has a benefit over loss-based one.
+First, we demonstrate empirically that delay-based \gls{cc} actually has a benefit over loss-based one.
 On one side, \autoref{fig:just_one_flow_tcp} shows that the loss-based \gls{cc} achieves a throughput of almost 10\,Mbps on a 10\,Mbps link, which means full utilization and an \gls{rtt} that is  higher than 100\,ms, which is not desirable.
-On the other side, \autoref{fig:just_one_flow_udp} shows an example flow using my simple delay-based \gls{cca} and a slow start algorithm similar to the one used by the classic TCP \gls{cca} \textit{New Reno} \citep{allman_tcp_1999}.
+On the other side, \autoref{fig:just_one_flow_udp} shows an example flow using our simple delay-based \gls{cca} and a slow start algorithm similar to the one used by the classic TCP \gls{cca} \textit{New Reno} \citep{allman_tcp_1999}.
 While the delay-based flow also achieves very high utilization, delay rarely exceeds 20\,ms, meaning that it is more than 5 times lower than the one of the loss-based \gls{cca} of \autoref{fig:just_one_flow_tcp}.
@@
 \subsection{Comparing flows with different Congestion Controls and AQM mechanisms}
-In this scenario, I show an example of a flow with my mechanism that shares a bottleneck link with a flow using a loss-based congestion control.
+In this scenario, we show an example of a flow with our mechanism that shares a bottleneck link with a flow using a loss-based congestion control.
 
 \subsubsection{No \gls{fq}}
-The examples in \autoref{fig:pfifo_tcp} and in \autoref{fig:pfifo_udp} show the loss-based \gls{cca} Cubic compete against my delay-based \gls{cca} on a link without \gls{fq}.
-It is important to note that I do not use my \gls{fq} detection mechanism here but instead choose delay-based \gls{cc} from the beginning to demonstrate the starvation of delay-based flows that happens when they compete against loss-based ones without \gls{fq}.
-I start the loss-based flow before to make sure that it has sufficient time to fill up the link: The loss-based flow starts 5 seconds earlier and after 5 seconds, the delay-based flow joins.
-I choose these 5 seconds so that the first flow can enter its stable state because I want to show how how the flow, even though it had the advantage of the 5 seconds in which it could use the link all for itself, loses to the loss-based \gls{cc}: While the delay-based flow gains some share of the link during startup, it is pushed away by the loss-based flow later on and ``starves''.
+The examples in \autoref{fig:pfifo_tcp} and in \autoref{fig:pfifo_udp} show the loss-based \gls{cca} Cubic compete against our delay-based \gls{cca} on a link without \gls{fq}.
+It is important to note that we do not use our \gls{fq} detection mechanism here but instead choose delay-based \gls{cc} from the beginning to demonstrate the starvation of delay-based flows that happens when they compete against loss-based ones without \gls{fq}.
+We start the loss-based flow before to make sure that it has sufficient time to fill up the link: The loss-based flow starts 5 seconds earlier and after 5 seconds, the delay-based flow joins.
+We choose these 5 seconds so that the first flow can enter its stable state because we want to show how how the flow, even though it had the advantage of the 5 seconds in which it could use the link all for itself, loses to the loss-based \gls{cc}: While the delay-based flow gains some share of the link during startup, it is pushed away by the loss-based flow later on and ``starves''.
 
@@
 }
-\caption{Throughput and delay of a flow controlled by my delay-based \gls{cca} on a link with a speed of 50\,Mbps, a delay of 10\,ms and a buffer size of 100 packets.
+\caption{Throughput and delay of a flow controlled by our delay-based \gls{cca} on a link with a speed of 50\,Mbps, a delay of 10\,ms and a buffer size of 100 packets.
 The bottleneck is controlled by a shared queue (no \gls{fq}) and is shared with the flow of \autoref{fig:pfifo_tcp}.}
@@
 }
-\caption{Throughput and delay of a flow controlled by my delay-based \gls{cca} on a link with a speed of 50\,Mbps, a delay of 10\,ms and a buffer size of 100 packets.
+\caption{Throughput and delay of a flow controlled by our delay-based \gls{cca} on a link with a speed of 50\,Mbps, a delay of 10\,ms and a buffer size of 100 packets.
 The bottleneck is controlled by \gls{fq} and is shared with the flow of \autoref{fig:fq_tcp}.}
@@
 
-For a link with fair queuing, the following example results show that throughput is identical for my flow (\autoref{fig:fq_udp}) and the Cubic flow (\autoref{fig:fq_tcp}), just that the Cubic flow's delay is more than two times higher than the delay-based one's.
-Thus, my mechanism detects that the bottleneck uses \gls{fq} and uses a delay-based congestion control because it is better suited in a scenario with \gls{fq}.
+For a link with fair queuing, the following example results show that throughput is identical for our flow (\autoref{fig:fq_udp}) and the Cubic flow (\autoref{fig:fq_tcp}), just that the Cubic flow's delay is more than two times higher than the delay-based one's.
+Thus, our mechanism detects that the bottleneck uses \gls{fq} and uses a delay-based congestion control because it is better suited in a scenario with \gls{fq}.
 
@@
 
-To evaluate if my proposed mechanism correctly recognizes \gls{fq} in a systematic manner, I perform experiments using a wide range of network configurations: I vary bandwidth from 5 to 50\,Mbps, delay from 10 to 100\,ms and the buffer size from 1 to 100 packets, using a grid of these parameters.
-For each parameter, I take 5 values, evenly spaced.
-Thus, I run experiments with 125 different configurations.
-I run each configuration once with \gls{fq} and once without.
-Results show that for experiments without \gls{fq} (using a shared queue), in over 99\% of cases, the absence of \gls{fq} was correctly detected and in less than 1\% of cases, my algorithm detected \gls{fq} even though there was no \gls{fq}.
+To evaluate if our proposed mechanism correctly recognizes \gls{fq} in a systematic manner, we perform experiments using a wide range of network configurations: we vary bandwidth from 5 to 50\,Mbps, delay from 10 to 100\,ms and the buffer size from 1 to 100 packets, using a grid of these parameters.
+For each parameter, we take 5 values, evenly spaced.
+Thus, we run experiments with 125 different configurations.
+We run each configuration once with \gls{fq} and once without.
+Results show that for experiments without \gls{fq} (using a shared queue), in over 99\% of cases, the absence of \gls{fq} was correctly detected and in less than 1\% of cases, our algorithm detected \gls{fq} even though there was no \gls{fq}.
 For experiments, during which \gls{fq} was actually deployed on the bottleneck link, this was correctly detected in 97\% of cases and wrongly in 3\% of the time.
-The overall accuracy is thus 98\%, which I consider sufficient.
+The overall accuracy is thus 98\%, which we consider sufficient.
 
-To verify the behavior of my algorithm under the presence of cross traffic, I repeated the above experiments but with one bulk traffic flow using Cubic running at the same time, which I start 5 seconds before so that it has sufficient time to gain full bandwidth.
-Also under these circumstances I also achieve 98\% detection accuracy using my algorithm.
-I investigated the reason for the detection accuracy not being 100\% and found the following: The PCC code relies on packet scheduling using timers, which is very computationally expensive because it needs a timer for every packet and thus the program has to be put to sleep/woken up repeatedly by the operating system.
+To verify the behavior of our algorithm under the presence of cross traffic, we repeated the above experiments but with one bulk traffic flow using Cubic running at the same time, which we start 5 seconds before so that it has sufficient time to gain full bandwidth.
+Also under these circumstances we also achieve 98\% detection accuracy using our algorithm.
+We investigated the reason for the detection accuracy not being 100\% and found the following: The PCC code relies on packet scheduling using timers, which is very computationally expensive because it needs a timer for every packet and thus the program has to be put to sleep/woken up repeatedly by the operating system.
 Thus, if processor load is high, the timers fail and PCC can't send as many packets as it would want to.
-My algorithm relies on flow 1 sending half the rate of flow 2 but this cannot always be guaranteed if CPU load is high.
-I tried to minimize all tasks which run in the background and achieved the accuracy I got.
-When I ran the experiments while I was browsing the web, I got lower accuracy.
+our algorithm relies on flow 1 sending half the rate of flow 2 but this cannot always be guaranteed if CPU load is high.
+We tried to minimize all tasks which run in the background and achieved the accuracy we got.
+When we ran the experiments while we was browsing the web, we got lower accuracy.
 Solutions to the problem are to not schedule packets using timers or to implement the algorithm not in user space but in kernel space.
@@
 
-Besides the examples I provided, I also conducted a more systematic comparison of the throughput and delay that my algorithm achieves: I use the same scenario as above and again vary bandwidth from 5 to 50\,Mbps, delay from 10 to 100\,ms and the buffer size from 1 to 100 packets.
-For each parameter, I take 5 values, evenly spaced, totaling 125 distinct scenarios.
-I use \gls{fq} at the bottleneck link.
-I run each scenario both using Cubic as well as my algorithm for 30 seconds each.
-I average the achieved throughputs and average delays from all experiments for both Cubic as well as my algorithm.
-For Cubic, the mean throughput is 21\,Mbit/s and the mean \gls{rtt} is 96\,ms while for my algorithm it is 26\,Mbit/s and 64\,ms.
-This shows that my algorithm can correctly determine that there is \gls{fq} and switch to the delay-based \gls{cc} and that my algorithm achieves significantly lower delay (loss-based Cubic's delay is 50\% higher) and -- surprisingly -- higher throughput.
+Besides the examples we provided, we also conducted a more systematic comparison of the throughput and delay that our algorithm achieves: we use the same scenario as above and again vary bandwidth from 5 to 50\,Mbps, delay from 10 to 100\,ms and the buffer size from 1 to 100 packets.
+For each parameter, we take 5 values, evenly spaced, totaling 125 distinct scenarios.
+We use \gls{fq} at the bottleneck link.
+We run each scenario both using Cubic as well as our algorithm for 30 seconds each.
+We average the achieved throughputs and average delays from all experiments for both Cubic as well as our algorithm.
+For Cubic, the mean throughput is 21\,Mbit/s and the mean \gls{rtt} is 96\,ms while for our algorithm it is 26\,Mbit/s and 64\,ms.
+This shows that our algorithm can correctly determine that there is \gls{fq} and switch to the delay-based \gls{cc} and that our algorithm achieves significantly lower delay (loss-based Cubic's delay is 50\% higher) and -- surprisingly -- higher throughput.
 This is astonishing considering that loss-based \glspl{cca} such as Cubic should be more aggressive in filling the link.
-When analyzing the results I identified that Cubic performs very poorly in scenarios with small buffers because the multiplicative decrease of Cubic is 0.7, meaning that the bandwidth is reduced by 30\% after every packet loss while my algorithm only reduces it by only 5\%.
+When analyzing the results we identified that Cubic performs very poorly in scenarios with small buffers because the multiplicative decrease of Cubic is 0.7, meaning that the bandwidth is reduced by 30\% after every packet loss while our algorithm only reduces it by only 5\%.
 Thus Cubic reduces bandwidth too much after packet loss which results in underutilization of the link.
@@
 
-During the implementation of my algorithm I noticed peculiar behavior when not using \gls{fq}: During startup, I make flow 1 have half the sending rate of flow 2.
-I would expect the receiving rates to be the same: flow 1 achieving around half the rate of flow 2.
-However, I noticed that usually flow 1 would have a receiving rate of 0.
-After investigating the issue I noticed that the reason was pacing: PCC (and other \glspl{cca}) such as BBR send packets in regular intervals, for example every millisecond, because sending regularly keeps the queues shorter.
+During the implementation of our algorithm we noticed peculiar behavior when not using \gls{fq}: During startup, we make flow 1 have half the sending rate of flow 2.
+We would expect the receiving rates to be the same: flow 1 achieving around half the rate of flow 2.
+However, we noticed that usually flow 1 would have a receiving rate of 0.
+After investigating the issue we noticed that the reason was pacing: PCC (and other \glspl{cca}) such as BBR send packets in regular intervals, for example every millisecond, because sending regularly keeps the queues shorter.
 This behavior is called pacing.
@@
 Then, when the packets arrive at the bottleneck, flow 2's packet, which arrives a bit earlier, fills up the buffer and flow 1's packet is consequently always dropped.
-As a solution, I do not send packets at regular intervals during startup but always add a small random value: For example, if the sending interval were 1\,ms, I would send the packet randomly in the interval from 0.5 to 1.5\,ms.
+As a solution, we do not send packets at regular intervals during startup but always add a small random value: For example, if the sending interval were 1\,ms, we would send the packet randomly in the interval from 0.5 to 1.5\,ms.
 This solves the aforementioned problem.
-The insight I gained from this is that while pacing is generally supportive, it is actually harmful if several flows compete at a bottleneck without \gls{fq} and one flow's sending rate is a multiple of the other's.
+The insight we gained from this is that while pacing is generally supportive, it is actually harmful if several flows compete at a bottleneck without \gls{fq} and one flow's sending rate is a multiple of the other's.
 
-One problem of my proposed solution occurs when the bottleneck link changes while a flow is running.
+One problem of our proposed solution occurs when the bottleneck link changes while a flow is running.
 For example, the bottleneck link could be the home router in the beginning, while later the bottleneck link could be another router deeper in the core network.
-The problem occurs if the home router supports \gls{fq} and my algorithm detects this while the router in the core network, which becomes the bottleneck later on, doesn't support \gls{fq}.
-Then, my algorithm is going to use a delay-based \gls{cc} because that's what it determined to be correct in the beginning of the flow.
-However, later when the bottleneck changes, my algorithm is still using delay-based \gls{cc} even though the current bottleneck doesn't support \gls{fq} anymore.
-As a result, it might be possible that my algorithm uses the wrong \gls{cc}.
+The problem occurs if the home router supports \gls{fq} and our algorithm detects this while the router in the core network, which becomes the bottleneck later on, doesn't support \gls{fq}.
+Then, our algorithm is going to use a delay-based \gls{cc} because that's what it determined to be correct in the beginning of the flow.
+However, later when the bottleneck changes, our algorithm is still using delay-based \gls{cc} even though the current bottleneck doesn't support \gls{fq} anymore.
+As a result, it might be possible that our algorithm uses the wrong \gls{cc}.
 One way to prevent this problem from occurring is to use periodic measurements.
-For example, my algorithm to determine \gls{fq} could be used every 5 or 10 seconds to measure again if the current bottleneck supports \gls{fq}.
+For example, our algorithm to determine \gls{fq} could be used every 5 or 10 seconds to measure again if the current bottleneck supports \gls{fq}.
 
-Another scenario that is worth noting is how my proposed solution deals with load balancing: In this research I assume that both of my flows, which I use during startup to determine the presence of \gls{fq} take the same path to the host on the other side.
+Another scenario that is worth noting is how our proposed solution deals with load balancing: In this research we assume that both of our flows, which we use during startup to determine the presence of \gls{fq} take the same path to the host on the other side.
 However, with load balancing it could happen that flow 1 takes a different path than flow 2.
 Fortunately, several load balancing algorithms have been proposed whose aim is to make sure that connections, which use multiple paths using Multipath TCP or QUIC, stay on the same path \citep{olivier_bonaventure_multipath_2018,olteanu_datacenter_2016,lienardy_towards_2016}.
-Since I envision my solution to be based on Multipath TCP or QUIC load balancing should not be a problem.
+Since we envision our solution to be based on Multipath TCP or QUIC load balancing should not be a problem.
 \textit{Apple Music}, for example, successfully load balances multipath connections \citep{olivier_bonaventure_apple_2019}.
 
-An alternative to my algorithm could be that all routers inform the endpoints if they support \gls{fq} or not.
+An alternative to our algorithm could be that all routers inform the endpoints if they support \gls{fq} or not.
 Such a solution could, for example, use an IP or a TCP extension, which every router uses to indicate if they're capable of \gls{fq}.
@@
 
-Concluding, I argue that my new flow startup method can help to increase the deployment of delay-based \gls{cc}.
+Concluding, we argue that our new flow startup method can help to increase the deployment of delay-based \gls{cc}.
 Especially for applications that are very delay-sensitive, such as video conferencing and cloud gaming, the adoption of delay-based \gls{cc} can result in a significant increase of \gls{qoe}.
-I demonstrated that my method works using a prototype implementation using PCC but I are also confident that a similar flow startup method can also be integrated with other \glspl{cca} such as BBR.
+We demonstrated that our method works using a prototype implementation using PCC but we are also confident that a similar flow startup method can also be integrated with other \glspl{cca} such as BBR.
 

### 2021-02-18 paper/paper_ieee.tex
Add related work

@@
 \usepackage{float}
+
+\usepackage[OT1]{fontenc}
 
@@
 
-\subsubsection{\gls{cc} using Active Measurements}
+\subsection{\gls{cc} using Active Measurements}
 Several approaches have been recently proposed which aim to incorporate active measurements into \glspl{cca}.
@@
 
-\subsubsection{Flow startup techniques} Besides the classical TCP slow start algorithm \citep{stevens_tcp_1997} some other proposals were made such as Quick Start \citep{jain_quick-start_2019}, which aims to speed up flow start by routers informing the end point, which rate they deem appropriate.
+\subsection{Flow startup techniques} Besides the classical TCP slow start algorithm \citep{stevens_tcp_1997} some other proposals were made such as Quick Start \citep{jain_quick-start_2019}, which aims to speed up flow start by routers informing the end point, which rate they deem appropriate.
 \citep{jarvinen_congestion_2019} investigate the impact of \gls{aqm} and \gls{cc} during flow startup.
@@
 %JF: Evtl.statt subsubsection Absätze mit Titel fett verwenden um doch einigen Platz zu sparen?
-\subsubsection{Detecting \gls{aqm} mechanisms} There are few publications that investigate methods to detect the presence of an \gls{aqm}.
+\subsection{Detecting \gls{aqm} mechanisms} There are few publications that investigate methods to detect the presence of an \gls{aqm}.
 \citep{kargar_bideh_tada_2016} propose a method to detect and distinguish certain popular \gls{aqm} mechanisms.
@@
 \gls{fq} is not considered, too.
+
+\subsection{Detecting the presence of a shared bottleneck} Several approaches have been proposed to detect whether in a set of flows, some of them share a common bottleneck \cite{hayes_practical_2014,ferlin_revisiting_2016,hayes_online_2020}.
+These approaches typically operate by comparing some statistics of flows to check whether they correlate between flows.
+For example, it can be checked if throughput correlates between two flows.
+If throughput correlates, it can mean that these flows share a common bottleneck link, and if additional flows join at the bottleneck link, all other flows' throughputs go down, which is the reason why they are correlated.
+While approaches to detect a shared bottleneck link appear to be similar to what this paper is aiming to do, there are differences: Two flows can share a common bottleneck link but the bottleneck link can have fair queuing.
+Thus the aims of shared bottleneck and fair queuing detection are different.
 
@@
 This design choice was first established in \citep{nichols_controlling_2012}.
-\item Otherwise (\textit{current rtt} > \textit{smallest rtt ever measured on the connection} + 5\textit{ms}) decrease the sending rate by 5\%.
+\item Otherwise (\textit{current rtt} $>$ \textit{smallest rtt ever measured on the connection} + 5\textit{ms}) decrease the sending rate by 5\%.
 \end{enumerate}
 
-%This \gls{cca} is meant as a simple proof-of-concept to demonstrate the efficacy of our \gls{fq} detection mechanism but we do not claim it to be optimal in any way.
+This \gls{cca} is meant as a simple proof-of-concept to demonstrate the efficacy of our \gls{fq} detection mechanism but we do not claim it to be optimal in any way.
 
@@
 This approach is similar to QUIC \citep{iyengar_quic_2018}.
-An advantage of implementing our approach on top of UDP as opposed to TCP is that it is easier to develop and modify since no kernel module needs to be created and that it is more secure since the code is isolated in a user space process.
+An advantage of implementing our approach on top of UDP as opposed to TCP is that it is easier to develop and modify since no kernel module needs to be created and that it is more secure since the code is isolated in a user space process and most attacks can only influence this process and not the entire kernel.
 
